---
title: "Stage DE figs"
date: "2024-01-05"
author: "Jesse Rop"
output:
  html_document:
    toc: true
    theme: united
    code_folding: hide
    df_print: paged
---

```{r}
suppressPackageStartupMessages({
  library(Seurat)
  # library(sctransform)
  # library(rgl)
  library(ggrepel)
  # library(ggpubr)
  library(dplyr) 
  library(stringr)
  library(purrr)
  library(readr)
  library(tibble)
  library(ggplot2)
  # library(ggnewscale)
  library(tidyr)
  library(cowplot)
  # library(scDblFinder)
  library(SeuratData)
  library(SeuratDisk)
  # library(scmap)
  library(SingleCellExperiment)
  library(pheatmap)
  library(plotly)
  # library(Hmisc)
  # library(rmarkdown)
  library(RColorBrewer)
  # library(MAST)
  library(ggvenn)
  # library("biomaRt")
  library(conflicted)
  # library(SoupX)
  library(MetBrewer)
  library(dreamlet)
  library(muscat)
  # library(ExperimentHub)
  # library(zenith)
  library(scater)
  library(slingshot)
  library(harmony)
  library(ggforce)
  # library(miloR)
  library(scater)
  library(scran)
  library(patchwork)
  # library(scRNAseq)
  library(scuttle)
  library(irlba)
  library(BiocParallel)
  library(tradeSeq)
  library(ComplexHeatmap)
  library(circlize)
  library(readxl)
  library(ggh4x)
  library(EnhancedVolcano)
  library(GGally)
  
  conflict_prefer("select", "dplyr")
  conflict_prefer("filter", "dplyr")
  conflict_prefer("rename", "dplyr")
  conflicts_prefer(dplyr::count)
  conflicts_prefer(base::as.data.frame)
  conflicts_prefer(base::as.factor)
  
})
```


```{r setup}
sample_set = "Mali2"

# knitr::opts_knit$set(root.dir = paste0("/Users/jr35/Google Drive/My Drive/PHD_documents/malaria_phd/", sample_set))
knitr::opts_knit$set(root.dir = paste0("/lustre/scratch126/tol/teams/lawniczak/users/jr35/phd/", sample_set))
```


#  integration of all MSC

### Common functions

#### Source commong functions, varriables and tables
```{r}
##Boxplot labeling function
source("../multipurpose_scripts_lnk/seurat_sce/seu_sce_common_fns.R")
source("../multipurpose_scripts_lnk/plotting_common_fns.R")
source("/nfs/users/nfs_j/jr35/Mali2/mali22_code/pf_common_vars.R")
```


```{r}
read_dir_scvi2seu <- "data/processed/Pf/m2_all/scvi2seu/"
read_dir_stage_DE_markers <- "data/processed/Pf/m2_all/stage_DE_markers/"
read_dir_integration <- "data/processed/Pf/m2_all/integration/"
save_dir_stage_DE_markers <- "data/processed/Pf/m2_all/stage_DE_markers/"
published_dset_dir <- "/lustre/scratch126/tol/teams/lawniczak/users/jr35/phd/published_dsets/"


```


#####!!! DE below in seurat. Already done using scvi. Not sure which one to go with

```{r}
# # Generated in "scvi_leiden_clusters_seuratDE.Rmd"
# all_jcC_seu_scvi_stg_noM55_afm_cln <- readRDS(paste0(read_dir_stage_DE_markers, 'all_jcC_seu_scvi_stg_noM55_afm_cln.RDS'))

# Generated in "scvi_leiden_clusters_doublet_investigations.Rmd"
all_jcC_seu_scvi_noM55_afm_cln <- readRDS(paste0(read_dir_integration, 'all_jcC_seu_scvi_noM55_afm_cln.RDS'))

```


```{r}
# Generated in 'scvi_leiden_clusters_seurat_pbulkDE_n_DE_comparisons'
de_wilcox_bulk_scvi <- readRDS(paste0(read_dir_stage_DE_markers, 'de_wilcox_bulk_scvi.RDS'))

```

```{r, warning=FALSE, message=FALSE}
# Generated in 'fem_subclusters_DE.Rmd'
seu_scvi_fem_sub_clusters <- map(c("f"), ~readRDS(paste0(save_dir_stage_DE_markers, 'seu_scvi_fem_sub_clusters_', .x,'.RDS')))

```

```{r}
map(seu_scvi_fem_sub_clusters, ~DimPlot(.x, group.by = "fem_subclusters"))

```

```{r}
map(seu_scvi_fem_sub_clusters, ~DimPlot(.x, group.by = "fem_subclusters") + scale_color_manual(values = sp_fld_colors2)) 

```


```{r}
# seu_scvi_fem_sub_clusters <- list(seu_scvi_fem_sub_clusters)

seu_scvi_fem_sub_clusters <- map(seu_scvi_fem_sub_clusters, ~SetIdent(.x, value = "fem_subclusters"))

```

##### DE on active female, active female LE, innert female, innert female LE compared versus the rest using pseudobulk DE 

```{r, warning=FALSE, message=FALSE}
## Remove pseudobulk groups with less than 50 cells - !!NOTE seems I had alread removed underepresented groups with less than 20 previously when doing pseudobulk DE 
map(seu_scvi_fem_sub_clusters, ~.x@meta.data %>% group_by(fem_subclusters, orig.ident) %>% count())

seu_scvi_fem_sub_clusters_cln4pbulkDE <- map(seu_scvi_fem_sub_clusters, ~.x[, .x@meta.data %>% rownames_to_column('bcode') %>% group_by(fem_subclusters, orig.ident) %>% filter(n() > 20) %>% pull(bcode)]) 
```

```{r, warning=FALSE, message=FALSE}
## Remove leiden clusters with less than 2 donors
map(seu_scvi_fem_sub_clusters_cln4pbulkDE, ~ .x@meta.data %>% group_by(fem_subclusters) %>% filter(n_distinct(orig.ident) < 2) %>% count(fem_subclusters, orig.ident))


seu_scvi_fem_sub_clusters_cln4pbulkDE <- map(seu_scvi_fem_sub_clusters_cln4pbulkDE, ~.x[, .x@meta.data %>% rownames_to_column('bcode') %>% group_by(fem_subclusters) %>% filter(n_distinct(orig.ident) >= 2) %>% pull(bcode)]) 
```

Retain only those genes expressed in atleast 20 cells for DE
```{r, warning=FALSE, message=FALSE}
seu_scvi_fem_sub_clusters_cln4pbulkDE <- map(seu_scvi_fem_sub_clusters_cln4pbulkDE, ~{
  
  counts <- GetAssayData(.x, slot = "counts")
  
  keep_genes <- rowSums(counts > 0) >= 20   # expressed in ≥10 cells
  # seurat_obj <- subset(seurat_obj, features = rowSums(GetAssayData(seurat_obj, slot="counts") > 0) > 10)
  
  subset(.x, features = rownames(counts)[keep_genes])
  
})
```


```{r, warning=FALSE, message=FALSE}
## Count the sufficient groups
map(seu_scvi_fem_sub_clusters_cln4pbulkDE, ~.x@meta.data %>% group_by(fem_subclusters, orig.ident) %>% count())

```



```{r, warning=FALSE, message=FALSE}
# Save for plotting
imap(seu_scvi_fem_sub_clusters_cln4pbulkDE, ~saveRDS(.x, paste0(save_dir_stage_DE_markers, 'seu_scvi_fem_sub_clusters_cln4pbulkDE_', .y,'.RDS')))

```

```{r}
# Get UMAPs and metadata for plotting
seu_scvi_fem_sub_clusters_cln4pbulkDE_umap_mdata <- map(seu_scvi_fem_sub_clusters_cln4pbulkDE, ~{
  .x@meta.data %>% 
    rownames_to_column("bcode") %>%
    left_join(., .x@reductions$scvi_umap@cell.embeddings %>% data.frame() %>% rownames_to_column("bcode"), by = "bcode") 
})
```

```{r, warning=FALSE, message=FALSE}
# Save for plotting
imap(seu_scvi_fem_sub_clusters_cln4pbulkDE_umap_mdata, ~saveRDS(.x, paste0(save_dir_stage_DE_markers, 'seu_scvi_fem_sub_clusters_cln4pbulkDE_umap_mdata_', .y,'.RDS')))

```

```{r, warning=FALSE, message=FALSE}
# pseudobulk by donor and leiden
pseudo_fem_hiresfem <- map(seu_scvi_fem_sub_clusters_cln4pbulkDE, ~AggregateExpression(.x, assays = "RNA", return.seurat = T, group.by = c("orig.ident", "fem_hires_subclusters")))

```

```{r, warning=FALSE, message=FALSE}
# Save for plotting
imap(pseudo_fem_hiresfem, ~saveRDS(.x, paste0(save_dir_stage_DE_markers, 'pseudo_fem_hiresfem_', .y,'.RDS')))

```

```{r, warning=FALSE, message=FALSE}
# pseudo_fem <- map(pseudo_fem, ~.x@meta.data %>%
#       mutate(cluster_condition = paste(fem_subclusters, condition)) %>%
#       select(cluster_condition) %>%
#       AddMetaData(.x, .)
# )

```


```{r, warning=FALSE, message=FALSE}
map(pseudo_fem_hiresfem, ~.x@meta.data %>% count(fem_hires_subclusters, orig.ident))

```

### Pseudobulk DE between each female subcluster and all other female subclusters

```{r, warning=FALSE, message=FALSE}
pseudo_fem_hiresfem <- map(pseudo_fem_hiresfem, ~SetIdent(.x, value = "fem_hires_subclusters"))

fem_hires_subclusters_bulk_de <- map(pseudo_fem_hiresfem, ~FindAllMarkers(object = .x, test.use = "DESeq2"))
```


```{r, warning=FALSE, message=FALSE}
fem_hires_subclusters_bulk_de_anot <- map(fem_hires_subclusters_bulk_de, possibly(~.x %>% 
                                                                                    # filter(p_val_adj < 0.05) %>% 
                                                                                    group_by(cluster) %>% arrange(p_val_adj, .by_group = T) %>% 
                                                                                    left_join(., Pf3D7_genes_plasmoDB[,c("gene_id", "Gene")], by = c("gene" = "gene_id"))))

fem_hires_subclusters_bulk_de_anot
```

```{r}
# Save for plotting
saveRDS(fem_hires_subclusters_bulk_de_anot, paste0(save_dir_stage_DE_markers, 'fem_hires_subclusters_bulk_de_anot.RDS'))

```


```{r, warning=FALSE, message=FALSE}
map(fem_hires_subclusters_bulk_de_anot, possibly(~.x %>% 
                                                   filter(p_val_adj < 0.05 & avg_log2FC > 0) %>% 
                                                   # filter(p_val_adj < 0.05) %>% 
                                                   arrange(desc(avg_log2FC)) %>% 
                                                   # arrange(desc(abs(avg_log2FC))) %>% 
                                                   group_by(cluster) %>% 
                                                   slice_min(order_by = p_val_adj, n = 20) %>% 
                                                   slice_max(avg_log2FC, n =10)))

```


```{r, warning=FALSE, message=FALSE}
map(fem_hires_subclusters_bulk_de_anot, possibly(~.x %>% 
                                                   filter(p_val_adj < 1e-8 & avg_log2FC > 0) %>% 
                                                   arrange(desc(abs(avg_log2FC))) %>% 
                                                   group_by(cluster) %>% 
                                                   slice_min(order_by = p_val_adj, n = 40) %>% 
                                                   slice_min(p_val_adj, n =10)))

```

```{r, fig.width=15, fig.height=3}
map(seu_scvi_fem_sub_clusters, ~FeaturePlot(.x, features = c("PF3D7-1006500", "PF3D7-0210900", "PF3D7-1421700",  "PF3D7-1214300"), reduction = "scvi_umap", order = T, ncol = 4, ))

```

```{r}
map(seu_scvi_fem_sub_clusters, ~VlnPlot(.x, features = "nFeature_RNA", group.by = "leiden", pt.size = 0))
map(seu_scvi_fem_sub_clusters, ~FeaturePlot(.x, features = "nFeature_RNA", reduction = "scvi_umap", label = T))
map(seu_scvi_fem_sub_clusters, ~DimPlot(.x, group.by = "leiden", reduction = "scvi_umap", label = T))

```


```{r}
map(seu_scvi_fem_sub_clusters, ~FeaturePlot(.x, features = c("PF3D7-1471900", "PF3D7-1323000"), reduction = "scvi_umap", label = F, order = F))
# map(seu_scvi_fem_sub_clusters, ~FeaturePlot(.x, features = "PF3D7-1141900", reduction = "scvi_umap", label = T))

```

```{r, warning=FALSE, message=FALSE}
top_de_tbl <- map(fem_hires_subclusters_bulk_de_anot, ~.x %>% 
                    # filter(gene %in% gns) %>% 
                    arrange(desc(avg_log2FC)) %>% 
                    filter(p_val_adj< 0.05 & avg_log2FC>0) %>%
                    group_by(cluster) %>% 
                    slice_min(order_by = p_val_adj, n = 40) %>% 
                    slice_max(order_by = avg_log2FC, n = 20) %>%
                    ungroup() %>%
                    mutate(across(cluster, as.character))
)

top_de_tbl
```

###### pairwise DE between the Female sub clusters including Female2 and Female4 which seem to be good quality mature terminal gametocytes that are not activated
```{r, warning=FALSE, message=FALSE}
# # pseudobulk by donor and leiden
# pseudo_fem_hiresfem <- map(seu_scvi_fem_sub_clusters_cln4pbulkDE, ~AggregateExpression(.x, assays = "RNA", return.seurat = T, group.by = c("orig.ident", "fem_hires_subclusters")))

```

```{r, warning=FALSE, message=FALSE}
# leiden_combis <- combn(c(2,3,4,8), m = 2, simplify = F)
fem_clusters_combis <- map(seu_scvi_fem_sub_clusters_cln4pbulkDE, ~combn(c(unique(.x$fem_hires_subclusters)), m = 2, simplify = F))

```

```{r, warning=FALSE, message=FALSE}
pseudo_fem_hiresfem <- map(pseudo_fem_hiresfem, ~SetIdent(.x, value = "fem_hires_subclusters"))

fem_hires_subclusters_bulk_pwDE <- map2(pseudo_fem_hiresfem, fem_clusters_combis, ~{
  obj = .x
  combis = .y
  map(combis, ~FindMarkers(obj, ident.1 = .x[1], ident.2 = .x[2], test.use = "DESeq2" )) %>% 
    set_names(map(combis, ~paste(., collapse = "vs")))
  
}) %>%
  unlist(., recursive = F)

# top10gns <- map(pseudo_fem_hiresfem, ~.x %>% map(.,\(obj) obj %>% ungroup() %>% select(Gene, gene) %>% deframe()))

```

```{r, warning=FALSE, message=FALSE}
fem_hires_subclusters_bulk_pwDE_anot <- map(fem_hires_subclusters_bulk_pwDE, ~.x %>% rownames_to_column("gene") %>% left_join(., Pf3D7_genes_plasmoDB[,c("gene_id", "Gene")], by = c("gene" = "gene_id")))

fem_hires_subclusters_bulk_pwDE_anot
```

```{r}
# Save for plotting
saveRDS(fem_hires_subclusters_bulk_pwDE_anot, paste0(save_dir_stage_DE_markers, 'fem_hires_subclusters_bulk_pwDE_anot.RDS'))

```


```{r, warning=FALSE, message=FALSE}
map(fem_hires_subclusters_bulk_pwDE_anot, ~.x %>% 
      slice_min(order_by = p_val_adj, n = 30) %>% 
      slice_max(order_by = avg_log2FC, n = 10))
```

```{r, warning=FALSE, message=FALSE}
map(fem_hires_subclusters_bulk_pwDE_anot, ~.x %>%  slice_min(order_by = p_val_adj, n = 30) %>%  slice_min(order_by = avg_log2FC, n = 10))
```

```{r, warning=FALSE, message=FALSE}
map(fem_hires_subclusters_bulk_pwDE_anot, ~.x %>%  slice_min(order_by = avg_log2FC, n = 30) %>% slice_min(order_by = p_val_adj, n = 10) )
```

```{r}
map(seu_scvi_fem_sub_clusters, ~FeaturePlot(.x, features = c("PF3D7-0417500", "PF3D7-1464400"), reduction = "scvi_umap", label = F, order = T, slot = "counts"))
map(seu_scvi_fem_sub_clusters, ~FeaturePlot(.x, features = c("PF3D7-0417500", "PF3D7-1464400"), reduction = "scvi_umap", label = F, order = T))

```


```{r}
map(seu_scvi_fem_sub_clusters, ~VlnPlot(.x, features = c("PF3D7-0417500", "PF3D7-1464400"), group.by = "fem_hires_subclusters", slot = "counts"))
map(pseudo_fem_hiresfem, ~VlnPlot(.x, features = c("PF3D7-0417500", "PF3D7-1464400"), group.by = "fem_hires_subclusters", slot = "counts"))

map(seu_scvi_fem_sub_clusters, ~VlnPlot(.x, features = c("PF3D7-0417500", "PF3D7-1464400"), group.by = "fem_hires_subclusters"))
map(pseudo_fem_hiresfem, ~VlnPlot(.x, features = c("PF3D7-0417500", "PF3D7-1464400"), group.by = "fem_hires_subclusters"))

```


```{r, fig.height=6, fig.width=6}
# vlcano_mrg <- map(c("bulkDseq", "seuWilcox"), ~{
#   EnhancedVolcano(de_wilcox_bulk_scvi %>% filter(cluster == "8"),
#                 lab = de_wilcox_bulk_scvi %>% filter(cluster == "8") %>% pull(!!rlang::sym(paste0("Gene_", .x))),

vlcano_actvFem <- imap(fem_hires_subclusters_bulk_pwDE_anot, ~EnhancedVolcano(.x ,
                                                                              lab = .x %>% pull(!!rlang::sym(paste0("Gene"))),
                                                                              # selectLab = .y,
                                                                              x = paste0("avg_log2FC"),
                                                                              y = paste0("p_val_adj"),
                                                                              pCutoff = 1e-10,
                                                                              FCcutoff = 1.2,
                                                                              pointSize = 1.2,
                                                                              labSize = 4.0,
                                                                              title = .y,
                                                                              subtitle = "",
                                                                              titleLabSize = 20,
                                                                              subtitleLabSize = 0,
                                                                              axisLabSize = 22,
                                                                              drawConnectors = TRUE,
                                                                              # boxedLabels = TRUE,
                                                                              # legendLabels =c( "p_val_adj"),
                                                                              gridlines.minor = FALSE
) +
  # labs(title = .y)+
  theme(legend.position = "none")
) 

vlcano_actvFem

```


###### Pairwise DE between female subclusters with cluster Female2 and Female4 combined

```{r, warning=FALSE, message=FALSE}
# pseudobulk the counts based on donor-female subclusters
pseudo_fem <- map(seu_scvi_fem_sub_clusters_cln4pbulkDE, ~AggregateExpression(.x, assays = "RNA", return.seurat = T, group.by = c("orig.ident", "fem_subclusters")))

```


```{r, warning=FALSE, message=FALSE}
# Save for plotting
imap(pseudo_fem, ~saveRDS(.x, paste0(save_dir_stage_DE_markers, 'pseudo_fem_', .y,'.RDS')))

```

```{r, warning=FALSE, message=FALSE}
# leiden_combis <- combn(c(2,3,4,8), m = 2, simplify = F)
fem_clusters_combis <- map(seu_scvi_fem_sub_clusters_cln4pbulkDE, ~combn(c(unique(.x$fem_subclusters)), m = 2, simplify = F))

```

```{r, warning=FALSE, message=FALSE}
pseudo_fem_pwDE <- map(pseudo_fem, ~SetIdent(.x, value = "fem_subclusters"))

fem_subclusters_bulk_pwDE <- map2(pseudo_fem_pwDE, fem_clusters_combis, ~{
  obj = .x
  combis = .y
  map(combis, ~FindMarkers(obj, ident.1 = .x[1], ident.2 = .x[2], test.use = "DESeq2" )) %>% 
    set_names(map(combis, ~paste(., collapse = "vs")))
  
}) %>%
  unlist(., recursive = F)

# top10gns <- map(pseudo_fem_hiresfem, ~.x %>% map(.,\(obj) obj %>% ungroup() %>% select(Gene, gene) %>% deframe()))

```

```{r, warning=FALSE, message=FALSE}
fem_subclusters_bulk_pwDE_anot <- map(fem_subclusters_bulk_pwDE, ~.x %>% rownames_to_column("gene") %>% left_join(., Pf3D7_genes_plasmoDB[,c("gene_id", "Gene")], by = c("gene" = "gene_id")))

fem_subclusters_bulk_pwDE_anot
```

```{r}
# Save for plotting
saveRDS(fem_subclusters_bulk_pwDE_anot, paste0(save_dir_stage_DE_markers, 'fem_subclusters_bulk_pwDE_anot.RDS'))

```


```{r, warning=FALSE, message=FALSE}
map(fem_subclusters_bulk_pwDE_anot, ~.x %>% 
      slice_min(order_by = p_val_adj, n = 30) %>% 
      slice_max(order_by = avg_log2FC, n = 10))
```

```{r, warning=FALSE, message=FALSE}
map(fem_subclusters_bulk_pwDE_anot, ~.x %>%  slice_min(order_by = p_val_adj, n = 30) %>%  slice_min(order_by = avg_log2FC, n = 10))
```

```{r, warning=FALSE, message=FALSE}
map(fem_subclusters_bulk_pwDE_anot, ~.x %>%  slice_min(order_by = avg_log2FC, n = 30) %>% slice_min(order_by = p_val_adj, n = 10) )
```

```{r}
map(seu_scvi_fem_sub_clusters, ~FeaturePlot(.x, features = c("PF3D7-0417500", "PF3D7-1464400"), reduction = "scvi_umap", label = F, order = T, slot = "counts"))
map(seu_scvi_fem_sub_clusters, ~FeaturePlot(.x, features = c("PF3D7-0417500", "PF3D7-1464400"), reduction = "scvi_umap", label = F, order = T))

```


```{r}
map(seu_scvi_fem_sub_clusters, ~VlnPlot(.x, features = c("PF3D7-0417500", "PF3D7-1464400"), group.by = "fem_subclusters", slot = "counts"))
map(pseudo_fem, ~VlnPlot(.x, features = c("PF3D7-0417500", "PF3D7-1464400"), group.by = "fem_subclusters", slot = "counts"))

map(seu_scvi_fem_sub_clusters, ~VlnPlot(.x, features = c("PF3D7-0417500", "PF3D7-1464400"), group.by = "fem_subclusters"))
map(pseudo_fem, ~VlnPlot(.x, features = c("PF3D7-0417500", "PF3D7-1464400"), group.by = "fem_subclusters"))

```


```{r, fig.height=6, fig.width=6}
# vlcano_mrg <- map(c("bulkDseq", "seuWilcox"), ~{
#   EnhancedVolcano(de_wilcox_bulk_scvi %>% filter(cluster == "8"),
#                 lab = de_wilcox_bulk_scvi %>% filter(cluster == "8") %>% pull(!!rlang::sym(paste0("Gene_", .x))),

vlcano_actvFem <- imap(fem_subclusters_bulk_pwDE_anot, ~EnhancedVolcano(.x ,
                                                                        lab = .x %>% pull(!!rlang::sym(paste0("Gene"))),
                                                                        # selectLab = .y,
                                                                        x = paste0("avg_log2FC"),
                                                                        y = paste0("p_val_adj"),
                                                                        pCutoff = 1e-10,
                                                                        FCcutoff = 1.2,
                                                                        pointSize = 1.2,
                                                                        labSize = 4.0,
                                                                        title = .y,
                                                                        subtitle = "",
                                                                        titleLabSize = 20,
                                                                        subtitleLabSize = 0,
                                                                        axisLabSize = 22,
                                                                        drawConnectors = TRUE,
                                                                        # boxedLabels = TRUE,
                                                                        # legendLabels =c( "p_val_adj"),
                                                                        gridlines.minor = FALSE
) +
  # labs(title = .y)+
  theme(legend.position = "none")
) 

vlcano_actvFem

```

### Leiden average expression

```{r, message=FALSE, warning=FALSE, eval=T, fig.width=6, fig.height=3.2}
avg_exprsn_fem_clusters_data <- map(seu_scvi_fem_sub_clusters, ~AverageExpression(.x, group.by = "fem_hires_subclusters", layer = "data") %>% .$RNA %>% as.data.frame() %>% rownames_to_column("gene"))
avg_exprsn_fem_clusters_counts <- map(seu_scvi_fem_sub_clusters, ~AverageExpression(.x, group.by = "fem_hires_subclusters", layer = "counts") %>% .$RNA %>% as.data.frame() %>% rownames_to_column("gene"))

```


```{r, message=FALSE, warning=FALSE, eval=T, fig.width=6, fig.height=3.2}
avg_exprsn_fem_clusters_data
```

```{r, message=FALSE, warning=FALSE, eval=T, fig.width=6, fig.height=3.2}
map(list(avg_exprsn_fem_clusters_data[[1]], avg_exprsn_fem_clusters_counts[[1]]), ~.x %>%
      ggplot(aes(x = `Female2`, y = `Female4`)) +
      geom_point(size = 0.01) +
      geom_abline()
)

```


```{r, message=FALSE, warning=FALSE, eval=T, fig.width=6, fig.height=3.2}
map(list(avg_exprsn_fem_clusters_data[[1]], avg_exprsn_fem_clusters_counts[[1]]), ~.x %>%
      filter(Female2 < 20 & Female4 < 20 ) %>%
      ggplot(aes(x = `Female2`, y = `Female4`)) +
      geom_point(size = 0.01) +
      geom_abline()
)
```

```{r, message=FALSE, warning=FALSE, eval=T, fig.width=6, fig.height=3.2}
map(list(avg_exprsn_fem_clusters_data[[1]], avg_exprsn_fem_clusters_counts[[1]]), ~.x %>%
      filter(Female2 < 20 & `Female (LE)` < 20 ) %>%
      ggplot(aes(x = `Female2`, y = `Female (LE)`)) +
      geom_point(size = 0.01) +
      geom_abline()
)
```

#### GGgally for pairwise comparisons across all clusters

```{r}
lowerfun <- function(data, mapping) {
  ggplot(data = data, mapping = mapping) +
    geom_point(size = 0.3, alpha = 0.3) +
    geom_abline(slope = 1, intercept = 0, color = "blue")
}

```

```{r}
## pairwise correlation between female leiden subclusters
map(list(avg_exprsn_fem_clusters_data[[1]], avg_exprsn_fem_clusters_counts[[1]]), ~.x %>%
      column_to_rownames("gene") %>%
      ggpairs(lower = list(continuous = GGally::wrap(lowerfun)),
              upper = list(continuous = wrap("cor", size = 5)),
              progress = FALSE)
)
```

```{r}
## pairwise correlation of each gene's expression averaged across all cells between female leiden subclusters - less than 50 transcripts
map(list(avg_exprsn_fem_clusters_data[[1]], avg_exprsn_fem_clusters_counts[[1]]), ~.x %>%
      column_to_rownames("gene") %>%
      filter(if_all(everything(), ~ .x <= 50)) %>%
      ggpairs(lower = list(continuous = GGally::wrap(lowerfun)),
              upper = list(continuous = wrap("cor", size = 5)),
              progress = FALSE)
)
```

```{r}
## pairwise correlation of each gene's expression averaged across all cells between female leiden subclusters - log transformed
map(list(avg_exprsn_fem_clusters_data[[1]], avg_exprsn_fem_clusters_counts[[1]]), ~.x %>%
      column_to_rownames("gene") %>%
      ggpairs(lower = list(continuous = GGally::wrap(lowerfun)),
              upper = list(continuous = wrap("cor", size = 5)),
              progress = FALSE) +
      
      scale_x_log10() +
      scale_y_log10() 
    
)
```

### Pseudobulk DE between each female subcluster and all other female subclusters

```{r, warning=FALSE, message=FALSE}
# pseudo_fem <- map(pseudo_fem, ~SetIdent(.x, value = "leiden"))

# fem_hires_subclusters_bulk_de_allmarkers <- map(pseudo_fem, ~FindAllMarkers(object = .x, test.use = "DESeq2"))
```

```{r, warning=FALSE, message=FALSE}
# fem_hires_subclusters_bulk_de_allmarkers_anot <- map(fem_hires_subclusters_bulk_de_allmarkers, possibly(~.x %>% 
#                                         # filter(p_val_adj < 0.05) %>% 
#                                         group_by(cluster) %>% arrange(p_val_adj, .by_group = T) %>% 
#                                         left_join(., Pf3D7_genes_plasmoDB[,c("gene_id", "Gene")], by = c("gene" = "gene_id"))))
# 
# fem_hires_subclusters_bulk_de_allmarkers_anot
```

Findmarkers edgeR but similar to above but dreamlet

```{r}
# seu_scvi_fem_sub_clusters_sce <- map(seu_scvi_fem_sub_clusters, ~as.SingleCellExperiment(.))
go()
```


```{r}
library(edgeR)
y <- Seurat2PB(seu_scvi_fem_sub_clusters[[1]], sample="orig.ident", cluster="leiden")
```

```{r}
summary(y$samples$lib.size)

keep.samples <- y$samples$lib.size > 5e4

table(keep.samples)

y <- y[, keep.samples]
```

```{r}
keep.genes <- filterByExpr(y, group=y$samples$cluster)

table(keep.genes)

y <- y[keep.genes, , keep=FALSE]
```

```{r}
y <- normLibSizes(y)
head(y$samples, n=10L)

summary(y$samples$norm.factors)
```

```{r}
cluster <- as.factor(y$samples$cluster)
limma::plotMDS(y, pch=16, col=c(2:8)[cluster], main="MDS")
legend("topleft", legend=paste0("cluster",levels(cluster)), pch=16, col=2:8, cex=0.8)
```

```{r}
donor <- factor(y$samples$sample)
design <- model.matrix(~ cluster + donor)
colnames(design) <- gsub("donor", "", colnames(design))
colnames(design)[1] <- "Int"
head(design)
```

```{r}
dim(design)
```

```{r}
y <- estimateDisp(y, design, robust=TRUE)

y$common.dispersion

plotBCV(y)
```

```{r}
fit <- glmQLFit(y, design, robust=TRUE)
plotQLDisp(fit)
```

```{r}
ncls <- nlevels(cluster)
contr <- rbind( matrix(1/(1-ncls), ncls, ncls), matrix(0, ncol(design)-ncls, ncls) )
diag(contr) <- 1
contr[1,] <- 0
rownames(contr) <- colnames(design)
colnames(contr) <- paste0("cluster", levels(cluster))
contr
```

```{r}
qlf <- list()
for(i in 1:ncls){
  qlf[[i]] <- glmQLFTest(fit, contrast=contr[,i])
  qlf[[i]]$comparison <- paste0("cluster", levels(cluster)[i], "_vs_others")
}
```

```{r}
qlf
```

```{r}
topTags(qlf[[1]], n=10L)

```

```{r}
qlf[[1]]$table %>% filter(PValue < 0.05) %>% filter(logFC > 0)
```

```{r}
dt <- lapply(lapply(qlf, decideTests), summary)
dt.all <- do.call("cbind", dt)
dt.all
```

```{r}
dt <- lapply(lapply(qlf, decideTests), summary)

```

```{r}
top <- 100
topMarkers <- list()
for(i in 1:ncls) {
  ord <- order(qlf[[i]]$table$PValue, decreasing=FALSE)
  up <- qlf[[i]]$table$logFC[ord] > 0
  topMarkers[[i]] <- rownames(y)[ord[up][1:top]]
}
topMarkers <- unique(unlist(topMarkers))
topMarkers
```

```{r}
lcpm <- edgeR::cpm(y, log=TRUE)
annot <- data.frame(cluster=paste0("cluster ", cluster))
rownames(annot) <- colnames(y)
ann_colors <- list(cluster=2:5)
names(ann_colors$cluster) <- paste0("cluster ", levels(cluster))
pheatmap::pheatmap(lcpm[topMarkers, ], breaks=seq(-2,2,length.out=101),
                   color=colorRampPalette(c("blue","white","red"))(100), scale="row",
                   cluster_cols=TRUE, border_color="NA", fontsize_row=5,
                   treeheight_row=70, treeheight_col=70, cutree_cols=7,
                   clustering_method="ward.D2", show_colnames=FALSE,
                   annotation_col=annot, annotation_colors=ann_colors)
```

## Compare DESeq2 vs edgeR Results

```{r}
# Extract edgeR results into data frame
edgeR_results <- map_df(qlf, ~.x$table %>% 
                          rownames_to_column("gene") %>%
                          mutate(comparison = .x$comparison) %>%
                          select(gene, comparison, logFC, PValue), 
                        .id = "contrast_idx") %>%
  select(-contrast_idx)

# Convert p-values to adjusted (approx using Benjamini-Hochberg)
edgeR_results <- edgeR_results %>%
  group_by(comparison) %>%
  mutate(PValue_adj = p.adjust(PValue, method = "BH")) %>%
  ungroup()

head(edgeR_results)
```


```{r}
# Extract DESeq2 results (cluster comparisons only)
deseq2_results <- fem_subclusters_bulk_de_allmarkers_anot %>%
  bind_rows(.id = "contrast_idx") %>%
  mutate(comparison = paste0("cluster",cluster,"_vs_others"), 
         test = "DESeq2") %>%
  select(gene, avg_log2FC, p_val_adj, comparison, test) 

head(deseq2_results)
```

```{r}
# Merge results by gene
comparison_df <- deseq2_results %>%
  rename(logFC_DESeq2 = avg_log2FC, padj_DESeq2 = p_val_adj) %>%
  inner_join(edgeR_results %>% 
               select(gene, logFC, PValue_adj, comparison) %>%
               rename(logFC_edgeR = logFC, padj_edgeR = PValue_adj),
             by = c("gene", "comparison")) %>%
  filter(!is.na(logFC_edgeR) & !is.na(logFC_DESeq2))

cat("Genes compared:", nrow(comparison_df), "\n")
cat("Correlation of logFC:", cor(comparison_df$logFC_DESeq2, comparison_df$logFC_edgeR, use="complete"), "\n")
cat("Correlation of -log10(p_adj):", 
    cor(-log10(comparison_df$padj_DESeq2 + 1e-300), 
        -log10(comparison_df$padj_edgeR + 1e-300), use="complete"), "\n")
```

```{r, fig.height=5, fig.width=10}
# Plot comparisons
p1 <- comparison_df %>%
  ggplot(aes(x = logFC_DESeq2, y = logFC_edgeR)) +
  geom_point(alpha = 0.3, size = 1) +
  geom_abline(intercept = 0, slope = 1, color = "red", linetype = "dashed") +
  labs(title = "LogFC: DESeq2 vs edgeR", x = "DESeq2 logFC", y = "edgeR logFC") +
  theme_bw()

p2 <- comparison_df %>%
  ggplot(aes(x = -log10(padj_DESeq2 + 1e-300), y = -log10(padj_edgeR + 1e-300))) +
  geom_point(alpha = 0.3, size = 1) +
  geom_abline(intercept = 0, slope = 1, color = "red", linetype = "dashed") +
  labs(title = "-log10(padj): DESeq2 vs edgeR", x = "DESeq2 -log10(padj)", y = "edgeR -log10(padj)") +
  theme_bw()

p1 | p2
```

