---
title: "Integration of donor only and with lab"
date: "2024-01-05"
author: "Jesse Rop"
output:
  html_document:
    toc: true
    theme: united
    code_folding: hide
    df_print: paged
---

```{r}
suppressPackageStartupMessages({
  library(Seurat)
  # library(sctransform)
  # library(rgl)
  library(ggrepel)
  library(ggpubr)
  library(dplyr) 
  library(stringr)
  library(purrr)
  library(readr)
  library(tibble)
  library(ggplot2)
  # library(ggnewscale)
  library(tidyr)
  library(cowplot)
  # library(scDblFinder)
  library(SeuratData)
  library(SeuratDisk)
  # library(scmap)
  library(SingleCellExperiment)
  library(pheatmap)
  library(plotly)
  # library(Hmisc)
  # library(rmarkdown)
  library(RColorBrewer)
  # library(MAST)
  library(ggvenn)
  # library("biomaRt")
  library(conflicted)
  # library(SoupX)
  library(MetBrewer)
  library(dreamlet)
  library(muscat)
  library(ExperimentHub)
  library(zenith)
  library(scater)
  library(slingshot)
  library(harmony)
  library(ggforce)
  library(miloR)
  library(scater)
  library(scran)
  library(patchwork)
  # library(scRNAseq)
  library(scuttle)
  library(irlba)
  library(BiocParallel)
  library(tradeSeq)
  library(BiocParallel)
  library(ggh4x)
  library(condiments)
  library(DESeq2)
  
  conflict_prefer("select", "dplyr")
  conflict_prefer("filter", "dplyr")
  conflict_prefer("rename", "dplyr")
  conflicts_prefer(dplyr::count)
  conflicts_prefer(base::as.data.frame)
  conflicts_prefer(base::as.factor)
  conflicts_prefer(dplyr::desc)
  
})
```


#  integration of all MSC

### Common functions
## setup 

```{r setup}
sample_set = "Mali2"

# knitr::opts_knit$set(root.dir = paste0("/Users/jr35/Google Drive/My Drive/PHD_documents/malaria_phd/", sample_set))
knitr::opts_knit$set(root.dir = paste0("/lustre/scratch126/tol/teams/lawniczak/users/jr35/phd/", sample_set))
```


#### Source commong functions, varriables and tables
```{r}
##Boxplot labeling function
source("../multipurpose_scripts_lnk/seurat_sce/seu_sce_common_fns.R")
source("../multipurpose_scripts_lnk/plotting_common_fns.R")
source("mali22_code_lnk/pf_common_vars.R")
```

```{r}
save_dir <- "data/processed/Pf/m2_all/asymptomatic_vs_symptomatic"

```

```{r setup}

# optimum_k_qcd <- c("MSC33" =8, "MSC48" =4, "MSC49" =6, "MSC55" =7, "MSC60" =7, "MSC66" =8, "MSC38" =4, "MSC40" =8, "MSC50_MACS" =1, "MSC53" =1, "MSC57" =1, "MSC24" =1, "MSC37" =5, "MSC39" =4, "MSC45" =1, "MSC50_SLO" =1, "MSC54" =1, "MSC67" =1, "MSC68" =3, "MSC70" =1)

optimum_k_qcd <- c("MSC33" =8, "MSC48" =4, "MSC49" =5, "MSC55" =7, "MSC60" =7, "MSC66" =8, "MSC38" =4, "MSC40" =8, "MSC50_MACS" =1, "MSC53" =1, "MSC57" =1, "MSC37" =5, "MSC39" =4, "MSC45" =1, "MSC50_SLO" =1, "MSC54" =1, "MSC67" =1, "MSC68" =3, "MSC70" =1)

optimum_k_qcd <- optimum_k_qcd[sort(names(optimum_k_qcd))]

```


read in files 
clean objects all donors
```{r}
# msc_w_cln_strns <- readRDS("data/processed/Pf/m2_all/msc_w_cln_strns.RDS")

```

```{r, eval=T}
# imap(msc_w_cln_strns, ~DimPlot(.x, group.by = "StrainDon")  + labs(title = .y))
```

```{r, eval=T}
# imap(msc_w_cln_strns, ~DimPlot(.x, group.by = "stage")  + labs(title = .y) + scale_color_manual(values = sp_fld_colors))
```

!!!NOTE - Reading in the metadata to add to the mixed species samples which were missed in the initial step - Will need to redo adding to the source
```{r}
mrc_sk21_summary <- read_csv("data/raw/MRC_Sample_Analysis_Summary_Summary.csv", skip = 1) 

# mrc_sk21_summary_mxd <- mrc_sk21_summary %>% 
#   filter(Sample_Name %in% c("MSC25" ,"MSC41", "MSC51")) %>%
#   mutate(across(Symptomatic, ~str_replace_all(., c("Y" = "Yes", "N" = "No")))) %>%
#   select(Sample_Name,Symptomatic, Hb_Type) %>% 
#   rename("donor" = Sample_Name)

mrc_sk21_summary_mxd <- mrc_sk21_summary %>% 
  select(Sample_Name, Hb_Type) %>% 
  rename("donor" = Sample_Name)

```



Remove mixed infections for now
```{r}
# msc_w_cln_strns <- msc_w_cln_strns %>% .[!(names(msc_w_cln_strns) %in% c("MSC25" ,"MSC41", "MSC51"))]
# msc_w_cln_strns <- msc_w_cln_strns %>% .[!(names(msc_w_cln_strns) %in% c("MSC25" , "MSC51"))] ## Retain MSC41 since it is mostly Pf with few Po
# msc_w_cln_strns <- msc_w_cln_strns %>% .[!(names(msc_w_cln_strns) %in% c("MSC25" , "MSC41"))] ## Retain MSC51 since MSC41 is pseudotime outlier with mostly trophs
# msc_w_cln_strns <- msc_w_cln_strns  ## Retain all to assess how they look
```


Strain info
```{r}
# strn_c_new_mdata_m22_min_qcd <- readRDS("data/processed/Pf/strn_c_new_mdata_m22_min_qcd.RDS")
```


Integrated asexual
```{r}
## subset where field parasites are for slingshot
# v3_m2_asex_harmony_ss_res_fld_pt <- readRDS("data/processed/Pf/m2_all/v3_m2_asex_harmony_ss_res_fld_pt.RDS")
# v3_m2_asex_harmony_ss_res_fld_pt <- readRDS("data/processed/Pf/m2_all/v3_m2_asex_harmony_res_fld_pt.RDS")
# v3_m2_asex_cr_filt_norm_harmony_pt <- readRDS("data/processed/Pf/v3_m2_asex_cr_filt_norm_harmony_pt.RDS")
v3_m2_asex_cr_filt_norm_harmony_pt <- readRDS("data/processed/Pf/asex_noM55_jcC_harmony_no_comtd_pt3.RDS")

v3_m2_asex_cr_filt_norm_harmony_pt <- v3_m2_asex_cr_filt_norm_harmony_pt[, !v3_m2_asex_cr_filt_norm_harmony_pt$Strain_DN %in% c("Doublet", "Negative")]

```


Integrated asexual
```{r}
v3_m2_asex_harmony_ss_res_fld_pt <- v3_m2_asex_cr_filt_norm_harmony_pt

```

```{r}
v3_m2_asex_harmony_ss_res_fld_pt@meta.data %>% glimpse

```

```{r}
## add haemoglobin type
# v3_m2_asex_harmony_ss_res_fld_pt_mxdSp <- v3_m2_asex_harmony_ss_res_fld_pt@meta.data %>% 
#   rownames_to_column("bcode") %>% 
#   left_join(., mrc_sk21_summary_mxd, by = "donor") %>%
#   # mutate(Symptomatic = coalesce(Symptomatic.x, Symptomatic.y), Hb_Type = coalesce(Hb_Type.x, Hb_Type.y)) %>% 
#   # select(bcode, Symptomatic, Hb_Type) %>% 
#   select(bcode,  Hb_Type) %>% 
#   column_to_rownames("bcode") %>% 
#   AddMetaData(v3_m2_asex_harmony_ss_res_fld_pt, .)
v3_m2_asex_harmony_ss_res_fld_pt_mxdSp<-v3_m2_asex_harmony_ss_res_fld_pt
```


Remove 1 early ring
```{r}
# v3_m2_asex_harmony_ss_res_fld_pt_mxdSp <- v3_m2_asex_harmony_ss_res_fld_pt_mxdSp[,!(v3_m2_asex_harmony_ss_res_fld_pt_mxdSp$StageFL == "Early ring")]
```

Remove msc12
```{r}
v3_m2_asex_harmony_ss_res_fld_pt_mxdSp <- v3_m2_asex_harmony_ss_res_fld_pt_mxdSp[,!(v3_m2_asex_harmony_ss_res_fld_pt_mxdSp$donor == "MSC12")]
```


```{r, fig.width=15, fig.height=5}
donor_mdata_slct <- readRDS("data/processed/Pf/m2_all/donor_mdata_slct.RDS")

```

add metadata

```{r, fig.width=15, fig.height=5}
# v3_m2_asex_harmony_ss_res_fld_pt_mxdSp <- v3_m2_asex_harmony_ss_res_fld_pt_mxdSp@meta.data  %>%
#   rownames_to_column("bcode") %>% 
#   mutate(donor_nm = str_remove_all(donor, "_SLO|_MACS")) %>%
#   select(bcode, donor_nm) %>%
#   left_join(., donor_mdata_slct, by=c("donor_nm"="short_name"), relationship = "many-to-one") %>%
#   select(bcode, any_of(colnames(donor_mdata_slct))) %>% 
#   column_to_rownames("bcode") %>% 
#   AddMetaData(v3_m2_asex_harmony_ss_res_fld_pt_mxdSp, .)

```


```{r}
v3_m2_asex_harmony_ss_res_fld_pt_mxdSp <- v3_m2_asex_harmony_ss_res_fld_pt_mxdSp@meta.data %>% 
  # mutate(across(Symptomatic, ~case_when(donor %in% c("MSC14", "MSC13", "MSC3", "MSC1") ~ "No", T ~ .)),
  mutate(Symptomatic = condition,
         across(Hb_Type, ~case_when(donor == "MSC14" ~ "HbC", T ~ .))) %>% 
  mutate(across(Symptomatic, ~factor(str_replace_all(.,  c("Yes" = "Symptomatic", "No" = "Asymptomatic")), levels=c("Symptomatic", "Asymptomatic")))) %>%
  select(Symptomatic, Hb_Type) %>% 
  AddMetaData(v3_m2_asex_harmony_ss_res_fld_pt_mxdSp, .)
```

Check which stage label to use
```{r}
# v3_m2_asex_harmony_ss_res_fld_pt_mxdSp@meta.data %>% count(stage, StageFL, StageFL, StageFL_ND)

v3_m2_asex_harmony_ss_res_fld_pt_mxdSp@meta.data %>% count(stage)
```


```{r}
## Creating new stage variables which are need to fit in with old code
v3_m2_asex_harmony_ss_res_fld_pt_mxdSp <- v3_m2_asex_harmony_ss_res_fld_pt_mxdSp@meta.data %>% 
  mutate(StageFL = stage,
         StageFL_ND = stage,
         dset = dset_labmrg,
         StrainDon = paste0(str_remove(donor, "SC"), "_", Strain_DN),
         pseudotime = slingPseudotime_1,
         pseudotime_fld_only = slingPseudotime_1,
         stage = stage,
         strain = Strain_DN) %>% 
  select(StageFL, StageFL_ND, dset,StrainDon,pseudotime, pseudotime_fld_only,  strain) %>% 
  AddMetaData(v3_m2_asex_harmony_ss_res_fld_pt_mxdSp, .)
```


Visualize all cells including mixed infections
```{r, fig.width=3, fig.height=3.5}
DimPlot(v3_m2_asex_harmony_ss_res_fld_pt_mxdSp, split.by = "dset", reduction = "umap.harmony_PC_man", group.by = "StageFL_ND", pt.size = 0.1) + scale_color_manual(values = sp_fld_colors)+
  theme(legend.position = "none", 
        axis.text = element_blank(), 
        plot.title  = element_blank())+
  # guides(x = "axis_nested") +
  labs(x = "UMAP_1", y="UMAP_2") 
```

Remove outlier MSC41 and lab for now
```{r, fig.width=7.5, fig.height=2}
# v3_m2_asex_harmony_ss_res_fld_pt_mxdSp_noOutlier <- v3_m2_asex_harmony_ss_res_fld_pt_mxdSp[,v3_m2_asex_harmony_ss_res_fld_pt_mxdSp$donor != "MSC41" | v3_m2_asex_harmony_ss_res_fld_pt_mxdSp$dset == "Lab"]

## Retain all donor cells
v3_m2_asex_harmony_ss_res_fld_pt_mxdSp_noOutlier <- v3_m2_asex_harmony_ss_res_fld_pt_mxdSp[,v3_m2_asex_harmony_ss_res_fld_pt_mxdSp$donor != "MSC12" | v3_m2_asex_harmony_ss_res_fld_pt_mxdSp$dset == "Lab"]

```


Visualize all cells including mixed infections
```{r, fig.width=3, fig.height=3.5}
DimPlot(v3_m2_asex_harmony_ss_res_fld_pt_mxdSp_noOutlier, split.by = "dset", reduction = "umap.harmony_PC_man", group.by = "StageFL_ND", pt.size = 0.1) + scale_color_manual(values = sp_fld_colors)+
  theme(legend.position = "none", 
        axis.text = element_blank(), 
        plot.title  = element_blank())+
  # guides(x = "axis_nested") +
  labs(x = "UMAP_1", y="UMAP_2") 
```


```{r, fig.width=7.5, fig.height=2}
tbl_all_umap <- v3_m2_asex_harmony_ss_res_fld_pt_mxdSp_noOutlier[,v3_m2_asex_harmony_ss_res_fld_pt_mxdSp_noOutlier$dset != "Lab"]@reductions$umap.harmony_PC_man@cell.embeddings %>% as.data.frame() %>% rownames_to_column("bcode") %>% left_join(., v3_m2_asex_harmony_ss_res_fld_pt_mxdSp_noOutlier@meta.data %>% rownames_to_column("bcode")) 

```

```{r setup, include=FALSE}
knitr::opts_chunk$set(dpi=300)
```

```{r, fig.width=10, fig.height=1.8, dpi=500}

tbl_all_umap %>%
  mutate(across(donor, ~str_remove_all(., "SC|_SLO|C"))) %>%
  # ggplot(., aes(x = weave_factors(Strain, donor), y = Counts, fill = donor)) +
  ggplot(., aes(x = umapharmonyPCman_1, y = umapharmonyPCman_2, colour = stage)) +
  geom_point(size = 0.1) +
  scale_color_manual(values = sp_fld_colors) +
  ggh4x::facet_nested_wrap(vars(condition, donor),  nrow = 1, scales = "free_x") +
  theme_classic()+
  theme(legend.position = "none", 
        axis.text = element_blank(), 
        axis.ticks = element_blank(),
        axis.title = element_blank(),
        strip.text =element_text(size=16,face="bold", colour = 'black'),
        # panel.border=element_rect(colour="black",fill='transparent')
        )+
  # guides(x = "axis_nested") +
  labs(x = "UMAP_1", y="UMAP_2") 
```


```{r, fig.width=16, fig.height=4}
## !! NOTE -remove strains with less than 20 - should be done at the beginning
sympt_fev_tbl1 <- v3_m2_asex_harmony_ss_res_fld_pt_mxdSp_noOutlier[,v3_m2_asex_harmony_ss_res_fld_pt_mxdSp_noOutlier@meta.data %>% rownames_to_column('bcode') %>% add_count(StrainDon, name="strn_n") %>% filter(strn_n>=20) %>% pull(bcode)] %>%
  .@meta.data %>%
  filter(dset != "Lab") %>%
  mutate(across(c(donor, StrainDon), ~str_remove_all(., "_SLO"))) %>% 
  # mutate(fev_sympt = case_when(fever == "Fever" & Symptomatic == "Symptomatic" ~ "Sympto_Fever", fever == "NoFever" & Symptomatic == "Asymptomatic" ~ "NonSympto_NoFever")) %>% 
  mutate(fev_sympt = case_when(fever == "Fever" & condition == "Symptomatic" ~ "Sympto_Fever", fever == "NoFever" & condition == "Asymptomatic" ~ "NonSympto_NoFever")) %>%
  add_count(StrainDon, name = "strain_n") #%>%
  # filter(strain_n > 20)

```


same as above without fever
```{r, fig.width=14, fig.height=4}
tbl <- sympt_fev_tbl1 %>%
  count(condition, donor, StrainDon) %>% 
  mutate(across(donor, ~str_remove_all(., "SC|C"))) %>%
  group_by(condition, donor) %>%
  add_count(name = "strn_n") %>% 
  ungroup() %>%
  mutate(prop_wdth =strn_n/n()) 

## Get width proportions of the different facets based on number of strains per group
p_width <- tbl%>% distinct(condition, donor, prop_wdth) %>% pull(prop_wdth) %>% round(., digits = 2) 

# # don_labls <- as_labeller(tbl%>%pull(donor) %>% unique() %>% str_replace(., "M", "M\n"))
# don_labls <- tbl%>%pull(donor) %>% unique() 
# don_key <- paste0(don_labls, "=", don_labls %>% str_replace(., "M", "M\n"))


tbl %>%
  mutate(donor_lbl = str_replace(donor, "M", "M\n")) %>%
  rename(c('Counts' =n, 'Strain' = StrainDon))  %>%
  # ggplot(., aes(x = weave_factors(Strain, donor), y = Counts, fill = donor)) +
  # ggplot(., aes(x = Strain, y = Counts/1000, fill = donor)) +
  ggplot(., aes(x = Strain, y = Counts/1000, fill = Strain)) +
  geom_col(position = position_dodge2(width = 0.9, preserve = "single")) +
  # scale_fill_manual(values = donor_cols %>% set_names(str_remove(names(donor_cols), "_SLO"))) +
  scale_fill_manual(values = strain_don_cols ) +
  geom_text(aes(label=Counts), angle = 90, hjust=0, vjust= 0.5, position = position_dodge2(width = 0.9, preserve = "single"), size = 5, show.legend = F, fontface ='bold') +
  scale_y_continuous(expand = expansion(mult = c(0.01, 0.78))) +
  ggh4x::facet_nested_wrap(vars(condition, donor_lbl),  nrow = 1, scales = "free_x") +
  theme_classic()+
  theme(legend.position = "none", 
        axis.title = element_text(size = 16),
        axis.text.y = element_text(size = 14),
        axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 15),
        strip.text =element_text(size=16,face="bold", colour = 'black'
                                 ),
        panel.border=element_rect(colour="black",fill='transparent')
        )+
  # guides(x = "axis_nested") +
  labs(x = "Strain", y = "Counts (X1000)") +
  force_panelsizes(cols = p_width, respect = TRUE, total_width = unit(12.6, "inches"), total_height = unit(1.3, "inches"))
```

Remove mixed infections for now
```{r}
v3_m2_asex_harmony_ss_res_fld_pt_pf <- v3_m2_asex_harmony_ss_res_fld_pt_mxdSp[,!(v3_m2_asex_harmony_ss_res_fld_pt_mxdSp$donor %in% c("MSC25" ,"MSC41", "MSC51", "MSC12"))]

# v3_m2_asex_harmony_ss_res_fld_pt_pf <- v3_m2_asex_harmony_ss_res_fld_pt_mxdSp[,!(v3_m2_asex_harmony_ss_res_fld_pt_mxdSp$donor %in% c( "MSC12"))] ## Retain all but since MSC51 for HbS/C vs HbN analysis is pseudotime outlier with mostly trophs

# v3_m2_asex_harmony_ss_res_fld_pt_pf <- v3_m2_asex_harmony_ss_res_fld_pt_mxdSp[,!(v3_m2_asex_harmony_ss_res_fld_pt_mxdSp$donor %in% c("MSC25", "MSC51"))] ## Retain MSC41 since it is mostly Pf with few Po
# v3_m2_asex_harmony_ss_res_fld_pt_pf <- v3_m2_asex_harmony_ss_res_fld_pt_mxdSp[,!(v3_m2_asex_harmony_ss_res_fld_pt_mxdSp$donor %in% c("MSC25", "MSC41"))] ## Retain MSC51 since MSC41 is pseudotime outlier with mostly trophs
# v3_m2_asex_harmony_ss_res_fld_pt_pf <- v3_m2_asex_harmony_ss_res_fld_pt_mxdSp ## Retain all to assess how they look

```



Visualize all cells
```{r, fig.width=3, fig.height=3.5}
DimPlot(v3_m2_asex_harmony_ss_res_fld_pt_pf, split.by = "dset", reduction = "umap.harmony_PC_man", group.by = "StageFL_ND", pt.size = 0.3) + scale_color_manual(values = sp_fld_colors)+
  theme(legend.position = "none", 
        axis.text = element_blank(), 
        plot.title  = element_blank(),
        strip.text =element_text(size=24,face="bold"),
        axis.title =element_text(size=20,face="bold"))+
  # guides(x = "axis_nested") +
  labs(x = "UMAP_1", y="UMAP_2") 
```

```{r, fig.width=1.5, fig.height=3}
DimPlot(v3_m2_asex_harmony_ss_res_fld_pt_pf, reduction = "umap.harmony_PC_man", group.by = "StageFL_ND", pt.size = 0.1) + scale_color_manual(values = sp_fld_colors)+
  theme(legend.position = "none", 
        axis.text = element_blank(), 
        plot.title  = element_blank())+
  # guides(x = "axis_nested") +
  labs(x = "UMAP_1", y="UMAP_2") 
```

Subset only the field parasites

```{r, fig.width=15, fig.height=5}
# visualize by batch and cell type annotation
m2_asex_harmony_tr <- v3_m2_asex_harmony_ss_res_fld_pt_pf[,v3_m2_asex_harmony_ss_res_fld_pt_pf@meta.data$dset == "Field"]
```

!!NOTE - May need to be reviewed in future incase MSC50_MACS is in the same object : Rename MSC0_SLO to MSC50 since MSC50_MACS doesn't have asexual and is not included in the object

```{r}
m2_asex_harmony_tr <- m2_asex_harmony_tr@meta.data %>% mutate(across(c(donor,StrainDon), ~str_remove(., "_SLO"))) %>% select(donor,StrainDon) %>% AddMetaData(m2_asex_harmony_tr, .)
```

!!NOTE - May need to be reviewed in future - filter out strains that are less than or equal to 20
```{r}
m2_asex_harmony_tr <- m2_asex_harmony_tr[,m2_asex_harmony_tr@meta.data %>% rownames_to_column('bcode') %>% add_count(StrainDon, name="strn_n") %>% filter(strn_n>=20) %>% pull(bcode)]
```

```{r}
m2_asex_harmony_tr_list <- list(m2_asex_harmony_tr)
```

#############START - BALANCE CELLS PSEUDOTIME CHUNKING START Calculate pseudotime chunks and downsample within chunks #############

```{r, warning=FALSE, message=FALSE}
## Calculate pseudotime chunks
## Name these  pseudotime chunks as downsampling_xxx since they are for downsampling

bal_df_asex_symptoDE <- map(m2_asex_harmony_tr_list, ~ptime_chunk_calc_df_fn(sc_df = .x@meta.data, ptime = "slingPseudotime_1", chunk_num = 20, p_cat_nm = "downsampling_ptime_cat", p_bal_cat_nm = "downsampling_ptime_bal_cat"))
```

```{r, warning=FALSE, message=FALSE}
## Calculate pseudotime chunks
m2_asex_harmony_tr_list <- map2(m2_asex_harmony_tr_list, bal_df_asex_symptoDE, ~AddMetaData(.x, .y) )
```

```{r}
## subset where field parasites are for slingshot
map(m2_asex_harmony_tr_list, ~DimPlot(.x, reduction = "umap.harmony_PC_man", group.by = "downsampling_ptime_cat" ))
map(m2_asex_harmony_tr_list, ~DimPlot(.x, reduction = "umap.harmony_PC_man", group.by = "downsampling_ptime_bal_cat" ))
```


```{r}
# Colours for plotting pseudotime chunks
ptc_cols <- c("#068B4CFF", "#4FB0C0FF", "#74DC09FF", "#3377ff", "#E769ACFF", 
"#8A1708FF", "#048461FF", "#3BDF3FFF", "#5559B5FF", "#557406FF", 
"#2A158CFF", "#B48606FF", "#CA2F4DFF", "#E75FFDFF", "#E35655FF", 
"#42CFC3FF", "#1BA0B9FF", "#D481E8FF", "#958FEBFF", "#4A7406FF", 
"#540B9CFF", "#341B4DFA") %>% set_names(paste0("PT.", 1:22))
```

```{r}
map(m2_asex_harmony_tr_list, ~DimPlot(.x, reduction = "umap.harmony_PC_man", group.by = "downsampling_ptime_bal_cat", cols = ptc_cols, label = T ) +
  theme(legend.position = "none"))
```

```{r}
# imap(m2_asex_harmony_tr_list, ~saveRDS(.x, paste0(save_dir, "afm_leiden_w_donors_pt_downsampling_chunks_", .y, ".RDS")))
```


```{r}
# Count the donors with atleast 2 strains each with atleast 100 cells
# don_100n <- imap(afm_leiden_w_donors_pt, ~.x@meta.data %>% count(orig.ident, StrainDon) %>% filter(n>100) %>% group_by(orig.ident) %>% filter(n_distinct(StrainDon) > 1) %>% pull(orig.ident) %>% unique() %>% str_remove(., ".mrna"))
# don_100n <- imap(m2_asex_harmony_tr_list, ~.x@meta.data %>% count(orig.ident, StrainDon) %>% pull(orig.ident) %>% unique() %>% str_remove(., ".mrna"))
```


```{r}
## Get only the asexuals
# m2_asex_seu <- map(msc_w_cln_strns[c("MSC49", "MSC60")], ~.x[,.x@meta.data$stage %in% c("Early trophozoite", "Late ring")] %>% AddMetaData(., ))
# dons_bal_sset <- c("M49_", "M60_", "M33_", "M66_", "M55_", "M49_|M55_")
# dons_bal_sset <- c("M49_", "M60_", "M33_", "M66_")
# dons_bal_sset <- c("M49C_", "M66C_")

# dons_bal_sset_nms <- str_remove_all(str_replace_all(dons_bal_sset, "M", "MSC"), "_$|C_$|\\|")
# dons_bal_sset_nms <- don_100n

# Retain donors with atleast 2 strains each with atleast 100 cells for each stage
# m2_afm_seu <- map2(afm_leiden_w_donors_pt, dons_bal_sset_nms, ~.x[, str_detect(.x@meta.data$orig.ident, paste(.y, collapse = "|")) & !str_detect(.x@meta.data$StrainDon, "Negative|Doublet")])
# m2_afm_seu <- map2(m2_asex_harmony_tr_list, dons_bal_sset_nms, ~.x[, str_detect(.x@meta.data$orig.ident, paste(.y, collapse = "|"))])
```

### Balance (subsample) the strain cells within pseudotime chunks


```{r, warning=FALSE, message=FALSE}
set.seed(989778)
# ds_grp <- rep('StrainDon', length(dons_bal_sset_nms))

# balanced_cells_tbl <- map2(m2_afm_seu, ds_grp, ~balanced_cells_fn(sc_df = .x@colData, dsample_condtn = .y , ptime_strategy = c("donor", "ptime_bal_cat"), multpl = 1, don_yes_no = "no"))
balanced_cells_tbl_sympto_DE <- map(m2_asex_harmony_tr_list, ~balanced_cells_fn(sc_df = .x@meta.data, dsample_condtn = c("condition") , ptime_strategy = c("downsampling_ptime_bal_cat"), multpl = 1, don_yes_no = "yes"))
```




```{r, warning=FALSE, message=FALSE}
######################### !!!!!NOTE - Unbalanced alternative to increase power replacing balanced with unbalanced - switch to return balanced or viceversa 
# m2_afm_bal_seu <- m2_afm_seu
m2_asex_bal_symptoDE <- map2(m2_asex_harmony_tr_list, balanced_cells_tbl_sympto_DE, ~.x[, .y$cell_bc])

```

```{r}
m2_asex_bal_symptoDE[[1]]@meta.data %>% data.frame() %>% dplyr::count(downsampling_ptime_bal_cat, condition, donor)
m2_asex_bal_symptoDE[[1]]@meta.data %>% data.frame() %>% dplyr::count(downsampling_ptime_bal_cat, condition)

```


```{r, warning=FALSE, message=FALSE}
## Calculate pseudotime chunks for DE
bal_df2_sympto_DE <- map(m2_asex_bal_symptoDE, ~ptime_chunk_calc_df_fn(sc_df = .x@meta.data, ptime = "slingPseudotime_1", chunk_num = 10, p_cat_nm = "ptime_cat", p_bal_cat_nm = "ptime_bal_cat"))

```


```{r, warning=FALSE, message=FALSE}
## Add pseudotime chunks to metadata
m2_asex_bal_symptoDE <- map2(m2_asex_bal_symptoDE, bal_df2_sympto_DE, ~AddMetaData(.x, .y) )

```


```{r}
m2_asex_bal_symptoDE[[1]]@meta.data %>% count(ptime_bal_cat, condition, donor)
```

```{r}
map(m2_asex_bal_symptoDE, ~DimPlot(.x, reduction = "umap.harmony_PC_man", group.by = "ptime_cat" , label = T ))
map(m2_asex_bal_symptoDE, ~DimPlot(.x, reduction = "umap.harmony_PC_man", group.by = "ptime_bal_cat" , label = T))
# go()
```

```{r}
## Subset to remove within each pseudotime chunk any strain with less than 10 cells to ensure there are no pseudotime chunks where some strains are overwhelmingly underrepresented
m2_asex_bal_symptoDE <- map(m2_asex_bal_symptoDE, ~.x[, .x@meta.data %>% 
                                               rownames_to_column('bcode') %>% 
                                               group_by(ptime_bal_cat, condition, donor) %>% 
                                               # filter(n() >=20) %>%
                                               filter(n() >=10) %>%
                                               group_by(ptime_bal_cat, condition) %>% 
                                               filter(n_distinct(donor) > 1) %>%
                                               # filter(n() >=100) %>% 
                                               pull(bcode)]
)


```

```{r}
m2_asex_bal_symptoDE[[1]]@meta.data %>% count(ptime_bal_cat, condition, donor) %>% group_by(condition, ptime_bal_cat) %>% mutate(dons = n_distinct(donor))

```

```{r, fig.width=14, fig.height=3}
map(m2_asex_bal_symptoDE, ~.x@meta.data %>% 
  count( condition, ptime_bal_cat,  donor) %>% 
  mutate(across(donor, ~str_remove_all(., "SC|SLO|C|_"))) %>% 
  group_by( condition, ptime_bal_cat) %>% 
  mutate(strn_n = n_distinct(donor)) %>% 
  filter(strn_n >1) %>%
  ggplot(aes(x = donor, y = n)) +
  geom_col() +
  # geom_col(position = position_dodge2()) + #width=0.4, position = position_dodge(width=0.5) 
  geom_text(aes(label=n), hjust=-0.05, vjust= 0.5, position = position_dodge(width = .9), size = 5, show.legend = F, fontface ='bold', angle = 90)+
  scale_y_continuous(expand = expansion(mult = c(0.01, 0.45))) +
  facet_nested_wrap(vars(condition, ptime_bal_cat), nrow = 1, scales = "free_x") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90),
        strip.text = element_text(size = 14),
        text = element_text(size = 14)) 
)
```

```{r}
map(m2_asex_bal_symptoDE, ~DimPlot(.x, reduction = "umap.harmony_PC_man", group.by = "ptime_bal_cat" , label = T ))

```
###################### END - BALANCE CELLS PSEUDOTIME CHUNKING START  Calculate pseudotime chunks and downsample within chunks  ############################################


###################### PSEUDOTIME CHUNKING START - SEU  ############################################

## pseudotime chunks used for all the DE methods below and can be used for DEseq2 above as well
### Calculate pseudotime chunks

```{r}
m2_asex_bal_symptoDE[[1]]$ptime_bal_cat10 <- m2_asex_bal_symptoDE[[1]]$ptime_bal_cat
```

```{r, warning=FALSE, message=FALSE}
## Calculate pseudotime chunks for DE
bal_df2_sympto_DE_coarse <- map(m2_asex_bal_symptoDE, ~ptime_chunk_calc_df_fn(sc_df = .x@meta.data, ptime = "slingPseudotime_1", chunk_num = 3, p_cat_nm = "ptime_cat", p_bal_cat_nm = "ptime_bal_cat"))

```


```{r, warning=FALSE, message=FALSE}
## Add pseudotime chunks to metadata
m2_asex_bal_symptoDE <- map2(m2_asex_bal_symptoDE, bal_df2_sympto_DE_coarse, ~AddMetaData(.x, .y) )

```

```{r}
m2_asex_harmony_tr_pt_chunks <- m2_asex_bal_symptoDE[[1]]
m2_asex_harmony_tr_pt_chunks$PT_grp <- m2_asex_harmony_tr_pt_chunks$ptime_bal_cat
m2_asex_harmony_tr_pt_chunks$PT_grp_unbal <- m2_asex_harmony_tr_pt_chunks$ptime_cat

```

```{r}
# ##Generate pseudotime chunks categories with range of highest and lowest pseudotime values and breaks numbers to ensure balanced representation of cells and drop the cells without strain information
# cat_n<-3
# 
# ##Remove those without pseudotime
# m2_asex_harmony_tr_pt_chunks <- m2_asex_harmony_tr[,!is.na(m2_asex_harmony_tr@meta.data[,'pseudotime'])]
# 
#       # print(sce_obj)
# m2_asex_harmony_tr_pt_chunks@meta.data[,c('PT_grp_unbal', 'PT_grp')] <- m2_asex_harmony_tr_pt_chunks@meta.data[,c( "pseudotime"), drop=F] %>%
#         data.frame() %>% 
#         mutate(PT_grp_unbal = cut(!!rlang::sym("pseudotime"),  
#                                breaks = c(-Inf, seq(min(!!rlang::sym("pseudotime")),  max(!!rlang::sym("pseudotime")), round((max(!!rlang::sym("pseudotime")) - min(!!rlang::sym("pseudotime")))/cat_n, 2)), Inf)),
#                pt_rank = dense_rank(!!rlang::sym("pseudotime")),
#                PT_grp = cut(pt_rank,  
#                                breaks = c(-Inf, seq(min(pt_rank),  max(pt_rank), round((max(pt_rank) - min(pt_rank))/cat_n, 2)), Inf))
#                ) %>%
#   mutate(ptime_grp_rnk = dense_rank(factor(PT_grp_unbal)),
#          ptime_bal_grp_rnk = dense_rank(factor(PT_grp)),
#          PT_grp_unbal = paste0("PT.", ptime_grp_rnk-1),
#          PT_grp = paste0("PT.", ptime_bal_grp_rnk-1)
#          ) %>%
#         dplyr::select(PT_grp_unbal, PT_grp)

```


```{r}
DimPlot(m2_asex_harmony_tr_pt_chunks, reduction = "umap.harmony_PC_man", group.by = "PT_grp")
```


```{r}
m2_asex_harmony_tr_pt_chunks@meta.data[,c("PT_grp", "donor")] %>% as.data.frame() %>% dplyr::count(PT_grp)
```

```{r}
DimPlot(m2_asex_harmony_tr_pt_chunks, reduction = "umap.harmony_PC_man", group.by = "PT_grp_unbal")
```

```{r}
m2_asex_harmony_tr_pt_chunks@meta.data[,c("PT_grp_unbal", "donor")] %>% as.data.frame() %>% dplyr::count(PT_grp_unbal)
```


```{r}
# m2_asex_harmony_tr_pt_sset <- m2_asex_harmony_tr_pt_chunks[,m2_asex_harmony_tr_pt_chunks$PT_grp_unbal %in% c("PT.3", "PT.4", "PT.5", "PT.6", "PT.7")]
# m2_asex_harmony_tr_pt_sset <- m2_asex_harmony_tr_pt_chunks[,m2_asex_harmony_tr_pt_chunks$PT_grp %in% c("PT.1", "PT.2", "PT.3")]
m2_asex_harmony_tr_pt_sset <- m2_asex_harmony_tr_pt_chunks[,m2_asex_harmony_tr_pt_chunks$PT_grp %in% c("PT.2", "PT.3", "PT.4")]

```

```{r}
m2_asex_harmony_tr_pt_sset$PT_grp <- str_replace_all(m2_asex_harmony_tr_pt_sset$PT_grp, c("PT.2" = "PT.1", "PT.3" = "PT.2", "PT.4" = "PT.3"))

```

```{r}
## Save DESeq2 for plotting
# saveRDS(m2_asex_harmony_tr_list, paste0(save_dir, "m2_asex_harmony_tr_list.RDS"))
saveRDS(m2_asex_harmony_tr_pt_sset, paste0(save_dir, "m2_asex_harmony_tr_pt_sset.RDS"))

```


```{r}
m2_asex_harmony_tr_pt_sset@meta.data[,c("PT_grp", "donor")] %>% as.data.frame() %>% dplyr::count(donor, PT_grp)%>% group_by(donor) %>% mutate(prop = round(n/sum(n), digits = 2))
```

Find markers for clusters
```{r}
DimPlot(m2_asex_harmony_tr_pt_sset, reduction = "umap.harmony_PC_man", label = T)

m2_asex_harmony_tr_pt_sset_hi_res <- FindClusters(m2_asex_harmony_tr_pt_sset, resolution = 1.5, verbose = F)
```

```{r}

DimPlot(m2_asex_harmony_tr_pt_sset_hi_res, reduction = "umap.harmony_PC_man", label = T, pt.size = 1)

```

############################################################ START OF ART MARKER ANALYSIS  ############################################################


```{r}
seu_markers <- FindAllMarkers(m2_asex_harmony_tr_pt_sset_hi_res, min.cells.group = 20)
```


```{r}
seu_markers %>% left_join(Pf3D7_genes_plasmoDB[, c("gene_id", "Gene")], by =c("gene" = "gene_id")) %>% group_by(cluster) %>% arrange(p_val_adj, desc(abs(avg_log2FC))) %>% slice_head(., n = 5)
```


```{r}
ARTP_markers <- readxl::read_xlsx("../published_dsets/zhu_ArtemisininResistanceTWAS_2022/resistance_genes.xlsx", sheet = "All marker transcripts", skip = 7, .name_repair = "universal")
```


```{r}
ARTP_markers_sset <- ARTP_markers %>% select(c("geneID", "desc", "symbol", "p", "fpr", "mean.difference.resistant.susceptible.", "marker.UP.Upregulation.DN.Downregulation....7", "marker.UP.Upregulation.DN.Downregulation....8")) %>% rename_with(~str_replace_all(., c(".UP.Upregulation.DN.Downregulation...." = "dir_", ".difference.resistant.susceptible."= "diff")))  %>% mutate(across(geneID, ~str_replace(., "_", "-")))

```


```{r}
ARTP_markers_sset %>% count(markerdir_7, markerdir_8 )

```

```{r}
ART_up <- ARTP_markers_sset %>% filter(markerdir_7 == "UP") %>% pull(geneID)
ART_down <- ARTP_markers_sset %>% filter(markerdir_7 == "DN") %>% pull(geneID)
ART_up_down <- ARTP_markers_sset %>% filter(markerdir_7 == "DN" | markerdir_7 == "UP") %>% pull(geneID)

```


```{r}
map(0:12, ~round(prop.table(table(seu_markers %>% filter(p_val_adj < 0.05 & abs(avg_log2FC) >1) %>% filter(cluster == .x) %>% pull(gene) %in% ART_up_down)), digits = 2))

```

```{r}
map(0:(length(unique(seu_markers$cluster))-1), ~table(seu_markers %>% filter(p_val_adj < 0.05 & abs(avg_log2FC) >1) %>% filter(cluster == .x) %>% pull(gene) %in% ARTP_markers_sset$geneID))

```

```{r}
map(0:(length(unique(seu_markers$cluster))-1), ~round(prop.table(table(seu_markers %>% filter(p_val_adj < 0.05 & abs(avg_log2FC) >1) %>% filter(cluster == .x) %>% pull(gene) %in% ARTP_markers_sset$geneID)), digits = 2)) %>% set_names(0:(length(unique(seu_markers$cluster))-1)) %>% bind_rows(., .id = "cluster") %>% filter(`TRUE` > 0.3)

```


```{r}
map(c(14,17,19), ~table(seu_markers %>% filter(p_val_adj < 0.05 & abs(avg_log2FC) >1) %>% filter(cluster == .x) %>% pull(gene) %in% ARTP_markers_sset$geneID))

```


```{r}
seu_markers %>% left_join(Pf3D7_genes_plasmoDB[, c("gene_id", "Gene")], by =c("gene" = "gene_id")) %>% group_by(cluster) %>% arrange(p_val_adj, desc(abs(avg_log2FC))) %>% filter(cluster %in% c(17)) %>% left_join(ARTP_markers_sset[, c("geneID", "meandiff", "markerdir_7", "markerdir_8")], by = c("gene"= "geneID")) %>% filter(!is.na(meandiff))
```


```{r}
map(c(14, 17, 19), ~seu_markers %>% filter(cluster %in% .x) %>% filter(p_val_adj<0.05) %>% inner_join(ARTP_markers_sset[, c("geneID", "meandiff", "markerdir_7", "markerdir_8")], by = c("gene"= "geneID")) %>% ggplot(aes(x=avg_log2FC, y=meandiff)) + geom_point()+ 
  geom_smooth(method='lm', formula= y~x))
```

```{r}

map(c(0:(length(unique(seu_markers$cluster))-1)), ~seu_markers %>% filter(cluster %in% .x) %>% filter(p_val_adj<0.05) %>% inner_join(ARTP_markers_sset[, c("geneID", "meandiff", "markerdir_7", "markerdir_8")], by = c("gene"= "geneID")) %>% ggplot(aes(x=avg_log2FC, y=meandiff)) + geom_point()+ 
  geom_smooth(method='lm', formula= y~x)+
                   stat_cor() + labs(title = paste0("cluster_",.x)))
```

```{r}

map(c(0:(length(unique(seu_markers$cluster))-1)), ~seu_markers %>% filter(cluster %in% .x) %>% filter(p_val_adj<0.05) %>% inner_join(ARTP_markers_sset[, c("geneID", "meandiff", "markerdir_7", "markerdir_8")], by = c("gene"= "geneID")) %>% dim())
```

```{r}

DimPlot(m2_asex_harmony_tr_pt_sset_hi_res, reduction = "umap.harmony_PC_man", cells.highlight = colnames(m2_asex_harmony_tr_pt_sset_hi_res[,m2_asex_harmony_tr_pt_sset_hi_res$seurat_clusters == 1]))

DimPlot(m2_asex_harmony_tr_pt_sset_hi_res, reduction = "umap.harmony_PC_man", cells.highlight = colnames(m2_asex_harmony_tr_pt_sset_hi_res[,m2_asex_harmony_tr_pt_sset_hi_res$seurat_clusters %in% c(8,15,20,23)]))
```


```{r}
Idents(m2_asex_harmony_tr_pt_sset_hi_res) <- "PT_grp"

pt_markers <- FindAllMarkers(m2_asex_harmony_tr_pt_sset_hi_res, min.cells.group = 20)
```


```{r}

map(paste0("PT.", c(1:3)), ~pt_markers %>% filter(cluster %in% .x) %>% filter(p_val_adj<0.05) %>% inner_join(ARTP_markers_sset[, c("geneID", "meandiff", "markerdir_7", "markerdir_8")], by = c("gene"= "geneID")) %>% ggplot(aes(x=avg_log2FC, y=meandiff)) + geom_point()+ 
  geom_smooth(method='lm', formula= y~x)+
                   stat_cor() + labs(title = paste0("Pseudotime_",.x)))
```


############################################################ END START OF ART MARKER ANALYSIS  ############################################################


```{r, results="asis", fig.width= 4.2, fig.height=3.2}
DimPlot(m2_asex_harmony_tr_pt_sset, reduction = "umap.harmony_PC_man", group.by = "stage", split.by = "condition", pt.size = 0.1)+
  scale_color_manual(values = sp_fld_colors)+
  theme(axis.text=element_blank(), 
        axis.title=element_blank(),
        axis.ticks = element_blank(),
        legend.position = "bottom",
        # legend.position = "none",
        legend.text = element_text(size = 15) ,
        legend.title = element_text(size = 15, face = "bold"),
        plot.title = element_blank(),
        strip.text =element_text(size=16,face="bold", colour = 'black')) +
    guides(color = guide_legend(title.position = "top", nrow =1, override.aes = list(size = 3))) +
  labs(x = "UMAP_1", y="UMAP_2") 
```

```{r, results="asis", fig.width= 2, fig.height=2.5}
DimPlot(m2_asex_harmony_tr_pt_sset, reduction = "umap.harmony_PC_man", group.by = "PT_grp", label = T)+
  theme(axis.text=element_blank(), 
        axis.title=element_blank(),
        # legend.position = "bottom",
        legend.position = "none",
        legend.text = element_text(size = 15) ,
        legend.title = element_text(size = 15, face = "bold"),
        plot.title = element_blank()) +
    guides(color = guide_legend(title.position = "top", nrow =2, override.aes = list(size = 2))) +
  labs(x = "UMAP_1", y="UMAP_2") 
```

```{r, fig.width=7.5, fig.height=2}
tbl_umap <- m2_asex_harmony_tr_pt_sset@reductions$umap.harmony_PC_man@cell.embeddings %>% as.data.frame() %>% rownames_to_column("bcode") %>% left_join(., m2_asex_harmony_tr_pt_sset@meta.data %>% rownames_to_column("bcode")) 

```


```{r, fig.width=7.5, fig.height=2}

tbl_umap %>%
  mutate(across(donor, ~str_remove_all(., "SC|C"))) %>%
  # ggplot(., aes(x = weave_factors(Strain, donor), y = Counts, fill = donor)) +
  ggplot(., aes(x = umapharmonyPCman_1, y = umapharmonyPCman_2, colour = stage)) +
  geom_point(size = 0.1) +
  scale_color_manual(values = sp_fld_colors) +
  ggh4x::facet_nested_wrap(vars(condition, donor),  nrow = 1, scales = "free_x") +
  theme_classic()+
  theme(legend.position = "none", 
        axis.text = element_blank(), 
        strip.text =element_text(size=12,face="bold", colour = 'black'),
        panel.border=element_rect(colour="black",fill='transparent')
        )+
  # guides(x = "axis_nested") +
  labs(x = "UMAP_1", y="UMAP_2") 
```

###################### PSEUDOTIME CHUNKING END  ############################################


Count asexuals per donor
```{r, fig.width=15, fig.height=5}
# visualize by batch and cell type annotation
m2_asex_harmony_tr_pt_sset@meta.data %>% count(donor, StageFL) %>% group_by(donor) %>% mutate(prop = round(n/sum(n), digits = 2))
```

Check the temp distribution between symptomatic vs asymptomatic
```{r}
m2_asex_harmony_tr_pt_sset@meta.data %>%  
  ggplot(., aes(x = condition, y = temperature_c)) +
  geom_boxplot() +
  geom_point()
```
Check the donor and strains for >37.5 fever
```{r}
m2_asex_harmony_tr_pt_sset@meta.data %>%  
  filter(temperature_c >= 37.5) %>%
  # count(donor, StrainDon, temperature_c, condition, StageFL)%>%
  count(donor, temperature_c, condition, StageFL)
```


Check the pseudotime distribution between symptomatic vs asymptomatic


```{r, fig.width=16, fig.height=4}
sympt_fev_tbl <- m2_asex_harmony_tr_pt_sset@meta.data %>% 
  mutate(fev_sympt = case_when(fever == "Fever" & condition == "Symptomatic" ~ "Sympto_Fever", fever == "NoFever" & condition == "Asymptomatic" ~ "NonSympto_NoFever")) %>%
  add_count(StrainDon, name = "strain_n") #%>%
  # filter(strain_n > 20)

```

```{r, fig.width=16, fig.height=4}
sympt_fev_tbl %>%
  ggplot(., aes(x = donor, y = pseudotime, colour = StrainDon)) +
  geom_boxplot() +
  facet_wrap(condition~., scales = "free_x")
```


```{r, fig.width=16, fig.height=4}

sympt_fev_tbl  %>%
  # filter(!is.na(fev_sympt)) %>%
  ggplot(., aes(x = donor, y = pseudotime, colour = StrainDon)) +
  geom_sina(size = 0.2) +
  # geom_boxplot() +
  facet_wrap(fever ~., scales = "free_x")
```

```{r, fig.width=16, fig.height=4}

sympt_fev_tbl  %>%
  filter(!is.na(fev_sympt)) %>%
  ggplot(., aes(x = donor, y = pseudotime, colour = StrainDon)) +
  geom_sina(size = 0.2) +
  # geom_boxplot() +
  facet_wrap(fev_sympt ~., scales = "free_x")
```

```{r, fig.width=16, fig.height=4}

sympt_fev_tbl  %>%
  filter(!is.na(fev_sympt)) %>%
  ggplot(., aes(x = donor, y = pseudotime_fld_only, colour = StrainDon)) +
  geom_sina(size = 0.2) +
  # geom_boxplot() +
  facet_wrap(fev_sympt ~., scales = "free_x")
```

```{r, fig.width=16, fig.height=4}

sympt_fev_tbl  %>%
  filter(!is.na(fev_sympt)) %>%
  ggplot(., aes(x = donor, y = smpl_wait_time, colour = StrainDon)) +
  geom_point() +
  facet_wrap(fev_sympt ~., scales = "free_x")
```


```{r, fig.width=8, fig.height=3}
sympt_fev_tbl %>%
  count(condition, donor, StrainDon) %>% 
  # mutate(across(donor, ~str_remove_all(., "SC|C")), across(StrainDon, ~str_remove(., "M.*_"))) %>% 
  mutate(across(donor, ~str_remove_all(., "SC|C"))) %>%
  rename(c('Counts' =n, 'Strain' = StrainDon))  %>%
  ggplot(., aes(x = weave_factors(Strain, donor), y = Counts, fill = donor)) +
  geom_col(position = position_dodge2(width = 0.9, preserve = "single")) +
  facet_grid(~condition, scales = "free_x", space = "free_x") + 
  theme(legend.position = "none", axis.text.x = element_text(angle = 90, hjust = 0, vjust = 0.5))+
  guides(x = "axis_nested") +
  labs(x = "Strain & Donor")
```


```{r, fig.width= 10, fig.height= 3}
# Example dataset
data <- data.frame(
  group = rep(c("A", "B"), each = 3),
  category = c("X1", "X2", "X3", "Y1", "Y2", "Y3"),
  value = c(5, 7, 6, 8, 9, 7)
)

# Basic bar plot
ggplot(data, aes(x = category, y = value, fill = group)) +
  geom_col()
```


```{r, fig.width=8, fig.height=3}
sympt_fev_tbl %>%
  count(condition, fever, donor, StrainDon) %>% 
  mutate(across(donor, ~str_remove_all(., "SC|C"))) %>%
  rename(c('Counts' =n, 'Strain' = StrainDon))  %>%
  ggplot(., aes(x = weave_factors(Strain, donor), y = Counts, fill = donor)) +
  geom_col(position = position_dodge2(width = 0.9, preserve = "single")) +
  geom_text(aes(label=Counts), angle = 90, hjust=0, vjust= 0.5, position = position_dodge2(width = 0.9, preserve = "single"), size = 5, show.legend = F, fontface ='bold')+
  ggh4x::facet_nested_wrap(vars(condition, fever),  nrow = 1, scales = "free_x") +
  theme(legend.position = "none", axis.text.x = element_text(angle = 90, hjust = 0, vjust = 0.5))+
  guides(x = "axis_nested") +
  labs(x = "Strain & Donor") +
  force_panelsizes(cols = c(1, 4, 2, 3), respect = TRUE)
```

!!PLOTTING - to be moved later
```{r, fig.width=9, fig.height=3}
tbl <- sympt_fev_tbl %>%
  count(condition, fever, donor, StrainDon) %>% 
  mutate(across(donor, ~str_remove_all(., "SC|C"))) %>%
  group_by(condition, fever, donor) %>%
  add_count(name = "strn_n") %>% 
  ungroup() %>%
  mutate(prop_wdth =strn_n/n()) 

## Get width proportions of the different facets based on number of strains per group
p_width <- tbl%>% distinct(condition, fever, donor, prop_wdth) %>% pull(prop_wdth) %>% round(., digits = 2)

tbl %>%
  rename(c('Counts' =n, 'Strain' = StrainDon))  %>%
  ggplot(., aes(x = weave_factors(Strain, donor), y = Counts, fill = donor)) +
  geom_col(position = position_dodge2(width = 0.9, preserve = "single")) +
  geom_text(aes(label=Counts), angle = 90, hjust=0, vjust= 0.5, position = position_dodge2(width = 0.9, preserve = "single"), size = 5, show.legend = F, fontface ='bold') +
  scale_y_continuous(expand = expansion(mult = c(0.01, 0.32))) +
  ggh4x::facet_nested_wrap(vars(condition, fever, donor),  nrow = 1, scales = "free_x") +
  theme_classic()+
  theme(legend.position = "none", axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))+
  # guides(x = "axis_nested") +
  labs(x = "Strain") +
  force_panelsizes(cols = p_width, respect = TRUE, total_width = unit(8, "inches"), total_height = unit(2, "inches"))
```

!!PLOTTING - to be moved later
same as above without fever
```{r, fig.width=9, fig.height=2.7}
tbl <- sympt_fev_tbl %>%
  count(condition, donor, StrainDon) %>% 
  mutate(across(donor, ~str_remove_all(., "SC|C"))) %>%
  group_by(condition, donor) %>%
  add_count(name = "strn_n") %>% 
  ungroup() %>%
  mutate(prop_wdth =strn_n/n()) 

## Get width proportions of the different facets based on number of strains per group
p_width <- tbl%>% distinct(condition, donor, prop_wdth) %>% pull(prop_wdth) %>% round(., digits = 2)

tbl %>%
  rename(c('Counts' =n, 'Strain' = StrainDon))  %>%
  # ggplot(., aes(x = weave_factors(Strain, donor), y = Counts, fill = donor)) +
  ggplot(., aes(x = Strain, y = Counts/1000, fill = donor)) +
  geom_col(position = position_dodge2(width = 0.9, preserve = "single")) +
  geom_text(aes(label=Counts), angle = 90, hjust=0, vjust= 0.5, position = position_dodge2(width = 0.9, preserve = "single"), size = 5, show.legend = F, fontface ='bold') +
  scale_y_continuous(expand = expansion(mult = c(0.01, 0.62))) +
  ggh4x::facet_nested_wrap(vars(condition, donor),  nrow = 1, scales = "free_x") +
  theme_classic()+
  theme(legend.position = "none", 
        axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1),
        strip.text =element_text(size=12,face="bold", colour = 'black'),
        panel.border=element_rect(colour="black",fill='transparent')
        )+
  # guides(x = "axis_nested") +
  labs(x = "Strain", y = "Counts (X1000)") +
  force_panelsizes(cols = p_width, respect = TRUE, total_width = unit(8.6, "inches"), total_height = unit(1.3, "inches"))
```


```{r, fig.width= 10, fig.height= 3}
sympt_fev_tbl %>% 
  count(condition, fever, donor, StrainDon) %>% 
  rename(c('Counts' =n, 'Strain' = StrainDon))  %>%
  ggplot(aes(x = donor, fill = Strain, y = Counts)) + 
  geom_col(position = position_dodge2(width = 0.9, preserve = "single")) + 
  geom_text(aes(label=Counts), angle = 90, hjust=0, vjust= 0.5, position = position_dodge2(width = 0.9, preserve = "single"), size = 5, show.legend = F, fontface ='bold') +
  scale_y_continuous(expand = expansion(mult = c(0.01, 0.24))) +
  ggh4x::facet_nested_wrap(vars(condition, fever),  nrow = 1, scales = "free_x") + 
  theme(legend.position = "none", axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)) +
  force_panelsizes(cols = c(1, 4, 2, 3), respect = TRUE)
```


```{r, fig.width = 13, fig.height = 3.5, message=FALSE, warning=FALSE, eval=FALSE}
k_ovlp_bar <- list(soupo_kselect_list_pp[c(1,4)], names(soupo_kselect_list_pp[c(1,4)]), optimum_k_pt_msc1_13_fake[c(1,4)]) %>%
  pmap(., ~{
    ..1 %>% 
  mutate("Strain_final" = !!rlang::sym(paste0("Strain_",..2,"_", ..3)),
         "Strain_2nd" = !!rlang::sym(paste0("Strain_",..2,"_", ..3-1)),
         "Strain_3rd" = !!rlang::sym(paste0("Strain_",..2,"_", ..3-2))) %>%
    filter(!Strain_final %in% c('Doublet', 'Negative')) %>% 
    filter(!Strain_2nd %in% c('Doublet', 'Negative')) %>%
    filter(!Strain_3rd %in% c('Doublet', 'Negative')) %>%
  dplyr::count(Strain_final, Strain_2nd, Strain_3rd) %>% 
  group_by(Strain_3rd) %>%  
  mutate(st3_f = paste0(Strain_final, collapse = '_'))  %>%
  group_by(Strain_2nd) %>% 
  mutate(st2_f = paste0(Strain_final, collapse = '_')) %>% 
  group_by(Strain_3rd) %>%  
  add_count(name = 'consist_strn') %>% 
  ungroup() %>% 
  mutate_at(vars(matches("st[0-9]_f")), ~ str_split(., "_")) %>%
  mutate_at(vars(matches("st[0-9]_f")), ~ map(., unique)) %>%
  mutate_at(vars(matches("st[0-9]_f")), ~ map_chr(., paste, collapse = "_")) %>% 
  # filter(consist_strn>1) %>% 
  rename(c('Counts' =n, 'Strain' = Strain_final))  %>%
  ggplot(., aes(x = Strain, y = Counts, fill = Strain)) +
  geom_col(position = position_dodge2()) + #width=0.4, position = position_dodge(width=0.5) 
  geom_text(aes(label=Counts), hjust=0.5, vjust= -0.3, position = position_dodge(width = .9), size = 5, show.legend = F, fontface ='bold')+
  scale_fill_manual(values = strain_cols_n) +
  ggh4x::facet_nested_wrap(vars(st3_f, st2_f),  nrow = 1)+
  scale_y_continuous(expand = expansion(mult = c(0.01, 0.24))) +
  scale_x_discrete(expand = expansion(add = c(1, 1))) +
  theme_classic() +
  labs( y = 'Counts')+
  theme(
    axis.text.y  = element_text(size = 14),
    axis.title.y  = element_text(size = 17, face = "bold"),
    axis.text.x  = element_text(size = 14, face = "bold", angle = 90),
    # axis.title.x = element_blank(),
    axis.title.x = element_text(size = 17, face = "bold"),
    legend.text = element_text(size = 17),
    legend.title = element_text(size = 17, face = "bold"),
    legend.position="bottom",
    strip.text =element_text(size=16,face="bold", colour = 'black'),
    panel.border=element_rect(colour="black",fill='transparent')
  ) + 
  guides(fill = guide_legend(title.position = 'left', nrow = 1, name = 'Strain (K8)'))

  })

k_ovlp_bar
```


```{r, fig.width=16, fig.height=4}

sympt_fev_tbl  %>%
  filter(!is.na(fev_sympt)) %>%
  ggplot(., aes(x = donor, y = pseudotime, colour = StrainDon)) +
  geom_sina(size = 0.2) +
  # geom_boxplot() +
  facet_wrap(Hb_Type ~., scales = "free_x")
```


```{r}
# # Create seurat 5 object to solve error "Error in LayerData.Assay5(object = x, layer = DefaultLayer(object = x)[1L],  : features are not found"
# options(Seurat.object.assay.version = "v5")
# 
# m2_asex_harmony_tr_pt_sset_v5 <- CreateSeuratObject(counts = m2_asex_harmony_tr_pt_sset[["RNA"]]$counts, meta.data = m2_asex_harmony_tr_pt_sset@meta.data)

```



!!!!NOTE - AggregateExpression recomended for Seurat v5 but gives error "Error in LayerData.Assay5(object = x, layer = DefaultLayer(object = x)[1L],  : features are not found" discussed here https://github.com/satijalab/seurat/issues/8176
Converting to PseudobulkExpression since not major difference as pointed out here https://github.com/satijalab/seurat/issues/8919
```{r, fig.width=15, fig.height=5}
# visualize by batch and cell type annotation
# pseudobulk the counts based on donor-condition-celltype
# pseudo_m2_asex <- AggregateExpression(m2_asex_harmony_tr_pt_sset, assays = "RNA", return.seurat = T, group.by = c("condition", "fever", "donor", "StageFL", "Hb_Type"))
# pseudo_m2_asex <- AggregateExpression(m2_asex_harmony_tr_pt_sset, assays = "RNA", return.seurat = T, group.by = c("condition", "fever", "donor",  "PT_grp", "Hb_Type"))
pseudo_m2_asex <- AggregateExpression(m2_asex_harmony_tr_pt_sset, assays = "RNA", return.seurat = T, group.by = c("condition", "fever", "donor",  "PT_grp", "Hb_Type"))

# pseudo_m2_asex <- AggregateExpression(m2_asex_harmony_tr_pt_sset_v5, assays = "RNA", return.seurat = T, group.by = c("condition", "fever", "donor",  "PT_grp", "Hb_Type"))

# pseudo_m2_asex <- AggregateExpression(m2_asex_harmony_tr_pt_sset_v5, assays = "RNA", return.seurat = T, group.by = c("condition", "fever", "donor",  "PT_grp", "Hb_Type"))

# each 'cell' is a donor-condition-celltype pseudobulk profile
tail(Cells(pseudo_m2_asex))
```


#####START########################## DESEQ2 ACCOUNTING COVARIATES ###################################

```{r, fig.width=15, fig.height=5}
counts <- pseudo_m2_asex@assays$RNA@layers$counts

colnames(counts) <- colnames(pseudo_m2_asex)
rownames(counts) <- rownames(pseudo_m2_asex)

counts
```

```{r, fig.width=15, fig.height=5}
mdata <- pseudo_m2_asex@meta.data

mdata

```


```{r, fig.width=15, fig.height=5}
## Get only specific PT
pseudo_m2_asex_pt <- pseudo_m2_asex[,pseudo_m2_asex@meta.data$PT_grp == "PT.3"]
counts <- pseudo_m2_asex_pt@assays$RNA@layers$counts

colnames(counts) <- colnames(pseudo_m2_asex_pt)
rownames(counts) <- rownames(pseudo_m2_asex_pt)

counts
```

```{r, fig.width=15, fig.height=5}
mdata <- pseudo_m2_asex_pt@meta.data

mdata

```

Get pseudotime as continous variable rather than chunks
```{r, fig.width=15, fig.height=5}
# pseutotime as continous
pseudo_m2_asex_pt_means <- AggregateExpression(m2_asex_harmony_tr_pt_sset, assays = "RNA", return.seurat = T, group.by = c("condition", "donor"))

```


```{r, fig.width=15, fig.height=5}
counts <- pseudo_m2_asex_pt_means@assays$RNA@layers$counts

colnames(counts) <- colnames(pseudo_m2_asex_pt_means)
rownames(counts) <- rownames(pseudo_m2_asex_pt_means)

counts

mdata <- pseudo_m2_asex_pt_means@meta.data

mdata

```

```{r, fig.width=15, fig.height=5}
m2_asex_harmony_tr_pt_sset@meta.data %>% 
  group_by(donor) %>% 
  mutate(mean_pt = mean(pseudotime) ) %>%
  ungroup() %>%
  count(condition, fever, donor, Hb_Type, parasitaemia_pf, smpl_wait_time, days_frm_ssn_start, age_years, mean_pt) %>% 
  mutate(id = paste(condition, donor, sep = "_"), 
         "cell_count" =n) %>%
  select(id, cell_count, parasitaemia_pf, smpl_wait_time, days_frm_ssn_start, age_years, mean_pt)

```

```{r, fig.width=15, fig.height=5}
## Get cell count
sympt_cell_count <- m2_asex_harmony_tr_pt_sset@meta.data %>% 
  group_by(donor) %>% 
  mutate(mean_pt = mean(pseudotime) ) %>%
  ungroup() %>%
  count(condition, fever, donor, Hb_Type, parasitaemia_pf, smpl_wait_time, days_frm_ssn_start, age_years, mean_pt) %>% 
  mutate(id = paste(condition, donor, sep = "_"), 
         "cell_count" =n) %>%
  select(id, fever, Hb_Type, cell_count, parasitaemia_pf, smpl_wait_time, days_frm_ssn_start, age_years, mean_pt)

sympt_mdata_all <- mdata %>% left_join(., sympt_cell_count, by = c("orig.ident" = "id"))
```


Scale metadata since some numeric variables result in warnings on colinearity - "Including numeric variables with large mean can induce collinearity with the intercept. Users should center and scale numeric variables in the design to improve GLM convergence".
```{r, fig.width=15, fig.height=5}
## Get cell count
# sympt_mdata_all$parasitaemia_pf_scaled <- scale(sympt_mdata_all$parasitaemia_pf, center = TRUE, scale = TRUE)

sympt_mdata_all <- sympt_mdata_all %>% 
  mutate(#PT_grp = factor(PT_grp, levels = c("PT.1", "PT.2", "PT.3")),
         condition = factor(condition, levels = c("Asymptomatic", "Symptomatic")),
         fever = factor(fever, levels = c("NoFever", "Fever")),
         parasitaemia_pf_scaled = scale(parasitaemia_pf, center = TRUE, scale = TRUE),
         age_years_scaled = scale(age_years, center = TRUE, scale = TRUE),
         smpl_wait_time_scaled = scale(smpl_wait_time, center = TRUE, scale = TRUE),
         mean_pt_scaled = scale(mean_pt, center = TRUE, scale = TRUE))
```


#### Investigate whether to split dataset into several PT_grps (Pseudotime groups) or perform one analysis adjusting for PT_grps as a covariate
```{r, fig.width=15, fig.height=5}
## Investigate whether to split dataset into several PT_grps (Pseudotime groups) or perform one analysis adjusting for PT_grps as a covariate

# Create DESeq2 dataset with interaction term
dds_full <- DESeqDataSetFromMatrix(counts, 
                              colData = sympt_mdata_all, 
                              # design = ~ condition +  fever + parasitaemia_pf + age_years,
                              # design = ~ condition,
                              # design = ~ condition + PT_grp + condition:PT_grp,
                              design = ~ condition + mean_pt + condition:mean_pt
                              )

# Run DESeq2 with full model
# dds_interaction <- DESeq(dds_full, test = "LRT", reduced = ~ condition + PT_grp)
dds_interaction <- DESeq(dds_full, test = "LRT", reduced = ~ condition + mean_pt)


# Extract results to identify genes with significant interaction effects
res_interaction <- results(dds_interaction)

# Inspect top significant genes for interaction
res_interaction <- res_interaction[order(res_interaction$padj), ]  # Order by adjusted p-value
res_interaction %>% as.data.frame() %>% arrange(padj)

# Interpreting Results
# 
#     The LRT tests whether including the pseudo_groups:condition interaction term significantly improves the model.
#     Genes with low adjusted p-values (padj < 0.05) have stage-dependent condition effects, meaning the effect of condition varies across pseudo_groups.
#     If many genes have significant interaction effects, it suggests that DE patterns differ across developmental stages, and analyzing each stage separately may be beneficial.

# VERDICT - Retain full dataset and just adjust as a covariate

```


```{r, fig.width=15, fig.height=5}
# Create DESeq2 object        
dds <- DESeqDataSetFromMatrix(counts, 
                              colData = sympt_mdata_all, 
                              # design = ~ condition +  fever + parasitaemia_pf + age_years,
                              # design = ~ condition,
                              # design = ~ condition + mean_pt + fever + parasitaemia_pf_scaled + age_years_scaled,
                              design = ~ condition + mean_pt_scaled + smpl_wait_time_scaled +  parasitaemia_pf_scaled + age_years_scaled
                              )
```

```{r, fig.width=15, fig.height=5}
# Transform counts for data visualization
rld <- rlog(dds, blind=TRUE)

# Plot PCA
DESeq2::plotPCA(rld, ntop = 500, intgroup = "condition")
DESeq2::plotPCA(rld, ntop = 500, intgroup = "mean_pt")
```





```{r, fig.width=15, fig.height=5}
DESeq2::plotPCA(rld, ntop = 500, intgroup = "cell_count")
DESeq2::plotPCA(rld, ntop = 500, intgroup = "parasitaemia_pf")
DESeq2::plotPCA(rld, ntop = 500, intgroup = "age_years")
DESeq2::plotPCA(rld, ntop = 500, intgroup = "days_frm_ssn_start")
```


```{r, fig.width=10, fig.height=8}
# Extract the rlog matrix from the object and compute pairwise correlation values
rld_mat <- assay(rld)
rld_cor <- cor(rld_mat)

# Plot heatmap
ComplexHeatmap::pheatmap(rld_cor, annotation = sympt_mdata_all[, c("condition"), drop=F])
```


```{r, fig.width=15, fig.height=5}
# Run DESeq2 differential expression analysis
dds <- DESeq(dds)
```


```{r, fig.width=15, fig.height=5}
# Plot dispersion estimates
plotDispEsts(dds)
```


```{r, fig.width=15, fig.height=5}
# library(apeglm)
# Check the coefficients for the comparison
resultsNames(dds)

# Generate results object
res <- results(dds, 
               name = "condition_Symptomatic_vs_Asymptomatic",
               alpha = 0.05)

# Shrink the log2 fold changes to be more appropriate using the apeglm method - should cite [paper]() when using this method
res <- lfcShrink(dds, 
                 coef = "condition_Symptomatic_vs_Asymptomatic",
                 res=res,
                 # type = "apeglm",
                 type = "normal")
```


```{r, fig.width=15, fig.height=5}
# Turn the DESeq2 results object into a tibble for use with tidyverse functions
res_tbl <- res %>%
  data.frame() %>%
  rownames_to_column(var = "gene") %>%
  as_tibble() %>%
  arrange(padj) #%>%
  # left_join(., Pf3D7_genes_plasmoDB[, c("gene_id", "Gene")], by = c("gene" = "gene_id")) %>%
  # mutate("gene" = Gene) %>%
  # select(-c(Gene))

# Check results output
res_tbl 

```


```{r, fig.width=15, fig.height=5}
# Set thresholds
padj_cutoff <- 0.05
# padj_cutoff <- 0.001

# Subset the significant results
sig_res <- dplyr::filter(res_tbl, padj < padj_cutoff) %>%
  dplyr::arrange(padj)

# Check significant genes output
sig_res

```

```{r, fig.width=15, fig.height=5}
# Set thresholds
log2fc_cutoff <- 0.58

# Count significantly up/down genes above threshold
n_sig_up <- dplyr::filter(sig_res, log2FoldChange >= log2fc_cutoff) %>% 
  nrow()
n_sig_up

n_sig_dn <- dplyr::filter(sig_res, log2FoldChange <= -log2fc_cutoff) %>% 
  nrow()
n_sig_dn
```

```{r, fig.width=15, fig.height=5}
# Scatterplot

## Extract normalized counts from dds object
normalized_counts <- counts(dds, normalized = TRUE)

## Extract top 20 DEG from resLFC (make sure to order by padj)
top20_sig_genes <- sig_res %>%
  dplyr::arrange(padj) %>%
  dplyr::pull(gene) %>%
  head(n = 20)

## Extract matching normalized count values from matrix
top20_sig_counts <- normalized_counts[rownames(normalized_counts) %in% top20_sig_genes, ]
top20_sig_counts

## Convert wide matrix to long data frame for ggplot2
top20_sig_df <- data.frame(top20_sig_counts)
top20_sig_df$gene <- rownames(top20_sig_counts)

top20_sig_df <- reshape2::melt(data.table::setDT(top20_sig_df), 
                     id.vars = c("gene"),
                     variable.name = "cluster_sample_id") %>% 
  data.frame()

top20_sig_df

## Replace "." by " " in cluster_sample_id variable (melt() introduced the ".")
top20_sig_df$cluster_sample_id <- gsub(" ","\\.", top20_sig_df$cluster_sample_id)
top20_sig_df

## Join counts data frame with metadata
top20_sig_df <- dplyr::left_join(top20_sig_df, base::as.data.frame(colData(dds)),
                           by = c("cluster_sample_id" = "orig.ident"))
top20_sig_df

## Generate plot
ggplot(top20_sig_df, aes(y = value, x = condition, col = condition)) +
  geom_jitter(height = 0, width = 0.15) +
  scale_y_continuous(trans = 'log10') +
  ylab("log10 of normalized expression level") +
  xlab("condition") +
  ggtitle("Top 20 Significant DE Genes") +
  theme(plot.title = element_text(hjust = 0.5)) +
  facet_wrap(~ gene)
```


```{r, fig.width=8, fig.height=5}
# Heatmap

# ## Extract normalized counts for significant genes only
# sig_counts <- normalized_counts[rownames(normalized_counts) %in% sig_res$gene, ]

## Extract normalized counts for top 20 significant genes only
sig_counts <- normalized_counts[rownames(normalized_counts) %in% top20_sig_genes, ]

sig_counts_gn_names <- Pf3D7_genes_plasmoDB %>% select(gene_id, Gene ) %>% column_to_rownames("gene_id") %>% .[rownames(sig_counts),]
rownames(sig_counts) <- sig_counts_gn_names

## Set a color-blind friendly palette
heat_colors <- rev(brewer.pal(11, "PuOr"))

## Run pheatmap using the metadata data frame for the annotation
ComplexHeatmap::pheatmap(sig_counts, 
         color = heat_colors, 
         cluster_rows = TRUE, 
         cluster_cols = F, 
         show_rownames = T,
         annotation_col = sympt_mdata_all[, c("condition", "fever")], 
         border_color = NA, 
         fontsize = 10, 
         scale = "row", 
         fontsize_row = 10, 
         height = 20)  
```


```{r, fig.width=8, fig.height=12}

# Volcano plot
res_table_thres <- res_tbl[!is.na(res_tbl$padj), ] %>% 
  mutate(threshold = padj < padj_cutoff & abs(log2FoldChange) >= log2fc_cutoff)
min(log10(res_table_thres$padj))

## Generate plot
ggplot(res_table_thres) +
  geom_point(aes(x = log2FoldChange, y = -log10(padj), colour = threshold)) +
  ggtitle("Volcano plot of stimulated B cells relative to control") +
  xlab("log2 fold change") +
  xlim(-4.5, 12) +
  ylab("-log10 adjusted p-value") +
  # scale_y_continuous(limits = c(0, 250)) +
  scale_color_manual(values = c("grey60", "red3")) +
  theme(legend.position = "none",
        plot.title = element_text(size = rel(1.3), hjust = 0.5),
        axis.title = element_text(size = rel(1.15)))                    


```

```{r, fig.width=8, fig.height=12}
# Heatmap

## Extract normalized counts for significant genes only
sig_counts <- normalized_counts[rownames(normalized_counts) %in% sig_res$gene, ]

## Set a color-blind friendly palette
heat_colors <- rev(brewer.pal(11, "PuOr"))

## Run pheatmap using the metadata data frame for the annotation
# ComplexHeatmap::pheatmap(sig_counts, 
#          color = heat_colors, 
#          cluster_rows = TRUE, 
#          cluster_cols = F, 
#          show_rownames = FALSE,
#          annotation_col = sympt_mdata_all[, c("condition", "fever", "PT_grp")], 
#          border_color = NA, 
#          fontsize = 10, 
#          scale = "row", 
#          fontsize_row = 10, 
#          height = 20)  
```


#####END########################## DESEQ2 ACCOUNTING COVARIATES ###################################


!!! NOTE - AggregateExpression doesn't work with seurat 5.0.3 so ran it using 5.0.1 and then read it back

```{r}
# gogo()
# saveRDS(m2_asex_harmony_tr_pt_sset, "data/processed/Pf/m2_all/m2_asex_harmony_tr_pt_sset.RDS")

```

```{r}
# gogo()
# pseudo_m2_asex <- readRDS("data/processed/Pf/m2_all/m2_asex_harmony_tr_pt_sset_pb.RDS")

```


```{r}
tail(Cells(pseudo_m2_asex))
```

```{r, fig.width=15, fig.height=5}
## Symptomatic disregarding fever

# batch by stage and symptomatic
pseudo_m2_asex$stage.hb <- paste(pseudo_m2_asex$PT_grp, pseudo_m2_asex$Hb_Type, sep = "_")
Idents(pseudo_m2_asex) <- "stage.hb"

# DE
bulk.de.hb <- map(c(paste0("PT.", 1:3)), ~ FindAllMarkers(object = pseudo_m2_asex[,pseudo_m2_asex$PT_grp == .x], test.use = "DESeq2", min.cells.group = 1) %>% rename('gene_id' = gene)) %>%
  set_names(c(paste0("PT.", 1:3)))

map(bulk.de.hb, ~.x %>% filter(p_val_adj < 0.05) %>% head( n = 10) %>% left_join(., Pf3D7_genes_plasmoDB%>% select(gene_id, Gene)))

```


```{r}

Idents(pseudo_m2_asex) <- "stage.hb"

imap(bulk.de.hb, ~{
  gns <- .x %>% filter(p_val_adj < 0.05) %>% pull(gene_id)
  stg <- .y
  # map(gns, ~VlnPlot(pseudo_m2_asex, features = .x, group.by = "stage.sympt", ncol = 1) )
  map(gns, ~VlnPlot(pseudo_m2_asex, features = .x, group.by = "stage.hb", ncol = 1) )
  })

```

PF3D7-1229300 https://www-ncbi-nlm-nih-gov.ezp.lib.cam.ac.uk/pmc/articles/PMC8527989/
PF3D7_1229200 https://www-ncbi-nlm-nih-gov.ezp.lib.cam.ac.uk/pmc/articles/PMC4286016/
https://www-ncbi-nlm-nih-gov.ezp.lib.cam.ac.uk/pmc/articles/PMC8527989/


```{r, fig.width=15, fig.height=5}
## Symptomatic disregarding fever

# batch by stage and symptomatic
pseudo_m2_asex$stage.sympt <- paste(pseudo_m2_asex$PT_grp, pseudo_m2_asex$condition, sep = "_")
Idents(pseudo_m2_asex) <- "stage.sympt"

# DE
bulk.de <- map(c(paste0("PT.", 1:3)), ~ FindMarkers(object = pseudo_m2_asex,  ident.1 = paste0(.x, "_Symptomatic"),   ident.2 = paste0(.x, "_Asymptomatic"), test.use = "DESeq2") %>% rownames_to_column('gene_id')) %>%
  set_names(c(paste0("PT.", 1:3)))

map(bulk.de, ~.x %>% head( n = 10) %>% left_join(Pf3D7_genes_plasmoDB%>% select(gene_id, Gene)))
```

```{r, fig.width=15, fig.height=5}
## Symptomatic disregarding fever

# batch by stage, symptomatic & fever
pseudo_m2_asex$stage.sympt.fever <- paste(pseudo_m2_asex$PT_grp, pseudo_m2_asex$condition, pseudo_m2_asex$fever,  sep = "_")
Idents(pseudo_m2_asex) <- "stage.sympt.fever"

# DE
bulk.de.SF <- map(c(paste0("PT.", 1:3)), ~ FindMarkers(object = pseudo_m2_asex,  ident.1 = paste0(.x, "_Symptomatic_Fever"),   ident.2 = paste0(.x, "_Asymptomatic_NoFever"), test.use = "DESeq2", min.cells.group = 2) %>% rownames_to_column('gene_id')) %>%
  set_names(c(paste0("PT.", 1:3)))

map(bulk.de.SF, ~.x %>% head( n = 10)  %>% left_join(Pf3D7_genes_plasmoDB%>% select(gene_id, Gene)))
```

PF3D7_1122900, ABCF1 - Apicoplast link to fever https://www-ncbi-nlm-nih-gov.ezp.lib.cam.ac.uk/pmc/articles/PMC8316339/, https://www-ncbi-nlm-nih-gov.ezp.lib.cam.ac.uk/pmc/articles/PMC6155542/
https://www.cell.com/trends/parasitology/fulltext/S1471-4922(22)00039-3?rss=yes#supplementaryMaterial 


```{r, fig.width=15, fig.height=5}
## Symptomatic disregarding fever

# batch by stage, fever
pseudo_m2_asex$stage.fever <- paste(pseudo_m2_asex$PT_grp, pseudo_m2_asex$fever,  sep = "_")
Idents(pseudo_m2_asex) <- "stage.fever"

# DE
bulk.de.F <- map(c(paste0("PT.", 1:3)), ~ FindMarkers(object = pseudo_m2_asex,  ident.1 = paste0(.x, "_Fever"),   ident.2 = paste0(.x, "_NoFever"), test.use = "DESeq2") %>% rownames_to_column('gene_id')) %>%
  set_names(c(paste0("PT.", 1:3)))

map(bulk.de.F, ~.x %>% head( n = 10)  %>% left_join(Pf3D7_genes_plasmoDB%>% select(gene_id, Gene)))
```


PF3D7-1023100 - https://www.biorxiv.org/content/10.1101/2020.12.10.419788v2.full
We found that HS-Sensitive mutants tended to be sensitive to both artemisinin derivatives and H2O2-induced oxidative stress, while HS-Tolerant mutants were less sensitive to either condition
In addition to the HS-Sensitive mutant PB4 (DHA, dynein, PF3D7_1122900) from the pilot library, we identified three HS-Sensitive dynein mutants in the 1K-library (two different mutants of DHA_12, PF3D7_1202300; one mutant of DHA_10, PF3D7_1023100), indicating the robusticity of the 1K-library screen (Fig. S8A- B).

#https://www.ncbi.nlm.nih.gov/pmc/articles/PMC6650087/ - PF3D7-1240400
#https://journals.asm.org/doi/10.1128/mbio.01636-21 https://www.frontiersin.org/articles/10.3389/fmala.2023.1075755/full  - hyp10
#http://dr.iiserpune.ac.in:8080/xmlui/bitstream/handle/123456789/7531/20153385_Abhishek_Kanyal_PhD_Thesis.pdf?sequence=1&isAllowed=y - PF3D7-1401300




```{r}
pmap(list(list("stage.sympt", "stage.sympt.fever", "stage.fever", "stage.hb"), list(bulk.de, bulk.de.SF, bulk.de.F, bulk.de.hb), list(c("_Symptomatic", "_Asymptomatic"), c("_Symptomatic_Fever", "_Asymptomatic_NoFever"), c("_Fever", "_NoFever"), c("_HbC", "_HbN"))), ~{ 
  condtn = ..1
  Idents(pseudo_m2_asex) <- condtn
  obj = ..2
  grp = ..3

  imap(obj, ~{
    gns <- .x %>% filter(p_val_adj < 0.05) %>% pull(gene_id)
    stg <- .y
    map(gns, ~VlnPlot(pseudo_m2_asex, features = .x, idents = c(paste0(stg, grp[1]), paste0(stg, grp[2])), group.by = condtn, ncol = 1) + labs(title= paste0(.x, grp)) )
    })
})
```


```{r, fig.width=15, fig.height=5}
# create a new column to annotate sample-condition-celltype in the single-cell dataset
m2_asex_harmony_tr_pt_sset$sympt_don <- paste(m2_asex_harmony_tr_pt_sset$condition, m2_asex_harmony_tr_pt_sset$donor, sep = "_")
m2_asex_harmony_tr_pt_sset$sympt_fever_don <- paste(m2_asex_harmony_tr_pt_sset$condition, m2_asex_harmony_tr_pt_sset$fever, m2_asex_harmony_tr_pt_sset$donor, sep = "_")
m2_asex_harmony_tr_pt_sset$fever_don <- paste(m2_asex_harmony_tr_pt_sset$fever, m2_asex_harmony_tr_pt_sset$donor, sep = "_")

m2_asex_harmony_tr_pt_sset$stage.sympt <- paste(m2_asex_harmony_tr_pt_sset$PT_grp, m2_asex_harmony_tr_pt_sset$condition, sep = "_")
m2_asex_harmony_tr_pt_sset$stage.sympt.fever <- paste(m2_asex_harmony_tr_pt_sset$PT_grp, m2_asex_harmony_tr_pt_sset$condition, m2_asex_harmony_tr_pt_sset$fever, sep = "_")
m2_asex_harmony_tr_pt_sset$stage.fever <- paste(m2_asex_harmony_tr_pt_sset$PT_grp, m2_asex_harmony_tr_pt_sset$fever, sep = "_")

```



```{r, fig.width= 14, fig.height=4}
pmap(list(list("stage.sympt", "stage.sympt.fever", "stage.fever"), list(bulk.de, bulk.de.SF, bulk.de.F), list(c("_Symptomatic", "_Asymptomatic"), c("_Symptomatic_Fever", "_Asymptomatic_NoFever"), c("_Fever", "_NoFever")), list("sympt_don", "sympt_fever_don", "fever_don")), ~{ 
  condtn = ..1
  Idents(m2_asex_harmony_tr_pt_sset) <- condtn
  obj = ..2
  grp = ..3
  grp_condtn = ..4

  imap(obj, ~{
    gns <- .x %>% filter(p_val_adj < 0.05) %>% pull(gene_id)
    stg <- .y
    map(gns, ~VlnPlot(m2_asex_harmony_tr_pt_sset, features = .x, idents = c(paste0(stg, grp[1]), paste0(stg, grp[2])), group.by = grp_condtn, ncol = 1) + labs(title= paste0(.x, grp)) )
    })
})
```


```{r, fig.width= 14, fig.height=4}
pmap(list(list("stage.sympt", "stage.sympt.fever", "stage.fever"), list(bulk.de, bulk.de.SF, bulk.de.F), list("sympt_don", "sympt_fever_don", "fever_don")), ~{ 
  condtn = ..1
  Idents(m2_asex_harmony_tr_pt_sset) <- condtn
  obj = ..2
  grp_condtn = ..3

  imap(obj, possibly(~{
    gns <- .x %>% filter(p_val_adj < 0.05) %>% pull(gene_id) 
    stg <- .y
    DotPlot(m2_asex_harmony_tr_pt_sset, features = gns, group.by = grp_condtn) + labs(title =  stg) 
    }), 
    "No significant genes"
    )
})
```

```{r, fig.width= 14, fig.height=4}
pmap(list(list(bulk.de, bulk.de.SF, bulk.de.F), list("stage.sympt", "stage.sympt.fever", "stage.fever")), ~{ 
  obj = ..1
  grp_condtn = ..2

  imap(obj, possibly(~{
    DotPlot(m2_asex_harmony_tr_pt_sset, features = .x %>% filter(p_val_adj < 0.05) %>% pull(gene_id), group.by = grp_condtn) + labs(title =  .y) 
    }), 
    "No significant genes"
    )
})
```


```{r, fig.width=15, fig.height=5}
# create a new column to annotate sample-condition-celltype in the single-cell dataset
m2_asex_harmony_tr_pt_sset$sympt_fever <- paste(m2_asex_harmony_tr_pt_sset$condition, m2_asex_harmony_tr_pt_sset$fever, sep = "_")

```

Plot collapsing the donors by the different conditions
```{r, fig.width= 16, fig.height=4}
pmap(list(list("stage.sympt", "stage.sympt.fever", "stage.fever", "stage.hb"), list(bulk.de, bulk.de.SF, bulk.de.F, bulk.de.hb), list("condition", "sympt_fever", "fever", "Hb_Type")), ~{ 
  condtn = ..1
  Idents(m2_asex_harmony_tr_pt_sset) <- condtn
  obj = ..2
  grp_condtn = ..3

  imap(obj, ~{
    gns <- .x %>% filter(p_val_adj < 0.05) %>% pull(gene_id)
    stg <- .y
    # map(gns, ~VlnPlot(m2_asex_harmony_tr_pt_sset, features = .x, idents = c(paste0(stg, grp[1]), paste0(stg, grp[2])), group.by = grp_condtn, ncol = 1) + labs(title= paste0(.x, grp)) )
    map(gns, ~FeaturePlot(m2_asex_harmony_tr_pt_sset[,m2_asex_harmony_tr_pt_sset$PT_grp == stg],  reduction = "umap.harmony_PC_man", features = .x, split.by = grp_condtn, order = T))
    })
})
```



```{r, fig.width=10, fig.height=5}
DimPlot(m2_asex_harmony_tr_pt_sset, reduction = "umap.harmony_PC_man", group.by = c("condition", "PT_grp"), ncol = 2)
DimPlot(m2_asex_harmony_tr_pt_sset[,m2_asex_harmony_tr_pt_sset$PT_grp == "PT.1"], reduction = "umap.harmony_PC_man", split.by = c("condition"), ncol = 2)
DimPlot(m2_asex_harmony_tr_pt_sset[,m2_asex_harmony_tr_pt_sset$PT_grp == "PT.2"], reduction = "umap.harmony_PC_man", split.by = c("condition"), ncol = 2)
DimPlot(m2_asex_harmony_tr_pt_sset[,m2_asex_harmony_tr_pt_sset$PT_grp == "PT.3"], reduction = "umap.harmony_PC_man", split.by = c("condition"), ncol = 2)

```



```{r, fig.width=10, fig.height=5}
FeaturePlot(m2_asex_harmony_tr_pt_sset,  reduction = "umap.harmony_PC_man", features = "PF3D7-0114500", split.by = c("condition"), ncol = 2, order = T)
FeaturePlot(m2_asex_harmony_tr_pt_sset,  reduction = "umap.harmony_PC_man", features = "PF3D7-1240400", split.by = c("condition"), ncol = 2, order = T)
```


```{r, fig.width=7.5, fig.height=2}
key_gns_exp <- FetchData(object = m2_asex_harmony_tr_pt_sset, vars = c('PF3D7-1240400', 'PF3D7-1229200','PF3D7-1229300', 'PF3D7-1107800'), slot = "data") 

```

```{r, fig.width=7.5, fig.height=2}
tbl_umap_gn <- m2_asex_harmony_tr_pt_sset@reductions$umap.harmony_PC_man@cell.embeddings %>% as.data.frame() %>% rownames_to_column("bcode") %>% left_join(., m2_asex_harmony_tr_pt_sset@meta.data %>% rownames_to_column("bcode")) %>%  left_join(., key_gns_exp %>% rownames_to_column("bcode"))

```


```{r, fig.width=7.3, fig.height=2.0}
# custom_colors <- colorRampPalette(c("lightgrey", "blue"))(100)

tbl_umap_gn %>%
  mutate(across(donor, ~str_remove_all(., "SC|C"))) %>%
  arrange(`PF3D7-1240400`) %>%
  # ggplot(., aes(x = weave_factors(Strain, donor), y = Counts, fill = donor)) +
  ggplot(., aes(x = umapharmonyPCman_1, y = umapharmonyPCman_2, colour = `PF3D7-1240400`))+
  geom_point(size = 0.1) +
  scale_color_distiller( direction = 1) +
  ggh4x::facet_nested_wrap(vars(condition, donor),  nrow = 1, scales = "free_x") +
  theme_classic()+
  theme(legend.position = "right", 
        legend.title = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        strip.text =element_text(size=12,face="bold", colour = 'black'))+
  # guides(x = "axis_nested") +
  labs(x = "UMAP_1", y="UMAP_2", title = "VAR_PF3D7_1240400") 
```


```{r, fig.width=7.3, fig.height=2.0}
# custom_colors <- colorRampPalette(c("lightgrey", "blue"))(100)

tbl_umap_gn %>%
  mutate(across(donor, ~str_remove_all(., "SC|C"))) %>%
  arrange(`PF3D7-1240400`) %>%
  # ggplot(., aes(x = weave_factors(Strain, donor), y = Counts, fill = donor)) +
  ggplot(., aes(x = umapharmonyPCman_1, y = umapharmonyPCman_2, colour = `PF3D7-1240400`))+
  geom_point(size = 0.1) +
  scale_color_distiller( direction = 1) +
  ggh4x::facet_nested_wrap(vars(condition, donor),  nrow = 1, scales = "free_x") +
  theme_classic()+
  theme(legend.position = "right", 
        legend.title = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        strip.text =element_text(size=12,face="bold", colour = 'black'))+
  # guides(x = "axis_nested") +
  labs(x = "UMAP_1", y="UMAP_2", title = "VAR_PF3D7_1240400") 
```

```{r, fig.width=6.8, fig.height=2.0}
# custom_colors <- colorRampPalette(c("lightgrey", "blue"))(100)

tbl_umap_gn %>%
  mutate(across(donor, ~str_remove_all(., "SC|C"))) %>%
  arrange(`PF3D7-1107800`) %>%
  # ggplot(., aes(x = weave_factors(Strain, donor), y = Counts, fill = donor)) +
  ggplot(., aes(x = umapharmonyPCman_1, y = umapharmonyPCman_2, colour = `PF3D7-1107800`))+
  geom_point(size = 0.1) +
  scale_color_distiller( direction = 1) +
  ggh4x::facet_nested_wrap(vars(condition, donor),  nrow = 1, scales = "free_x") +
  theme_classic()+
  theme(legend.position = "right", 
        legend.title = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        strip.text =element_text(size=12,face="bold", colour = 'black'))+
  # guides(x = "axis_nested") +
  labs(x = "UMAP_1", y="UMAP_2", title = "VAR_PF3D7_1107800") 
```


```{r, fig.width=5.8, fig.height=2.8}
# custom_colors <- colorRampPalette(c("lightgrey", "blue"))(100)

tbl_umap_gn %>%
  mutate(across(donor, ~str_remove_all(., "SC|C"))) %>%
  arrange(`PF3D7-1107800`) %>%
  # ggplot(., aes(x = weave_factors(Strain, donor), y = Counts, fill = donor)) +
  ggplot(., aes(x = donor, y = `PF3D7-1107800`))+
  geom_boxplot() +
  scale_color_distiller( direction = 1) +
  ggh4x::facet_nested_wrap(vars(condition),  nrow = 1, scales = "free_x") +
  theme_classic()+
  theme(legend.position = "bottom", 
        # legend.position = "none", 
        axis.text = element_blank(),
        strip.text =element_text(size=12,face="bold", colour = 'black'))+
  # guides(x = "axis_nested") +
  labs(x = "UMAP_1", y="UMAP_2", title = "VAR_PF3D7_1107800") 
```


```{r, fig.width=7.5, fig.height=3}
# custom_colors <- colorRampPalette(c("lightgrey", "blue"))(100)

# tbl_umap_gn %>%ds <- DESeqDataSetFromMatrix(counts
#   mutate(across(donor, ~str_remove_all(., "SC|C")),
#          across(Hb_Type, ~factor(.x, levels = c("HbC", "HbS", "HbN")))) %>%
#   arrange(`PF3D7-1229300`) %>%
#   # ggplot(., aes(x = weave_factors(Strain, donor), y = Counts, fill = donor)) +
#   ggplot(., aes(x = umapharmonyPCman_1, y = umapharmonyPCman_2, colour = `PF3D7-1229300`))+
#   geom_point(size = 0.1) +
#   scale_color_distiller( direction = 1) +
#   ggh4x::facet_nested_wrap(vars(Hb_Type, donor),  nrow = 1, scales = "free_x") +
#   theme_classic()+
#   theme(legend.position = "right", 
#         legend.title = element_blank(),
#         axis.text = element_blank(),
#         axis.ticks = element_blank(),
#         strip.text =element_text(size=12,face="bold", colour = 'black'))+
#   # guides(x = "axis_nested") +
#   labs(x = "UMAP_1", y="UMAP_2", title = "PF3D7-1229300") 
```

```{r, fig.width=7.5, fig.height=3}
# custom_colors <- colorRampPalette(c("lightgrey", "blue"))(100)

tbl_umap_gn %>%
  mutate(across(donor, ~str_remove_all(., "SC|C")),
         across(Hb_Type, ~factor(.x, levels = c("HbC", "HbS", "HbN")))) %>%
  arrange(`PF3D7-1229300`) %>%
  # ggplot(., aes(x = weave_factors(Strain, donor), y = Counts, fill = donor)) +
  ggplot(., aes(x = umapharmonyPCman_1, y = umapharmonyPCman_2, colour = `PF3D7-1229300`))+
  geom_point(size = 0.1) +
  scale_color_distiller( direction = 1) +
  ggh4x::facet_nested_wrap(vars(Hb_Type, donor),  nrow = 1, scales = "free_x") +
  theme_classic()+
  theme(legend.position = "bottom", 
        axis.text = element_blank(),
        strip.text =element_text(size=12,face="bold", colour = 'black'))+
  # guides(x = "axis_nested") +
  labs(x = "UMAP_1", y="UMAP_2", title = "PF3D7_1229300") 
```


```{r, fig.width=6.8, fig.height=3.0}
# custom_colors <- colorRampPalette(c("lightgrey", "blue"))(100)

tbl_umap_gn %>% 
  pivot_longer(cols = starts_with("PF3D7-"), names_to = "Gene", values_to = "Expression") %>%
  mutate(across(donor, ~str_remove_all(., "SC|C")),
         across(Hb_Type, ~factor(.x, levels = c("HbC", "HbS", "HbN")))) %>%
  filter(Gene %in% c("PF3D7-1229200", "PF3D7-1229300")) %>%
  group_by(Gene) %>%
  arrange(Expression, .by_group = T) %>%
  # ggplot(., aes(x = weave_factors(Strain, donor), y = Counts, fill = donor)) +
  ggplot(., aes(x = umapharmonyPCman_1, y = umapharmonyPCman_2, colour = Expression))+
  geom_point(size = 0.1) +
  scale_color_distiller( direction = 1) +
  ggh4x::facet_nested(Gene ~ Hb_Type+ donor, scales = "free_x") +
  theme_classic()+
  theme(legend.position = "right", 
        legend.title = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        strip.text =element_text(size=12,face="bold", colour = 'black'))+
  # guides(x = "axis_nested") +
  labs(x = "UMAP_1", y="UMAP_2") 
```

################## START  edgeR DE ################## 

https://bioconductor.org/packages/release/bioc/vignettes/edgeR/inst/doc/edgeRUsersGuide.pdf

```{r, fig.width=15, fig.height=5}
## Create pseudo-bulk samples
y <- Seurat2PB(m2_asex_harmony_tr_pt_sset[,m2_asex_harmony_tr_pt_sset$PT_grp == "PT.3"], sample="donor", cluster="condition")
# m2_asex_harmony_tr_pt_sset, assays = "RNA", return.seurat = T, group.by = c("condition", "fever", "donor",  "PT_grp", "Hb_Type"

dim(y)

head(y$samples, n=10L)
```


!! NOTE edgeR PSEUDOBULK BY SYMPTOMATIC STATUS DOESN'T GIVE DE GENES - SO PSEUDOBULKING BY PSEUDOTIME
```{r, fig.width=15, fig.height=5}
## Create pseudo-bulk samples
y <- Seurat2PB(m2_asex_harmony_tr_pt_sset, sample="donor", cluster=c("PT_grp"))
# m2_asex_harmony_tr_pt_sset, assays = "RNA", return.seurat = T, group.by = c("condition", "fever", "donor",  "PT_grp", "Hb_Type"

dim(y)

head(y$samples, n=10L)
```

Filtering and normalization

```{r, fig.width=15, fig.height=5}
## filter examine the library size
summary(y$samples$lib.size)

keep.samples <- y$samples$lib.size > 5e4
table(keep.samples)
```



```{r, fig.width=15, fig.height=5}
## filter by gene expression
keep.genes <- filterByExpr(y, group=y$samples$cluster)

table(keep.genes)

y <- y[keep.genes, , keep=FALSE]
```


```{r, fig.width=15, fig.height=5}
## 
y <- normLibSizes(y)
head(y$samples, n=10L)

summary(y$samples$norm.factors)
```


```{r}
## Visualize
cluster <- as.factor(y$samples$cluster)
limma::plotMDS(y, pch=16, col=c(2:8)[cluster], main="MDS")
legend("bottomleft", legend=paste0("cluster",levels(cluster)), pch=16, col=2:8, cex=0.8)
```

```{r, fig.width=15, fig.height=5}
## Design matrix
donor <- factor(y$samples$sample)
design <- model.matrix(~ cluster + donor )
colnames(design) <- gsub("donor", "", colnames(design))
colnames(design)[1] <- "Int"
head(design)
```

```{r, fig.width=15, fig.height=5}
## Dispersion estimation
y <- estimateDisp(y, design, robust=TRUE)
y$common.dispersion

plotBCV(y)
```


```{r, fig.width=15, fig.height=5}
fit <- glmQLFit(y, design, robust=TRUE)
plotQLDisp(fit)
```


```{r, fig.width=15, fig.height=5}
## Marker genes identification
ncls <- nlevels(cluster)
contr <- rbind( matrix(1/(1-ncls), ncls, ncls),matrix(0, ncol(design)-ncls, ncls) )
diag(contr) <- 1
contr[1,] <- 0
rownames(contr) <- colnames(design)
colnames(contr) <- paste0("cluster", levels(cluster))
contr
```


```{r, fig.width=15, fig.height=5}
## quasi-likelihood F-test for each testing contras
qlf <- list()
for(i in 1:ncls){
 qlf[[i]] <- glmQLFTest(fit, contrast=contr[,i])
 qlf[[i]]$comparison <- paste0("cluster", levels(cluster)[i], "_vs_others")
}


```


```{r, fig.width=15, fig.height=5}
## top most significant DE genes of cluster 0 vs other clusters
topTags(qlf[[1]], n=10L)
```



```{r, fig.width=15, fig.height=5}
## numbers of DE genes under each comparison

dt <- lapply(lapply(qlf, decideTests), summary)
dt.all <- do.call("cbind", dt)
dt.all

```


```{r, fig.width=15, fig.height=5}
## top 20 marker (up-regulated) genes for each cluster
top <- 20
topMarkers <- list()
for(i in 1:ncls) {
 ord <- order(qlf[[i]]$table$PValue, decreasing=FALSE)
 up <- qlf[[i]]$table$logFC[ord] > 0
 topMarkers[[i]] <- rownames(y)[ord[up][1:top]]
 }
topMarkers <- unique(unlist(topMarkers))
topMarkers
```


```{r, fig.width=6, fig.height=6}
## heat map is produced to visualize the top marker genes across all the pseudo-bulk samples
lcpm <- edgeR::cpm(y, log=TRUE)
annot <- data.frame(cluster=paste0("cluster ", cluster))
rownames(annot) <- colnames(y)
ann_colors <- list(cluster=2:4)
names(ann_colors$cluster) <- paste0("cluster ", levels(cluster))
pheatmap::pheatmap(lcpm[topMarkers, ], breaks=seq(-2,2,length.out=101),
 color=colorRampPalette(c("blue","white","red"))(100), scale="row",
 cluster_cols=TRUE, border_color="NA", fontsize_row=5,
 treeheight_row=70, treeheight_col=70, cutree_cols=7,
 clustering_method="ward.D2", show_colnames=FALSE,
 annotation_col=annot, annotation_colors=ann_colors)
```


################## END  edgeR DE ################## 

###### MSC49 strain DE assessing var genes

```{r, fig.width=7.5, fig.height=2}
m49_gns_exp <- FetchData(object = m2_asex_harmony_tr_pt_sset, vars = c('PF3D7-0400400', "PF3D7-0412400", 'PF3D7-0421300', "PF3D7-0412700", 'PF3D7-0712900', 'PF3D7-1041300', "PF3D7-0632800", 'PF3D7-0937800', 'PF3D7-1240600', 'PF3D7-1240400', 'PF3D7-0808700', 'PF3D7-0900100'), slot = "data") 

```


```{r, fig.width=7.5, fig.height=2}
m49_umap_gns <- m2_asex_harmony_tr_pt_sset@reductions$umap.harmony_PC_man@cell.embeddings %>% as.data.frame() %>% rownames_to_column("bcode") %>% left_join(., m2_asex_harmony_tr_pt_sset@meta.data %>% rownames_to_column("bcode")) %>%  left_join(., m49_gns_exp %>% rownames_to_column("bcode"))

```

```{r, fig.width=18, fig.height=5}
# custom_colors <- colorRampPalette(c("lightgrey", "blue"))(100)

m49_umap_gns %>% 
  filter(donor == "MSC49C") %>%
  pivot_longer(cols = starts_with("PF"), names_to = "Vars", values_to = "Expression") %>%
  mutate(across(donor, ~str_remove_all(., "SC|C")),
         # across(strain, ~factor(.x, levels = c("SC1", "SC2")))
         ) %>%
  arrange(Vars,Strain_DN,Expression) %>%
  # ggplot(., aes(x = weave_factors(Strain, donor), y = Counts, fill = donor)) +
  ggplot(., aes(x = umapharmonyPCman_1, y = umapharmonyPCman_2, colour = Expression))+
  geom_point(size = 0.3) +
  scale_color_distiller( direction = 1) +
  ggh4x::facet_nested_wrap(vars(Vars, Strain_DN),  nrow = 1, scales = "free_x") +
  theme_classic()+
  theme(legend.position = "bottom", 
        axis.text = element_blank(),
        strip.text =element_text(size=12,face="bold", colour = 'black'))+
  # guides(x = "axis_nested") +
  labs(x = "UMAP_1", y="UMAP_2", title = "Var genes") 
```

###### END MSC49 strain DE assessing var genes

```{r, results="asis", fig.width= 20, fig.height=2.8}
FetchData(object = m2_asex_harmony_tr_pt_sset, vars = 'PF3D7-1240400', slot = "data") 

FeaturePlot(m2_asex_harmony_tr_pt_sset, reduction = "umap.harmony_PC_man", features = "PF3D7-0114500", split.by = c("donor"), order = T)+
  # scale_color_manual(values = sp_fld_colors)+
  theme(axis.text=element_blank(), 
        axis.title=element_blank(),
        legend.position = "bottom",
        # legend.position = "none",
        legend.text = element_text(size = 10) ,
        legend.title = element_text(size = 10, face = "bold"),
        plot.title = element_blank(),
        strip.text =element_text(size=10,face="bold", colour = 'black')) +
    guides(color = guide_legend(title.position = "top", nrow =1, override.aes = list(size = 3))) +
  labs(x = "UMAP_1", y="UMAP_2") 
```

Feature plots of HbC and HbS upregulated genes stratified by strains only showing strain with more than 20 cells
```{r, fig.width=27, fig.height=3}
# FeaturePlot(m2_asex_harmony_tr_pt_sset[,m2_asex_harmony_tr_pt_sset@meta.data %>% filter(donor %in% c("MSC14", "MSC66", "MSC51")) %>% rownames_to_column('bcode') %>% group_by(StrainDon) %>% filter(n()>20) %>% pull(bcode)], features = "PF3D7-1229200", split.by = c("StrainDon"), ncol = 5, order = T)
FeaturePlot(m2_asex_harmony_tr_pt_sset[,m2_asex_harmony_tr_pt_sset@meta.data %>% filter(donor %in% c("MSC14", "MSC66", "MSC51")) %>% rownames_to_column('bcode') %>% group_by(StrainDon) %>% filter(n()>20) %>% pull(bcode)], features = "PF3D7-1229300", split.by = c("StrainDon"), ncol = 5, order = T)

```

Feature plots of HbC and HbS upregulated genes stratified by strains showing all strains 
```{r, fig.width=15, fig.height=3}
# DotPlot(m2_asex_harmony_tr_pt_sset, features = c("PF3D7-1229200", "PF3D7-1229300"), group.by = c("StrainDon")) + coord_flip() + theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1), legend.position = "bottom")
DotPlot(m2_asex_harmony_tr_pt_sset, features = c( "PF3D7-1229300"), group.by = c("StrainDon")) + coord_flip() + theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1), legend.position = "bottom")
```


```{r, fig.width=14, fig.height=3}
DimPlot(m2_asex_harmony_tr_pt_sset,  reduction = "umap.harmony_PC_man", split.by = c("donor"), order = T, group.by = "condition") + theme(legend.position = "bottom", axis.text = element_blank(), axis.ticks = element_blank())
```



```{r, fig.width=15, fig.height=5}
# visualize by batch and cell type annotation
# m2_asex_harmony_tr_pt_sset <- v3_m2_asex_harmony_ss_res_fld_pt_pf[,v3_m2_asex_harmony_ss_res_fld_pt_pf@meta.data$dset == "Field"]
```




```{r}
v3_m2_asex_harmony_ss_res_fld_pt_pf[,v3_m2_asex_harmony_ss_res_fld_pt_pf@meta.data$dset == "Field"]@meta.data %>% count(harmony_clusters_ss)
```



```{r}
fld_clusters <- v3_m2_asex_harmony_ss_res_fld_pt_pf[,v3_m2_asex_harmony_ss_res_fld_pt_pf@meta.data$dset == "Field"]@meta.data %>% count(harmony_clusters_ss) %>% pull(harmony_clusters_ss) 

fld_clusters_cutoff <- v3_m2_asex_harmony_ss_res_fld_pt_pf[,v3_m2_asex_harmony_ss_res_fld_pt_pf@meta.data$dset == "Field"]@meta.data %>% count(harmony_clusters_ss) %>% filter(n>200) %>% pull(harmony_clusters_ss) 
```

```{r}
## Retained
DimPlot(v3_m2_asex_harmony_ss_res_fld_pt_pf[,v3_m2_asex_harmony_ss_res_fld_pt_pf@meta.data$harmony_clusters_ss %in% fld_clusters_cutoff], reduction = "umap.harmony_PC_man", split.by = "dset", label = T)
```

```{r}
## Removed
# DimPlot(v3_m2_asex_harmony_ss_res_fld_pt_pf[,!(v3_m2_asex_harmony_ss_res_fld_pt_pf@meta.data$harmony_clusters_ss %in% fld_clusters_cutoff)], reduction = "umap.harmony_PC_man", split.by = "dset", label = T)
```


```{r}
## subset where field parasites are for slingshot
# v3_m2_asex_harmony_ss_res_fld_pt_pf <- v3_m2_asex_harmony_ss_res_fld_pt_pf[,v3_m2_asex_harmony_ss_res_fld_pt_pf@meta.data$harmony_clusters_ss %in% fld_clusters_cutoff]
```


```{r, fig.width=8, fig.height=5}
DimPlot(v3_m2_asex_harmony_ss_res_fld_pt_pf,  reduction = "umap.harmony_PC_man", split.by = c("condition"))
```

## Trajectory inference of integrated object
Adjust to get appropriate clustering so as to input into slingshot
```{r}
## Adjusting resolution to find the optimum for slingshot trajectory inference to avoid error  "values must be length 150, but FUN(X[[1]]) result is length 38689"

DimPlot(v3_m2_asex_harmony_ss_res_fld_pt_pf, reduction = "umap.harmony_PC_man", label = T)
# msc_gam_combi_noLE <- FindClusters(msc_gam_combi_noLE, resolution = 0.32, verbose = F)

```

```{r}
## Adjusting resolution to find the optimum for slingshot trajectory inference to avoid error  "values must be length 150, but FUN(X[[1]]) result is length 38689"
# DimPlot(v3_m2_asex_harmony_ss_res_fld_pt_pf, reduction = "umap.harmony_PC_man", group.by = "harmony_clusters_ss", pt.size = 0.5, dims = c(1,2), label = T)

```

!! Plotting
```{r, results="asis", fig.width= 7, fig.height=5}

# umap_meta_df <- m2_asex_harmony_tr_pt_sset@reductions$umap.harmony_PC_man@cell.embeddings %>% as.data.frame() %>%rownames_to_column("cell_bc") %>% left_join(., m2_asex_harmony_tr_pt_sset@meta.data[,c("stage", "StageFL", "pseudotime", "pseudotime_fld_only", "orig.ident", "donor", "StrainDon", "condition", "dset", "harmony_clusters", "harmony_clusters_ss")] %>% as.data.frame() %>% rownames_to_column("cell_bc"), by = "cell_bc") %>% mutate(across(donor, ~str_remove_all(., "SC|C"))) %>% left_join(., m2_asex_harmony_tr_pt_sset@reductions$integrated.harmony@cell.embeddings %>% as.data.frame() %>%rownames_to_column("cell_bc")) #%>% 

umap_meta_df <- m2_asex_harmony_tr_pt_sset@reductions$umap.harmony_PC_man@cell.embeddings %>% as.data.frame() %>%rownames_to_column("cell_bc") %>% left_join(., m2_asex_harmony_tr_pt_sset@meta.data[,c("stage", "StageFL", "pseudotime", "pseudotime_fld_only", "orig.ident", "donor", "StrainDon", "condition", "dset", "harmony_PC_man_clusters", "harmony_clusters_ss")] %>% as.data.frame() %>% rownames_to_column("cell_bc"), by = "cell_bc") %>% mutate(across(donor, ~str_remove_all(., "SC|C"))) %>% left_join(., m2_asex_harmony_tr_pt_sset@reductions$integrated.harmony@cell.embeddings %>% as.data.frame() %>%rownames_to_column("cell_bc")) #%>% 
# mutate(across(strain_dset, ~factor(., levels = rev(strain_dset_lvls_don)))) %>% arrange(strain_dset

```

```{r, results="asis", fig.width= 1.5, fig.height=3}

map(c("pseudotime", "pseudotime_fld_only"), ~{
  pt = .x
  map(c("harmony_", "umapharmonyPCman_"), ~{
  umap_meta_df %>% 
  ggplot(aes(x = !!rlang::sym(paste0(.x, "1")), y =!!rlang::sym(paste0(.x, "2")), color = !!rlang::sym(pt))) + geom_point(size=0.5) + 
  scale_color_distiller(name="Pseudotime", direction = 1, palette = "Blues") + 
  theme_classic()+
  theme(axis.text=element_blank(), 
        axis.title=element_blank(),
        axis.ticks = element_blank(),
        legend.position = "bottom",
        legend.text = element_text(size = 15) ,
        legend.title = element_text(size = 15, face = "bold") ) +
    guides(color = guide_colourbar(title.position = "top"))
  })
})


```

```{r, results="asis", fig.width= 5.5, fig.height=4.8}

thm_std <- theme(axis.text=element_text(size=26),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), 
        axis.title=element_text(size=28,face="bold"), 
        legend.text = element_text(size = 26) ,
        legend.title = element_text(size = 28, face = "bold"),
        # legend.box.margin = margin(t = 0, r = 4.1, b = 0, l = 0, unit = "cm"),#######
        legend.position="bottom",
        # legend.direction="vertical"
  )

gbox_theme<- theme(axis.text=element_text(size=15),
        axis.title=element_text(size=15,face="bold"), 
        legend.text = element_text(size = 15) ,
        legend.title = element_text(size = 15, face = "bold"),
        legend.position = "bottom")
        
dens_thme <- theme(axis.text=element_text(size=22),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), 
        axis.title=element_text(size=22,face="bold"), 
        legend.text = element_text(size = 20) ,
        legend.title = element_text(size = 20, face = "bold"),
        legend.position = "bottom")
  
map(c("pseudotime", "pseudotime_fld_only"), ~ umap_meta_df %>% 
  filter(dset == "Field") %>%
  ggplot(., aes(x = !!rlang::sym(.x), col=StageFL, fill=stage)) + 
  scale_color_manual(name = "Stage", values = sp_fld_colors)+ 
  scale_fill_manual(name = "Stage", values = sp_fld_colors)+
  geom_density(aes(y = ..count..*0.18), size=2, bw = 0.18, alpha=.2)+
  # geom_freqpoly()+
  theme_classic()+
  thm_std+
  labs(x = "Pseudotime", y="Counts") +
  guides(color = guide_legend(override.aes = list(size = 4), title.position = "top"), fill = 'none')
)

map(c("pseudotime", "pseudotime_fld_only"), ~umap_meta_df %>% 
  filter(dset == "Field") %>%
  ggplot(., aes(x = !!rlang::sym(.x), col=donor, fill=donor)) + 
  geom_density(aes(y = ..count..*0.18), size=2, bw = 0.18, alpha=.2)+
  # geom_freqpoly()+
  theme_classic()+
  thm_std+
  labs(x = "Pseudotime", y="Counts") +
  guides(color = guide_legend(override.aes = list(size = 4), title.position = "top", nrow = 4))
)


symp_dens_pt_plt <- map(c("pseudotime", "pseudotime_fld_only"), ~umap_meta_df %>% 
  filter(dset == "Field") %>%
  ggplot(., aes(x = !!rlang::sym(.x), col=condition, fill=condition)) + 
  geom_density(aes(y = ..count..*0.18), size=1.5, bw = 0.18, alpha=.2)+
  # geom_freqpoly()+
  theme_classic()+
  dens_thme+
  labs(x = "Pseudotime", y="Counts") +
  guides(color = guide_legend(override.aes = list(size = 2), title.position = "top", title = "Status"),
         fill = "none")
)

symp_dens_pt_plt

symp_box_pt_plt <-map(c("pseudotime", "pseudotime_fld_only"), ~umap_meta_df %>% 
  filter(dset == "Field") %>%
  ggpubr::ggboxplot(., y =  .x, x="condition", col="condition", size=1,) + 
  ggpubr::stat_compare_means(aes(label = paste0("P = ", after_stat(p.format))), size = 6, method = "wilcox.test", label.x = 1,
  label.y = 4) +
    gbox_theme+
    theme(axis.text.x = element_blank(), 
        axis.title=element_text(size=22,face="bold")#,
        # legend.position = "bottom"
        )+
  labs(x = "Status", y="Pseudotime", color="Strain")
)
  
symp_box_pt_plt

```

```{r, fig.width=7.5, fig.height=4}

str_lege <- cowplot::get_plot_component(symp_dens_pt_plt[[2]], 'guide-box-bottom', return_all = TRUE) 

plot_grid(
  plot_grid(symp_dens_pt_plt[[1]]+ theme(legend.position = "none") , 
            symp_box_pt_plt[[1]] + theme(legend.position = "none", axis.title.x = element_blank()), 
            nrow = 1,
            rel_widths = c(6.0, 4.0)),
  str_lege,
  nrow = 2,
  rel_heights = c(8,2)
)
```


######### START CONDIMENTS PACKAGE ###########

```{r, fig.width=7.5, fig.height=4}

ggplot(umap_meta_df, aes(x = harmony_1, y =harmony_2, col = condition)) +
  geom_point() +
  scale_color_brewer(type = "qual")

ggplot(umap_meta_df, aes(x = umapharmonyPCman_1, y =umapharmonyPCman_2, col = condition)) +
  geom_point() +
  scale_color_brewer(type = "qual")


```


```{r}

data("toy_dataset", package = "condiments")
df <- toy_dataset$sd
toy_dataset %>% glimpse()
```

```{r, fig.width=7.5, fig.height=4}
p <- ggplot(umap_meta_df, aes(x = umapharmonyPCman_1, y =umapharmonyPCman_2, col = condition)) +
  geom_point(alpha = .5) +
  # geom_point(data = toy_dataset$mst, size = 2) +
  # geom_path(data = toy_dataset$mst, aes(group = lineages), size = 1.5) +
  scale_color_brewer(type = "qual") + 
  facet_wrap(~condition) +
  guides(col = FALSE)
p

```




```{r}

scores <- imbalance_score(Object = umap_meta_df %>% select(umapharmonyPCman_1, umapharmonyPCman_2) %>% as.matrix(),
                          conditions = umap_meta_df$condition)

umap_meta_df$scores <- scores$scores
umap_meta_df$scaled_scores <- scores$scaled_scores
```


```{r}
ggplot(umap_meta_df, aes(x = umapharmonyPCman_1, y = umapharmonyPCman_2, col = scores)) +
  geom_point() +
  scale_color_viridis_c(option = "C")

```


```{r, fig.width=7.5, fig.height=4}
ggplot(umap_meta_df, aes(x = umapharmonyPCman_1, y = umapharmonyPCman_2, col = scaled_scores)) +
  geom_point() +
  scale_color_viridis_c(option = "C")
```


```{r, fig.width=7.5, fig.height=4}
ggplot(umap_meta_df, aes(x = umapharmonyPCman_1, y = umapharmonyPCman_2, col = harmony_clusters_ss)) +
  geom_point() 
```

```{r}

rd <- as.matrix(umap_meta_df[, c("umapharmonyPCman_1", "umapharmonyPCman_2")])
# sds <- slingshot(rd, umap_meta_df$harmony_PC_man_clusters)
# sds <- slingshot(rd, umap_meta_df$harmony_clusters_ss, end.clus = "3")
sds <- slingshot(rd, umap_meta_df$harmony_clusters_ss, end.clus = "2")


```

```{r, eval=F}

## Takes ~1m30s to run
top_res <- topologyTest(sds = sds, conditions = df$condition)
```

```{r, fig.width=7.5, fig.height=4, eval=F}
ggplot(umap_meta_df, aes(x = umapharmonyPCman_1, y = umapharmonyPCman_2, col = harmony_clusters_ss)) +
  geom_point()
top_res <- topologyTest(sds = sds, conditions = df$conditions)
```


```{r, fig.width=7.5, fig.height=4}
ggplot(umap_meta_df, aes(x = umapharmonyPCman_1, y = umapharmonyPCman_2, col = harmony_clusters_ss)) +
  geom_point() +
  geom_path(data =  slingCurves(sds, as.df = TRUE) %>% arrange(Order),
            aes(group = Lineage), col = "black", size = 2)
```


```{r}
psts <- slingPseudotime(sds) %>%
  as.data.frame() %>%
  mutate(cells = rownames(.),
         conditions = umap_meta_df$condition) %>%
  pivot_longer(starts_with("Lineage"), values_to = "pseudotime", names_to = "lineages")

```


```{r}
ggplot(psts, aes(x = pseudotime, fill = conditions)) +
  geom_density(alpha = .5) +
  scale_fill_brewer(type = "qual") +
  facet_wrap(~lineages) +
  theme(legend.position = "bottom")

```


```{r, fig.width=4, fig.height=8}
psts$donor <- umap_meta_df$donor

ggplot(psts, aes(x = pseudotime, fill = conditions)) +
  geom_density(alpha = .5) +
  scale_fill_brewer(type = "qual") +
  facet_nested(vars(conditions, donor)) +
  theme(legend.position = "bottom")

```

```{r}
prog_res <- progressionTest(sds, conditions = umap_meta_df$condition, global = TRUE, lineages = TRUE)
knitr::kable(prog_res)

```


```{r}
# # library(lme4)
# library(lmerTest)
# # Fit a mixed model to pseudotime values
# model <- lmerTest::lmer(pseudotime ~ conditions + (1|donor), data = psts)
# summary(model)


```


```{r}
# Fit a mixed model to pseudotime values
model2 <- lmerTest::lmer(pseudotime ~ condition + (1|donor), data = umap_meta_df)
# summary(model2)

# Fit a reduced model without the fixed effect
model_reduced <- lmerTest::lmer(pseudotime ~ 1 + (1 | donor), data = umap_meta_df)

# Perform a likelihood ratio test
anova(model_reduced, model2)
```


```{r}
saveRDS(umap_meta_df, "data/processed/Pf/m2_all/umap_meta_df.RDS")
```

```{r}


```

```{r, fig.width=7.5, fig.height=4}
gogo()

```


######### END CONDIMENTS PACKAGE ###########



## Placing the more sophisticated DE at the top to save on time since MAST findmarkers run takes long

## DREAMLET
## Proper DE accounting for factors
```{r, fig.width=15, fig.height=5}
# visualize by batch and cell type annotation
v3_m2_asex_harmony_ss_res_fld_pt_pf@meta.data %>% filter(!str_detect(StrainDon, "Doublet|Negative") & StageFL %in% c("Early trophozoite", "Late ring"))  %>% group_by(StrainDon,StageFL) %>% filter(n() > 10) %>% dplyr::count(StrainDon, StageFL)
```


```{r, fig.width=15, fig.height=5}
# visualize by batch and cell type annotation
v3_m2_asex_harmony_ss_res_fld_pt_pf@meta.data %>% filter(!str_detect(StrainDon, "Doublet|Negative") & StageFL %in% c("Early trophozoite", "Late ring"))  %>% group_by(StrainDon, StageFL) %>% filter(n() > 10) %>% ungroup() %>% add_count(donor, name = "don_count") %>% add_count(StrainDon, name = "strn_count") %>% mutate(strn_prop = strn_count/don_count) %>% dplyr::count(StrainDon, StageFL,don_count,strn_count,strn_prop)
```


Check pseudotime by sampling time
```{r, fig.width=15, fig.height=5}
map(c("smpl_wait_time","parasitaemia_pf"), ~{
  
  m2_asex_harmony_tr_pt_sset@meta.data %>% 
    ggplot(aes(x = !!rlang::sym(.x), y = pseudotime, colour = condition)) +
    geom_point() 
})
```

Check pseudotime by sampling time
```{r, fig.width=15, fig.height=5}
map(c("smpl_wait_time","parasitaemia_pf"), ~{
  
  m2_asex_harmony_tr_pt_sset@meta.data %>% 
    ggplot(aes(x = !!rlang::sym(.x), y = mean(pseudotime), colour = condition)) +
    geom_point() +
  geom_smooth(method = "lm")
  
})
```

```{r, fig.width=15, fig.height=5}
m2_asex_harmony_tr_pt_sset@meta.data %>% 
  lm(formula =pseudotime ~ smpl_wait_time + parasitaemia_pf, data = .)
```

```{r}
st_dist <- m2_asex_harmony_tr_pt_sset@meta.data %>% count(donor, condition, StrainDon) %>% filter(n>20) %>% group_by(donor) 

```

```{r}
st_dist %>%  ggplot(aes(x = donor, fill = StrainDon, y = n)) + geom_col(position = position_dodge2()) + facet_wrap(condition~., scales = "free_x")
```


```{r, fig.width= 10, fig.height= 4}
m2_asex_harmony_tr_pt_sset@meta.data %>% count(donor, condition) %>% filter(n>20) %>% ggplot(aes(x = donor, fill = donor, y = n)) + geom_col(position = position_dodge2()) + facet_wrap(condition~., scales = "free_x")
```


```{r, fig.width= 6, fig.height= 2.5}
m2_asex_harmony_tr_pt_sset@meta.data %>% count(donor, condition, StrainDon) %>% filter(n>20) %>% ggplot(aes(x = donor, fill = StrainDon, y = n)) + geom_col(position = position_dodge2()) + facet_wrap(condition~., scales = "free_x") + theme(legend.position = "none", axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))
```


```{r, fig.width= 10, fig.height= 4}
m2_asex_harmony_tr_pt_sset@meta.data %>% count(donor, condition, StrainDon) %>% filter(n>20) %>% ggplot(aes(x = donor, fill = StrainDon, y = n)) + geom_col(position = position_dodge2()) + facet_wrap(condition~., scales = "free_x") + theme(legend.position = "right") + coord_flip()
```

```{r}
## Only those with asexuals and included in analysis
map(c("parasitaemia_pf", "age_years", "smpl_wait_time"), ~{
  donor_mdata_slct %>% 
  filter(short_name %in% str_remove(unique(st_dist$donor), "_SLO")) %>%
  ggplot(aes(x = symptomatic, y = !!rlang::sym(.x))) +
  geom_boxplot() +
  geom_point()
})
```



Subset to retain only field asexual strains with great than 20 strains per straindon group. Remove stages with few cells - early rings and schizonts which are a minority and maybe poorly labelled
```{r, results="asis"}
qcd_asex <- m2_asex_harmony_tr_pt_sset@meta.data %>% rownames_to_column("bcode") %>% add_count(donor, condition, StrainDon) %>% filter(n>20) %>% pull(bcode)
  
```

```{r, fig.width=15, fig.height=5, eval=FALSE}
# visualize by batch and cell type annotation
m2_asex_harmony_tr_pt_sset_strn20 <- m2_asex_harmony_tr_pt_sset[,qcd_asex]
m2_asex_tr_sce <- as.SingleCellExperiment(m2_asex_harmony_tr_pt_sset_strn20)
```

```{r, fig.width=15, fig.height=5}
m2_asex_tr_sce <- as.SingleCellExperiment(m2_asex_harmony_tr_pt_sset)
```

## check the strains going into the analysis

```{r, results="asis"}
# only keep cells with sufficient reads
m2_asex_tr_sce <- m2_asex_tr_sce[rowSums(counts(m2_asex_tr_sce) > 0) > 0, ]

## 

# compute QC metrics
m2_asex_qc <- perCellQCMetrics(m2_asex_tr_sce)

# remove cells with few or many detected genes
ol <- isOutlier(metric = m2_asex_qc$detected, nmads = 2, log = TRUE)

table(ol)

```


```{r, results="asis"}
# only keep cells with sufficient reads
m2_asex_tr_sce <- m2_asex_tr_sce[, !ol]
plotReducedDim(m2_asex_tr_sce, dimred = "UMAP.HARMONY_PC_MAN")
```

###################### PSEUDOTIME CHUNKING START  ############################################

## pseudotime chunks used for all the DE methods below and can be used for DEseq2 above as well
### Calculate pseudotime chunks
```{r, eval=F}
##Generate pseudotime chunks categories with range of highest and lowest pseudotime values and breaks numbers to ensure balanced representation of cells and drop the cells without strain information
cat_n<-3

##Remove those without pseudotime
m2_asex_tr_sce_pt_chunks <- m2_asex_tr_sce[,!is.na(m2_asex_tr_sce@colData[,'pseudotime'])]

      # print(sce_obj)
m2_asex_tr_sce_pt_chunks@colData[,c('PT_grp_unbal', 'PT_grp')] <- m2_asex_tr_sce_pt_chunks@colData[,c( "pseudotime"), drop=F] %>%
        data.frame() %>% 
        mutate(PT_grp_unbal = cut(!!rlang::sym("pseudotime"),  
                               breaks = c(-Inf, seq(min(!!rlang::sym("pseudotime")),  max(!!rlang::sym("pseudotime")), round((max(!!rlang::sym("pseudotime")) - min(!!rlang::sym("pseudotime")))/cat_n, 2)), Inf)),
               pt_rank = dense_rank(!!rlang::sym("pseudotime")),
               PT_grp = cut(pt_rank,  
                               breaks = c(-Inf, seq(min(pt_rank),  max(pt_rank), round((max(pt_rank) - min(pt_rank))/cat_n, 2)), Inf))
               ) %>%
  mutate(ptime_grp_rnk = dense_rank(factor(PT_grp_unbal)),
         ptime_bal_grp_rnk = dense_rank(factor(PT_grp)),
         PT_grp_unbal = paste0("PT.", ptime_grp_rnk),
         PT_grp = paste0("PT.", ptime_bal_grp_rnk)
         ) %>%
        dplyr::select(PT_grp_unbal, PT_grp)

```


```{r, eval=FALSE}
plotReducedDim(m2_asex_tr_sce_pt_chunks, dimred = "UMAP.HARMONY_PC_MAN", colour_by = "PT_grp")
```


```{r, eval=FALSE}
m2_asex_tr_sce_pt_chunks@colData[,c("PT_grp", "donor")] %>% as.data.frame() %>% dplyr::count(PT_grp)
```

```{r, eval=FALSE}
plotReducedDim(m2_asex_tr_sce_pt_chunks, dimred = "UMAP.HARMONY_PC_MAN", colour_by = "PT_grp_unbal")
```

```{r, eval=FALSE}
m2_asex_tr_sce_pt_chunks@colData[,c("PT_grp_unbal", "donor")] %>% as.data.frame() %>% dplyr::count(PT_grp_unbal)
```

```{r, eval=FALSE}
m2_asex_tr_sce_pt_chunks_sset <- m2_asex_tr_sce_pt_chunks[,m2_asex_tr_sce_pt_chunks$PT_grp_unbal %in% c("PT.3", "PT.4", "PT.5", "PT.6", "PT.7")]
m2_asex_tr_sce_pt_chunks_sset <- m2_asex_tr_sce_pt_chunks[,m2_asex_tr_sce_pt_chunks$PT_grp %in% c("PT.2", "PT.3", "PT.4", "PT.5", "PT.7")]
```

```{r, eval=FALSE}
m2_asex_tr_sce_pt_chunks_sset@colData[,c("PT_grp", "donor")] %>% as.data.frame() %>% dplyr::count(PT_grp)
```

```{r}
# ##Generate pseudotime chunks categories with range of highest and lowest pseudotime values and breaks numbers to ensure balanced representation of cells and drop the cells without strain information
# cut_pt_denom <- rep(c(6,6,6,6), 2)
# # pts <- rep(c("female_pt", "female_pt", "male_pt", "male_pt" ),4)
# pts <- rep(c("female_pt", "male_pt" ),4)
# 
# 
# msc_gcombi_noLE_sub_mf_sce_unbal_auto_ci_ptchunks <- list(msc_gcombi_noLE_sub_mf_sce_unbal_auto_ci, pts, cut_pt_denom) %>%
#   pmap(
#     # possibly(
#     ~{
#       sce_obj <- ..1[,!is.na(..1@colData[,'dset'])]
#       # print(sce_obj)
#       sce_obj@colData[,'PT_grp_unbal'] <- sce_obj@colData[,c( ..2), drop=F] %>%
#         data.frame() %>% 
#         mutate(PT_grp_unbal = cut(!!rlang::sym(..2),  breaks = c(-Inf, seq(min(!!rlang::sym(..2)),  max(!!rlang::sym(..2)), round((max(!!rlang::sym(..2)) - min(!!rlang::sym(..2)))/..3, 2)), Inf))) %>%
#         dplyr::select(PT_grp_unbal)
#       
#       # print(glimpse(a))
#       return(sce_obj)
#     }
#     # , 'e')
#   )
# 
# ##Retain only for MSC14 as we dont want to compare field vs lab for the other samples
# # msc_gcombi_noLE_sub_mf_sce_unbal_auto_ci_ptchunks <- msc_gcombi_noLE_sub_mf_sce_unbal_auto_ci_ptchunks[str_detect(names(msc_gcombi_noLE_sub_mf_sce_unbal_auto_ci_ptchunks), "msc14")]
# msc_gcombi_noLE_sub_mf_sce_unbal_auto_ci_ptchunks

```

### Calculate pseudotime chunks donor strains
```{r}
# ##Generate pseudotime chunks categories with range of highest and lowest pseudotime values and breaks numbers to ensure balanced representation of cells and drop the cells without strain information
# cut_pt_denom <- rep(c(6,6), 9)
# pts <- rep(c("female_pt",  "male_pt" ), 9)
# 
# 
# msc_gcombi_noLE_sub_sce_unbal_ptchunks <- list(msc_gcombi_noLE_sub_mf_sce_unbal_auto_ci_strainsAll, pts, cut_pt_denom) %>%
#   pmap(
#     # possibly(
#     ~{
#       sce_obj <- ..1[,!is.na(..1@colData[,'lf_strain'])]
#       # print(sce_obj)
#       sce_obj@colData[,'PT_grp_unbal'] <- sce_obj@colData[,c( ..2), drop=F] %>%
#         data.frame() %>% 
#         mutate(PT_grp_unbal = cut(!!rlang::sym(..2),  breaks = c(-Inf, seq(min(!!rlang::sym(..2)),  max(!!rlang::sym(..2)), round((max(!!rlang::sym(..2)) - min(!!rlang::sym(..2)))/..3, 2)), Inf))) %>%
#         dplyr::select(PT_grp_unbal)
#       
#       # print(glimpse(a))
#       return(sce_obj)
#     }
#     # , 'e')
#   )
#   

```


```{r}
m2_asex_tr_sce_pt_chunks <- m2_asex_tr_sce
```


###################### PSEUDOTIME CHUNKING END  ############################################

###################### DREAMLET START   ############################################

## DREAMLET & MUSCAT  - DONOR WITH RANDOM EFFECTS

Tabulate and get the pseudotime chunks with most numbers 3,4,5
```{r, results="asis", eval=T}
m2_asex_tr_sce_pt_chunks@colData[, c("PT_grp", "condition", "donor")] %>% data.frame() %>% count(PT_grp, condition)
```

```{r, results="asis", eval=T}
# m2_asex_tr_sce_pt_chunks_345 <- m2_asex_tr_sce_pt_chunks[,m2_asex_tr_sce_pt_chunks@colData$PT_grp %in% c("PT.1", "PT.2", "PT.3")]
m2_asex_tr_sce_pt_chunks_345 <- m2_asex_tr_sce_pt_chunks

```

```{r, results="asis", eval=T}
m2_asex_tr_sce_pt_chunks_345@colData[, c("PT_grp", "condition", "donor", "StrainDon")] %>% data.frame() %>% count(PT_grp, condition, StrainDon)
```

```{r, results="asis"}
# only keep cells with sufficient reads
plotReducedDim(m2_asex_tr_sce_pt_chunks_345, dimred = "UMAP.HARMONY_PC_MAN", color_by = "PT_grp")
```

```{r, results="asis"}
## Save for tradeseq
saveRDS(m2_asex_tr_sce_pt_chunks_345, "data/raw/Pf/m2_asex_tr_sce_pt_chunks_345.RDS")
```

Write out annotated asexual lab+field in .h5ad 
```{r, message=FALSE, warning=FALSE, eval=T}
##Change factor metadata to character as these are not handled properly when converting to h5ad
m2_asex_tr_sce_pt_chunks_345_raw <- as.Seurat(m2_asex_tr_sce_pt_chunks_345)
# m2_asex_tr_sce_pt_chunks_345_raw@assays$RNA@layers$data.1 <- NULL
m2_asex_tr_sce_pt_chunks_345_raw@meta.data <- m2_asex_tr_sce_pt_chunks_345_raw@meta.data %>% mutate(across(where(is.factor), as.character))

## !!!NB - NEEDS SORTING IN FUTURE UPGRADING TO SEURAT 5 RESULTED IN ERROR AS DESCRIBED https://github.com/mojaveazure/seurat-disk/issues/147#top
# m2_asex_tr_sce_pt_chunks_345_raw[["RNA3"]] <- as(object = m2_asex_tr_sce_pt_chunks_345_raw[["RNA"]], Class = "Assay")
# DefaultAssay(m2_asex_tr_sce_pt_chunks_345_raw) <- "RNA3"
# m2_asex_tr_sce_pt_chunks_345_raw[["RNA"]] <- NULL
# m2_asex_tr_sce_pt_chunks_345_raw <- RenameAssays(object = m2_asex_tr_sce_pt_chunks_345_raw, RNA3 = 'RNA')

# SaveH5Seurat(m2_asex_tr_sce_pt_chunks_345_raw, filename =  paste0("/Users/jr35/Google Drive/My Drive/PHD_documents/malaria_phd/",sample_set,"/data/processed/Pf/m2_asex_tr_sce_pt_chunks_345_raw.h5Seurat"), overwrite = T)
SaveH5Seurat(m2_asex_tr_sce_pt_chunks_345_raw, filename =  paste0("data/processed/Pf/m2_all/m2_asex_tr_sce_pt_chunks_345_raw.h5Seurat"), overwrite = T)

# Convert(paste0("/Users/jr35/Google Drive/My Drive/PHD_documents/malaria_phd/",sample_set,"/data/processed/Pf/m2_asex_tr_sce_pt_chunks_345_raw.h5Seurat"), dest = "h5ad", overwrite = T)
Convert(paste0("data/processed/Pf/m2_all/m2_asex_tr_sce_pt_chunks_345_raw.h5Seurat"), dest = "h5ad", overwrite = T)

```

## Dreamlet - pseudobulk by Strain and donor as random intercept effect

```{r, results="asis"}
# Since 'ind' is the individual and 'StimStatus' is the stimulus status,
# create unique identifier for each sample
# m2_asex_tr_sce$id <- paste0(m2_asex_tr_sce$condition, m2_asex_tr_sce$donor, m2_asex_tr_sce$StrainDon)
# m2_asex_tr_sce_pt_chunks_345$id <- paste0( m2_asex_tr_sce_pt_chunks_345$StrainDon)

m2_asex_tr_sce_pt_chunks_345$id <- paste0(m2_asex_tr_sce_pt_chunks_345$condition, m2_asex_tr_sce_pt_chunks_345$StrainDon)
m2_asex_tr_sce_pt_chunks_345$constant_cluster <- "all"


# Create pseudobulk data by specifying cluster_id and sample_id
# Count data for each cell type is then stored in the `assay` field
# assay: entry in assayNames(sce) storing raw counts
# cluster_id: variable in colData(sce) indicating cell clusters
# sample_id: variable in colData(sce) indicating sample id for aggregating cells
m2_asex_tr_pb <- aggregateToPseudoBulk(m2_asex_tr_sce_pt_chunks_345,
  assay = "counts",
  # cluster_id = "StageFL",
  # cluster_id = "constant_cluster",
  cluster_id = "PT_grp",
  sample_id = "id",
  verbose = FALSE
)

# one 'assay' per cell type
assayNames(m2_asex_tr_pb)
```


```{r, results="asis"}
# Since 'ind' is the individual and 'StimStatus' is the stimulus status,
# saveRDS(m2_asex_tr_sce_pt_chunks_345, "/lustre/scratch126/tol/teams/lawniczak/users/jr35/phd/Mali2/data/processed/Pf/m2_all/m2_asex_tr_sce_pt_chunks_345")
```

```{r, results="asis"}
# one 'assay' per cell type
assayNames(m2_asex_tr_pb)

metadata(m2_asex_tr_pb)$aggr_means
```


```{r, results="asis"}

# form <- ~ condition + (1|donor) + fever
# form <- ~ condition + (1|donor) + temperature_c
# form <- ~ condition + (1|donor) 
# form <- ~ condition + (1|donor) + pseudotime + fever + parasitaemia_pf + age_years
# form <- ~ condition + (1|donor) + pseudotime + fever
form <- ~ condition + (1|donor) 

# Normalize and apply voom/voomWithDreamWeights
res.proc <- processAssays(m2_asex_tr_pb, form, min.count = 2)

# the resulting object of class dreamletProcessedData stores
# normalized data and other information
res.proc
```


```{r, results="asis"}
# Normalize and apply voom/voomWithDreamWeights
details(res.proc)
```

```{r, results="asis"}
# show voom plot for each cell clusters
plotVoom(res.proc)
```

```{r, results="asis"}

# run variance partitioning analysis
vp.lst <- fitVarPart(res.proc, form)
```

```{r, results="asis", eval=FALSE}
# Show variance fractions at the gene-level for each cell type
genes <- vp.lst$gene[1:2]
plotPercentBars(vp.lst[vp.lst$gene %in% genes, ])
```


```{r, results="asis", fig.width=10, fig.height=10, eval=FALSE}
# Summarize variance fractions genome-wide for each cell type
plotVarPart(vp.lst, label.angle = 60)
```

```{r, results="asis"}
# Differential expression analysis within each assay,
# evaluated on the voom normalized data
# res.dl <- dreamlet(res.proc, form, method = "random")
res.dl <- dreamlet(res.proc, form)

# names of estimated coefficients
coefNames(res.dl)
```

```{r, results="asis"}
# the resulting object of class dreamletResult
# stores results and other information
res.dl
```

```{r, results="asis"}
plotVolcano(res.dl, coef = "conditionSymptomatic")
```

```{r, results="asis"}
# genes <- c("PF3D7-1240400", "PF3D7-1240900")
# plotGeneHeatmap(res.dl, coef = "conditionSymptomatic", genes = genes)
```

```{r, results="asis"}
conflicts_prefer(dreamlet::topTable)
# results from full analysis
topTable(res.dl, coef = "conditionSymptomatic") 
```

```{r, results="asis"}
# only B cells
topTable(res.dl[["PT.1"]], coef = "conditionSymptomatic") %>% arrange(adj.P.Val)
```

```{r, results="asis"}
plotForest(res.dl, coef = "conditionSymptomatic", gene = c("PF3D7-1240400"))
```

```{r, results="asis"}
genes <- c("PF3D7-1240400", "PF3D7-0609100", "PF3D7-1027200")
plotGeneHeatmap(res.dl, coef = "conditionSymptomatic", genes = genes)
```


```{r, results="asis"}
m2_asex_tr_sce_pt_chunks_345_seu <- as.Seurat(m2_asex_tr_sce_pt_chunks_345) 
```


```{r, results="asis"}
FeaturePlot(m2_asex_tr_sce_pt_chunks_345_seu, features = c("PF3D7-1128200", "PF3D7-1240400"), split.by = "condition") 
```

```{r, results="asis"}
# get data
df <- extractData(res.proc, "PT.2", genes = "PF3D7-1128200")

# set theme
thm <- theme_classic() +
  theme(aspect.ratio = 1, plot.title = element_text(hjust = 0.5))

# make plot
ggplot(df, aes(condition, `PF3D7-1128200`, colour = donor)) +
  geom_boxplot() +
  thm +
  ylab(bquote(Expression ~ (log[2] ~ CPM))) +
  ggtitle("PF3D7-1128200")
```


```{r, results="asis"}
# create a contrasts called 'Diff' that is the difference between expression
# in the stimulated and controls.
# More than one can be specified
contrasts <- c(Diff = "conditionSymptomatic - conditionAsymptomatic")

# Evalaute the regression model without an intercept term.
# Instead estimate the mean expression in stimulated, controls and then
# set Diff to the difference between the two
res.dl2 <- dreamlet(res.proc, ~ 0 + condition, contrasts = contrasts)

# see estimated coefficients
coefNames(res.dl2)
```

```{r, results="asis"}
# Volcano plot of Diff
plotVolcano(res.dl2[1:2], coef = "Diff")
```

```{r, results="asis"}
# results from full analysis
topTable(res.dl, coef = "conditionSymptomatic")

```


```{r, results="asis"}
# results from full analysis
# topTable(res.dl, coef = "conditionAsymptomatic")

```


```{r, results="asis", eval=FALSE}
# results from full analysis
topTable(res.dl, coef = "conditionAsymptomatic")

```

```{r, results="asis", eval=FALSE}
gogo()
```


## Dreamlet - pseudobulk by strain - test fever

```{r, results="asis"}
# Since 'ind' is the individual and 'StimStatus' is the stimulus status,
# create unique identifier for each sample
# m2_asex_tr_sce$id <- paste0(m2_asex_tr_sce$condition, m2_asex_tr_sce$donor, m2_asex_tr_sce$StrainDon)
# m2_asex_tr_sce_pt_chunks_345$symp_don <- paste0(m2_asex_tr_sce_pt_chunks_345$condition, m2_asex_tr_sce_pt_chunks_345$donor)

m2_asex_tr_sce_pt_chunks_345$symp_don <- paste0(m2_asex_tr_sce_pt_chunks_345$fever, m2_asex_tr_sce_pt_chunks_345$StrainDon)


# Create pseudobulk data by specifying cluster_id and sample_id
# Count data for each cell type is then stored in the `assay` field
# assay: entry in assayNames(sce) storing raw counts
# cluster_id: variable in colData(sce) indicating cell clusters
# sample_id: variable in colData(sce) indicating sample id for aggregating cells
m2_asex_tr_pb_don <- aggregateToPseudoBulk(m2_asex_tr_sce_pt_chunks_345,
  assay = "counts",
  # cluster_id = "StageFL",
  cluster_id = "PT_grp",
  sample_id = "symp_don",
  verbose = FALSE
)

# one 'assay' per cell type
assayNames(m2_asex_tr_pb_don)
```


```{r, results="asis"}
# one 'assay' per cell type
assayNames(m2_asex_tr_pb_don)

metadata(m2_asex_tr_pb_don)$aggr_means
```



```{r, results="asis"}

# form <- ~ (1|donor) + condition + pseudotime
# form <- ~ condition + (1|donor)
form <- ~ fever + (1|donor)

# Normalize and apply voom/voomWithDreamWeights
res.proc_don <- processAssays(m2_asex_tr_pb_don, form, min.count = 2)

# the resulting object of class dreamletProcessedData stores
# normalized data and other information
res.proc_don
```


```{r, results="asis"}
# Normalize and apply voom/voomWithDreamWeights
details(res.proc_don)
```

```{r, results="asis"}
# show voom plot for each cell clusters
plotVoom(res.proc_don)
```


```{r, results="asis"}

# run variance partitioning analysis
vp.lst_don <- fitVarPart(res.proc_don, form)
```


```{r, results="asis"}
# Show variance fractions at the gene-level for each cell type
genes <- vp.lst_don$gene[1:2]
# plotPercentBars(vp.lst_don[vp.lst_don$gene %in% genes, ])
```


```{r, results="asis", fig.width=10, fig.height=10}
# Summarize variance fractions genome-wide for each cell type
# plotVarPart(vp.lst_don, label.angle = 60)
```


```{r, results="asis"}
# Differential expression analysis within each assay,
# evaluated on the voom normalized data
# res.dl <- dreamlet(res.proc_don, form, method = "random")
res.dl_don <- dreamlet(res.proc_don, form)

# names of estimated coefficients
coefNames(res.dl_don)
```


```{r, results="asis"}
# the resulting object of class dreamletResult
# stores results and other information
res.dl_don
```

```{r, results="asis"}
plotVolcano(res.dl_don, coef = "feverNoFever")
```


```{r, results="asis"}
conflicts_prefer(dreamlet::topTable)
# results from full analysis
topTable(res.dl_don, coef = "feverNoFever") 
```

## Dreamlet - pseudobulk by donor and donor as random intercept effect

```{r, results="asis"}
# Since 'ind' is the individual and 'StimStatus' is the stimulus status,
# create unique identifier for each sample
# m2_asex_tr_sce$id <- paste0(m2_asex_tr_sce$condition, m2_asex_tr_sce$donor, m2_asex_tr_sce$StrainDon)
# m2_asex_tr_sce_pt_chunks_345$symp_don <- paste0(m2_asex_tr_sce_pt_chunks_345$condition, m2_asex_tr_sce_pt_chunks_345$donor)

m2_asex_tr_sce_pt_chunks_345$symp_don <- paste0(m2_asex_tr_sce_pt_chunks_345$condition, m2_asex_tr_sce_pt_chunks_345$donor)


# Create pseudobulk data by specifying cluster_id and sample_id
# Count data for each cell type is then stored in the `assay` field
# assay: entry in assayNames(sce) storing raw counts
# cluster_id: variable in colData(sce) indicating cell clusters
# sample_id: variable in colData(sce) indicating sample id for aggregating cells
m2_asex_tr_pb_don <- aggregateToPseudoBulk(m2_asex_tr_sce_pt_chunks_345,
  assay = "counts",
  # cluster_id = "StageFL",
  cluster_id = "PT_grp",
  sample_id = "symp_don",
  verbose = FALSE
)

# one 'assay' per cell type
assayNames(m2_asex_tr_pb_don)
```


```{r, results="asis"}
# one 'assay' per cell type
assayNames(m2_asex_tr_pb_don)

metadata(m2_asex_tr_pb_don)$aggr_means
```



```{r, results="asis"}

# form <- ~ (1|donor) + condition + pseudotime
# form <- ~ condition + (1|donor)
form <- ~ condition + parasitaemia_pf

# Normalize and apply voom/voomWithDreamWeights
res.proc_don <- processAssays(m2_asex_tr_pb_don, form, min.count = 2)

# the resulting object of class dreamletProcessedData stores
# normalized data and other information
res.proc_don
```


```{r, results="asis"}
# Normalize and apply voom/voomWithDreamWeights
details(res.proc_don)
```

```{r, results="asis"}
# show voom plot for each cell clusters
plotVoom(res.proc_don)
```


```{r, results="asis"}

# run variance partitioning analysis
vp.lst_don <- fitVarPart(res.proc_don, form)
```


```{r, results="asis"}
# Show variance fractions at the gene-level for each cell type
genes <- vp.lst_don$gene[1:2]
# plotPercentBars(vp.lst_don[vp.lst_don$gene %in% genes, ])
```


```{r, results="asis", fig.width=10, fig.height=10}
# Summarize variance fractions genome-wide for each cell type
# plotVarPart(vp.lst_don, label.angle = 60)
```


```{r, results="asis"}
# Differential expression analysis within each assay,
# evaluated on the voom normalized data
# res.dl <- dreamlet(res.proc_don, form, method = "random")
res.dl_don <- dreamlet(res.proc_don, form)

# names of estimated coefficients
coefNames(res.dl_don)
```


```{r, results="asis"}
# the resulting object of class dreamletResult
# stores results and other information
res.dl_don
```

```{r, results="asis"}
plotVolcano(res.dl_don, coef = "conditionSymptomatic")
```


```{r, results="asis"}
conflicts_prefer(dreamlet::topTable)
# results from full analysis
topTable(res.dl_don, coef = "conditionSymptomatic") 
```


```{r, results="asis"}
# only B cells
topTable(res.dl_don[["PT.2"]], coef = "conditionSymptomatic") %>% arrange(adj.P.Val)
```

```{r, results="asis"}
plotForest(res.dl_don, coef = "conditionSymptomatic", gene = c("PF3D7-1346400"))
```

```{r, results="asis"}
genes <- c("PF3D7-1346400", "PF3D7-1478600", "PF3D7-0205000")
plotGeneHeatmap(res.dl_don, coef = "conditionSymptomatic", genes = genes)
```


```{r, fig.width= 30, fig.height=8}
m2_asex_tr_sce_pt_chunks_345_seu$symp_don <- paste(m2_asex_tr_sce_pt_chunks_345_seu$condition, m2_asex_tr_sce_pt_chunks_345_seu$donor, sep = "_")
FeaturePlot(m2_asex_tr_sce_pt_chunks_345_seu, features = c("PF3D7-1346400", "PF3D7-0205000"), split.by = c("symp_don")) 
```


```{r, results="asis"}
# get data
df <- extractData(res.proc_don, "PT.3", genes = "PF3D7-1346400")

# set theme
thm <- theme_classic() +
  theme(aspect.ratio = 1, plot.title = element_text(hjust = 0.5))

# make plot
ggplot(df, aes(condition, `PF3D7-1346400`, colour = donor)) +
  geom_point() +
  thm +
  ylab(bquote(Expression ~ (log[2] ~ CPM))) +
  ggtitle("PF3D7-1346400")
```

```{r, results="asis"}
# get data
df <- extractData(res.proc_don, "PT.3", genes = "PF3D7-0205000")

# set theme
thm <- theme_classic() +
  theme(aspect.ratio = 1, plot.title = element_text(hjust = 0.5))

# make plot
ggplot(df, aes(condition, `PF3D7-0205000`, colour = donor)) +
  geom_point() +
  thm +
  ylab(bquote(Expression ~ (log[2] ~ CPM))) +
  ggtitle("PF3D7-0205000")
```


```{r, results="asis"}
# create a contrasts called 'Diff' that is the difference between expression
# in the stimulated and controls.
# More than one can be specified
contrasts <- c(Diff = "conditionSymptomatic - conditionAsymptomatic")

# Evalaute the regression model without an intercept term.
# Instead estimate the mean expression in stimulated, controls and then
# set Diff to the difference between the two
res.dl2 <- dreamlet(res.proc_don, ~ 0 + condition, contrasts = contrasts)

# see estimated coefficients
coefNames(res.dl2)
```


```{r, results="asis"}
gogo()

```

##Stop here for the thesis


## Dreamlet following guidance from https://medium.com/@grioni.andrea/mastering-single-cell-rna-seq-with-dreamlet-41ead4650723 

```{r, results="asis"}
# create a contrasts called 'Diff' that is the difference between expression
m2_asex_tr_sce_pt_chunks_345
```



```{r, results="asis"}
# Extract read counts for each Channel/celltype
ctorder = c("PT.1", "PT.2", "PT.3")
df_counts <- map_dfr(ctorder, function(x) {
  data <- assay(pb, x)
  data.frame(
    celltype = x,
    ID = colnames(data),
    readCounts = Matrix::colSums(data)
  )
})

```

```{r, results="asis"}
# Convert celltype to a factor with the specified order
df_counts$celltype <- factor(df_counts$celltype, levels = ctorder)
```


```{r, results="asis"}
# Removing cells with 0 reads.
df_counts <- df_counts[df_counts$readCounts > 0, ]
```


```{r, results="asis"}
# Generate violin plot to visualize the distribution of read counts for each cell type
ggplot(df_counts, aes(x = celltype, y = readCounts, fill = celltype)) +  
  # Violin plot
  geom_violin() + 
  
  # Boxplot inside violin
  geom_boxplot(width = 0.1, outlier.size = 0.1) +
  
  # Theme and aesthetics
  theme_bw() + 
  theme(
    aspect.ratio = 1, 
    plot.title = element_text(hjust = 0.5),
    legend.position = "none",
    axis.title.y = element_blank()
  ) + 
  
  # Log scale for y-axis
  scale_y_log10() +
  
  # Horizontal layout
  coord_flip() +
  
  # Labels and title
  labs(
    y = "Number of reads observed for each CellType",
    x = NULL,
    title = 'Reads per cell cluster for each CellType'
  )
```

Show the normalized reads for cell ratio.
```{r, results="asis"}
# Convert cell counts to dataframe and pivot to longer format
df_rate <- cellCounts(pb) |>
  as_tibble(rownames = "ID") |>
  tidyr::pivot_longer(cols = -ID, values_to = "ncells", names_to = "celltype") |>
  dplyr::mutate(celltype = factor(celltype, ctorder))

# Join dataframes and compute normalized reads
df_joined <- df_counts |>
  dplyr::inner_join(df_rate, by = c("celltype", "ID")) |>
  dplyr::mutate(normalizedReads = readCounts/ncells)
```


```{r, results="asis"}
# Create a violin plot to visualize the normalized read counts for each 'celltype'
df_joined |>
  ggplot(aes(celltype, normalizedReads, fill = celltype)) +  
    # Display distribution of normalized read counts using a violin plot
    geom_violin(color = NA) + 
    # Add a boxplot inside the violin plot for a summarized view
    geom_boxplot(width = 0.1, outlier.size = 0.1) +
    # Use the classic theme for a polished appearance
    theme_classic() + 
    # Adjust theme settings for optimal visualization
    theme(
      aspect.ratio = 1, 
      plot.title = element_text(hjust = 0.5),
      legend.position = "none"
    ) + 
    # Apply a log scale to the y-axis for better data representation
    scale_y_log10() +
    # Present the plot horizontally for clarity
    coord_flip() +
    # Define labels for the plot
    labs(
      y = 'Normalized Reads (Reads per cell)', 
      x = NULL, 
      title = 'Normalized Reads per Cell for Each Channel'
    )
```



```{r, results="asis"}
df <- cellTypeSpecificity( pb )

plotViolin(df, assays=ctorder)
```


```{r, results="asis"}
# Process the 'df' dataframe to extract top marker genes for each cell type
my_markers <- df |> 
  # Convert the dataframe to a tibble and assign row names to 'gene' column
  tibble::as_tibble(rownames = "gene") |>

  # Exclude rows with 'gene' values containing the pattern "ENSG"
  dplyr::filter(!str_detect(gene, "ENSG")) |>

  # Transform the dataframe to a long format, with 'cellType' and 'CPM' columns
  tidyr::pivot_longer(
    cols = !c(gene, totalCPM), 
    names_to = "cellType", 
    values_to = "CPM"
  ) |>

  # Compute the percentage of CPM for each gene relative to 'totalCPM'
  dplyr::mutate(percentCPM = CPM / totalCPM * 100) |>

  # Organize data by 'cellType' and order within each group by 'percentCPM'
  dplyr::group_by(cellType) |>
  dplyr::arrange(desc(percentCPM)) |>

  # Retain only the top 3 genes for each cell type
  dplyr::slice(1:3) |>

  # Keep only the 'cellType' and 'gene' columns
  dplyr::select(cellType, gene) |>

  # Eliminate any repeated rows
  dplyr::distinct() |>

  # Convert the tibble to a named vector with 'cellType' as names and 'gene' as values
  tibble::deframe()
```



```{r, results="asis"}
# Plot percent bars for the specified genes and assays
plotPercentBars(df, genes = unique(my_markers), assays = ctorder)
```



```{r, results="asis"}
dreamlet::plotHeatmap(
  df,
  genes=unique(my_markers),
  assays=ctorder)
```



```{r, results="asis"}
# Create a formula for normalization
form = ~ (1|SubID) + (1|poolID) + Sex + scale(Age) + Dx
```


```{r, results="asis"}

```

###################### DREAMLET END  ############################################

###################### MILO START   ############################################

```{r, results="asis"}
traj_milo <- Milo(m2_asex_tr_sce)
```


```{r, results="asis"}
# traj_milo <- buildGraph(traj_milo, k = 30, d = 30, reduced.dim = "PCA")
traj_milo <- buildGraph(traj_milo, k = 30, d = 30, reduced.dim = "INTEGRATED.HARMONY_PC_MAN")

```

```{r, results="asis"}
# traj_milo <- makeNhoods(traj_milo, prop = 0.1, k = 30, d=30, refined = TRUE, reduced_dims = "PCA")
traj_milo <- makeNhoods(traj_milo, prop = 0.1, k = 30, d=30, refined = TRUE, reduced_dims = "INTEGRATED.HARMONY_PC_MAN")


```

```{r, results="asis"}
plotNhoodSizeHist(traj_milo)
```


```{r, results="asis"}
# traj_milo <- countCells(traj_milo, meta.data = data.frame(colData(traj_milo)), samples="StrainDon")
traj_milo <- countCells(traj_milo, meta.data = data.frame(colData(traj_milo)), sample="donor")
```

```{r, results="asis"}
head(nhoodCounts(traj_milo))
```

```{r, results="asis"}
traj_design <- data.frame(colData(traj_milo))[,c("donor", "condition")]
# traj_design <- data.frame(colData(traj_milo))[,c("StrainDon", "condition", "donor")]
traj_design <- distinct(traj_design)
# rownames(traj_design) <- traj_design$StrainDon
rownames(traj_design) <- traj_design$donor
## Reorder rownames to match columns of nhoodCounts(milo)
# traj_design <- traj_design[colnames(nhoodCounts(traj_milo)), , drop=FALSE]

traj_design

# embryo_design <- data.frame(colData(embryo_milo))[,c("sample", "stage", "sequencing.batch")]
# 
# ## Convert batch info from integer to factor
# embryo_design$sequencing.batch <- as.factor(embryo_design$sequencing.batch) 
# embryo_design <- distinct(embryo_design)
# rownames(embryo_design) <- embryo_design$sample
# 
# embryo_design
```

```{r, results="asis"}
traj_milo <- calcNhoodDistance(traj_milo, d=30)
```

```{r, results="asis"}
da_results <- testNhoods(traj_milo, design = ~ condition, design.df = traj_design)
# da_results <- testNhoods(traj_milo, design = ~ donor + condition, design.df = traj_design)
```

```{r, results="asis"}
da_results %>%
  arrange(SpatialFDR) %>%
  head() 
```


```{r, results="asis"}
ggplot(da_results, aes(PValue)) + geom_histogram(bins=50)
```

```{r, results="asis"}
ggplot(da_results, aes(logFC, -log10(SpatialFDR))) + 
  geom_point() +
  geom_hline(yintercept = 1) ## Mark significance threshold (10% FDR)
```


```{r, results="asis"}
traj_milo <- buildNhoodGraph(traj_milo)

```

```{r, results="asis"}
conflicts_prefer(graphics::layout)
## Plot single-cell UMAP
plotReducedDim(traj_milo, dimred = "UMAP.HARMONY_PC_MAN", colour_by="condition", text_by = "seurat_clusters", 
                          text_size = 3, point_size=0.5)+
  guides(fill="none") + 
  plotNhoodGraphDA(traj_milo, da_results, alpha=0.05, layout = "UMAP.HARMONY_PC_MAN") #+
  # plot_layout(guides="collect")



# ## Plot neighbourhood graph
# nh_graph_pl <- plotNhoodGraphDA(embryo_milo, da_results, layout="umap",alpha=0.1) 
#   
# umap_pl + nh_graph_pl +
#   plot_layout(guides="collect")
```

```{r, results="asis"}
# da_results <- annotateNhoods(traj_milo, da_results, coldata_col = "PT_grp_unbal")
da_results <- annotateNhoods(traj_milo, da_results, coldata_col = "seurat_clusters")
head(da_results)
```


```{r, results="asis"}
ggplot(da_results, aes(seurat_clusters_fraction)) + geom_histogram(bins=50)
# ggplot(da_results, aes(PT_grp_unbal_fraction)) + geom_histogram(bins=50)
```

```{r, results="asis"}
da_results$seurat_clusters <- ifelse(da_results$seurat_clusters_fraction < 0.7, "Mixed", da_results$seurat_clusters)
```


```{r, results="asis"}
plotDAbeeswarm(da_results, group.by = "seurat_clusters")
```

```{r, results="asis"}
## Add log normalized count to Milo object
traj_milo <- logNormCounts(traj_milo)

da_results$NhoodGroup <- as.numeric(da_results$SpatialFDR < 0.1 & da_results$logFC < 0)
# da_nhood_markers <- findNhoodGroupMarkers(traj_milo, da_results, subset.row = rownames(traj_milo)[1:10])

```

```{r, results="asis"}
da_nhood_markers <- findNhoodGroupMarkers(traj_milo, da_results, subset.row = rownames(traj_milo), 
                                          aggregate.samples = TRUE, sample_col = "donor")

```

```{r, results="asis"}
head(da_nhood_markers)
```

Automatic grouping of neighbourhoods

```{r, results="asis"}
## Run buildNhoodGraph to store nhood adjacency matrix
traj_milo <- buildNhoodGraph(traj_milo)

## Find groups
da_results <- groupNhoods(traj_milo, da_results, max.lfc.delta = 2)
head(da_results)
```

```{r, results="asis"}
plotNhoodGroups(traj_milo, da_results, layout="UMAP.HARMONY_PC_MAN") 
```



```{r, results="asis"}
plotDAbeeswarm(da_results, group.by = "NhoodGroup")
```

```{r, results="asis"}
# trial with parameters adjusted
plotDAbeeswarm(groupNhoods(traj_milo, da_results, max.lfc.delta = 1) , group.by = "NhoodGroup") + ggtitle("max LFC delta=1")
plotDAbeeswarm(groupNhoods(traj_milo, da_results, max.lfc.delta = 2)   , group.by = "NhoodGroup") + ggtitle("max LFC delta=2")
plotDAbeeswarm(groupNhoods(traj_milo, da_results, max.lfc.delta = 3)   , group.by = "NhoodGroup") + ggtitle("max LFC delta=3")
```


```{r, results="asis"}
# trial with parameters adjusted
plotDAbeeswarm(groupNhoods(traj_milo, da_results, max.lfc.delta = 5, overlap=1) , group.by = "NhoodGroup") + ggtitle("overlap=5")
plotDAbeeswarm(groupNhoods(traj_milo, da_results, max.lfc.delta = 5, overlap=5)   , group.by = "NhoodGroup") + ggtitle("overlap=10")
plotDAbeeswarm(groupNhoods(traj_milo, da_results, max.lfc.delta = 5, overlap=10)   , group.by = "NhoodGroup") + ggtitle("overlap=20")
```


```{r, results="asis"}
set.seed(42)
da_results <- groupNhoods(traj_milo, da_results, max.lfc.delta = 10, overlap=1)
plotNhoodGroups(traj_milo, da_results, layout="UMAP.HARMONY_PC_MAN")
```


```{r, results="asis"}
plotDAbeeswarm(da_results, group.by = "NhoodGroup")
```



Finding gene signatures for neighbourhoods

```{r, results="asis"}
## Exclude zero counts genes
keep.rows <- rowSums(logcounts(traj_milo)) != 0
traj_milo <- traj_milo[keep.rows, ]

## Find HVGs
dec <- modelGeneVar(traj_milo)
hvgs <- getTopHVGs(dec, n=200)
head(hvgs)
```

one-vs-all differential gene expression for each neighbourhood group
```{r, results="asis"}
nhood_markers <- findNhoodGroupMarkers(traj_milo, da_results, subset.row = hvgs, 
                                       aggregate.samples = TRUE, sample_col = "donor")

head(nhood_markers)
```

```{r, results="asis"}
gr2_markers <- nhood_markers[c("GeneID", "logFC_6", "adj.P.Val_6")] 
colnames(gr2_markers) <- c("GeneID", "logFC", "adj.P.Val")

head(gr2_markers[order(gr2_markers$adj.P.Val), ])
```
subset.groups tests 8-VS-all markers
```{r, results="asis"}
nhood_markers8 <- findNhoodGroupMarkers(traj_milo, da_results, subset.row = hvgs, 
                                       aggregate.samples = TRUE, sample_col = "donor",
                                       subset.groups = c("8")
                                       )

head(nhood_markers8)
```

Compare neighbourhoods - seems like it is each neighbour hood versus all the rest otherwise there should be only one p value
```{r, results="asis"}
nhood_markers2 <- findNhoodGroupMarkers(traj_milo, da_results, subset.row = hvgs,
                                       subset.nhoods = da_results$NhoodGroup %in% c('1','2'),
                                       aggregate.samples = TRUE, sample_col = "donor")

head(nhood_markers2)
```

```{r, results="asis"}
ggplot(nhood_markers2, aes(logFC_2,-log10(adj.P.Val_2 ))) + 
  geom_point(alpha=0.5, size=0.5) +
  geom_hline(yintercept = 2)
```

```{r, results="asis"}
markers <- nhood_markers2$GeneID[nhood_markers$adj.P.Val_2 < 0.3 & nhood_markers2$logFC_2 > 0]
```

```{r, results="asis"}
plotNhoodExpressionGroups(traj_milo, da_results, features=markers[1:20],
                          subset.nhoods = da_results$NhoodGroup %in% c('3','2'), scale=TRUE,
                          grid.space = "fixed")
```

DGE testing within a group
```{r, results="asis"}
dge_9 <- testDiffExp(traj_milo, da_results, design = ~ condition, meta.data = data.frame(colData(traj_milo)), subset.nhoods=da_results$NhoodGroup=="3")
```

```{r, results="asis"}
dge_9
```

Milo GLLM

```{r, results="asis"}
# traj_milo <- Milo(m2_asex_tr_sce_pt_chunks)
```

```{r, results="asis"}
traj_sce <- m2_asex_tr_sce_pt_chunks
```

```{r, results="asis"}
# uncomment the row below to allow multi-processing and comment out the SerialParam line.
# bpparam <- MulticoreParam(workers=4)
bpparam <- SerialParam()
register(bpparam)
```

```{r, results="asis"}
# set.seed(42)
# # remove sparser cells
# drop.cells <- colSums(counts(traj_sce)) < 600
# traj_sce <- computePooledFactors(traj_sce[, !drop.cells], BPPARAM=bpparam)
# traj_sce <- logNormCounts(traj_sce)
# 
# pbmc.hvg <- modelGeneVar(traj_sce)
# pbmc.hvg$FDR[is.na(pbmc.hvg$FDR)] <- 1
# 
# traj_sce <- runPCA(traj_sce, subset_row=rownames(traj_sce)[pbmc.hvg$FDR < 0.5], scale=TRUE, ncomponents=30, assay.type="logcounts", name="PCA", BPPARAM=bpparam)
# traj_sce
```

```{r, results="asis"}
# set.seed(42)
# traj_sce <- runUMAP(traj_sce, n_neighbors=30, pca=20, BPPARAM=bpparam) # add a UMAP for plotting results later
# 
# traj_milo_gllm <- Milo(traj_sce) # from the SCE object
# reducedDim(traj_milo_gllm, "UMAP") <- reducedDim(traj_sce, "UMAP")
# 
# plotUMAP(traj_milo_gllm, colour_by="StageFL") + plotUMAP(traj_milo_gllm, colour_by="donor")
```



```{r, results="asis"}
# set.seed(42)
# traj_milo_gllm <- Milo(m2_asex_tr_sce_pt_chunks) # from the SCE object
traj_milo_gllm <- Milo(m2_asex_tr_sce) # from the SCE object
reducedDim(traj_milo_gllm, "UMAP") <- reducedDim(m2_asex_tr_sce, "UMAP.HARMONY_PC_MAN")

plotUMAP(traj_milo_gllm, colour_by="StageFL") + plotUMAP(traj_milo_gllm, colour_by="donor")
```

```{r, results="asis"}
set.seed(42)
# we build KNN graph
traj_milo_gllm <- buildGraph(traj_milo_gllm, k = 60, d = 30)
```

```{r, results="asis"}
traj_milo_gllm <- makeNhoods(traj_milo_gllm, prop = 0.2, k = 60, d=30, refined = TRUE, refinement_scheme="graph") # make nhoods using graph-only as this is faster
```

```{r, results="asis"}
# colData(traj_milo_gllm)$ObsID <- paste(colData(traj_milo_gllm)$tenx_lane, colData(traj_milo_gllm)$sampleid, sep="_")
traj_milo_gllm <- countCells(traj_milo_gllm, meta.data = data.frame(colData(traj_milo_gllm)), samples="StrainDon") # make the nhood X sample counts matrix
```

```{r, results="asis"}
head(nhoodCounts(traj_milo_gllm))
```

```{r, results="asis"}
plotNhoodSizeHist(traj_milo_gllm)
```

```{r, results="asis"}
symp.design <- data.frame(colData(traj_milo_gllm))[,c("donor", "StrainDon", "condition")]
symp.design <- distinct(symp.design)
rownames(symp.design) <- symp.design$StrainDon
## Reorder rownames to match columns of nhoodCounts(milo)
symp.design <- symp.design[colnames(nhoodCounts(traj_milo_gllm)), , drop=FALSE]
table(symp.design$condition)
```


```{r, results="asis"}
symp.design
```

```{r, results="asis", eval=FALSE}
set.seed(42)
rownames(symp.design) <- symp.design$StrainDon

da_results <- testNhoods(traj_milo_gllm, design = ~ condition + (1|donor), design.df = symp.design, fdr.weighting="graph-overlap", glmm.solver="HE-NNLS", REML=TRUE, norm.method="TMM", BPPARAM = bpparam, fail.on.error=FALSE, max.iters = 1000)
```

```{r, results="asis", eval=FALSE}
table(da_results$SpatialFDR < 0.99)
```

```{r, results="asis", eval=FALSE}
which(is.na(da_results$logFC))
```

```{r, results="asis", eval=FALSE}
which(checkSeparation(traj_milo_gllm, design.df=symp.design, condition="condition", min.val=5))
```

```{r, results="asis", eval=FALSE}
length(which(checkSeparation(traj_milo_gllm, design.df=symp.design, condition="condition", min.val=5)))
```

```{r, results="asis", fig.width=30, fig.height=20, eval=FALSE}
plotNhoodCounts(traj_milo_gllm, which(checkSeparation(traj_milo_gllm, design.df=symp.design, condition="condition", min.val=5)), design.df=symp.design, condition="condition")
```

```{r, results="asis", eval=FALSE}
head(da_results[!is.na(da_results$logFC), ])
```

```{r, results="asis", eval=FALSE}
da_results[!is.na(da_results$logFC) & da_results$Converged == T, ] %>% arrange(PValue)
```

```{r, results="asis", eval=FALSE}
traj_milo_gllm <- buildNhoodGraph(traj_milo_gllm, overlap=25)

# we need to subset the plotting results as it can't handle the NAs internally.
plotUMAP(traj_milo_gllm, colour_by="condition") + plotNhoodGraphDA(traj_milo_gllm, da_results[!is.na(da_results$logFC), ], subset.nhoods=!is.na(da_results$logFC), alpha=0.1) +  plot_layout(guides="auto" )
```


```{r, results="asis", eval=FALSE}
set.seed(42)
# we need to use place the test variable at the end of the formula
glm_results <- testNhoods(traj_milo_gllm, design = ~condition, design.df = symp.design, fdr.weighting="graph-overlap",
                          REML=TRUE, norm.method="TMM", BPPARAM = bpparam)
```

```{r, results="asis", eval=FALSE}
table(glm_results$SpatialFDR < 0.1)
```

```{r, results="asis", eval=FALSE}
plotUMAP(traj_milo_gllm, colour_by="condition") + plotNhoodGraphDA(traj_milo_gllm, glm_results, alpha=0.1) +
  plot_layout(guides="auto" )
```

```{r, results="asis", fig.width=12, fig.height=4, eval=FALSE}
comp.da <- merge(da_results, glm_results, by='Nhood')
comp.da$Sig <- "none"
comp.da$Sig[comp.da$SpatialFDR.x < 0.1 & comp.da$SpatialFDR.y < 0.1] <- "Both"
comp.da$Sig[comp.da$SpatialFDR.x >= 0.1 & comp.da$SpatialFDR.y < 0.1] <- "GLM"
comp.da$Sig[comp.da$SpatialFDR.x < 0.1 & comp.da$SpatialFDR.y >= 0.1] <- "GLMM"

ggplot(comp.da, aes(x=logFC.x, y=logFC.y)) +
    geom_point(data=comp.da[, c("logFC.x", "logFC.y")], aes(x=logFC.x, y=logFC.y),
               colour='grey80', size=1) +
    geom_point(aes(colour=Sig)) +
    labs(x="GLMM LFC", y="GLM LFC") +
    facet_wrap(~Sig) +
    NULL
```


```{r, results="asis", eval=FALSE}
gogo()
```

```{r, results="asis", eval=FALSE}
library(scRNAseq)
# uncomment the row below to allow multi-processing and comment out the SerialParam line.
# bpparam <- MulticoreParam(workers=4)
bpparam <- SerialParam()
register(bpparam)

pbmc.sce <- KotliarovPBMCData(mode="rna", ensembl=TRUE) # download the PBMC data from Kotliarov _et al._
```

```{r, results="asis", eval=FALSE}
# downsample cells to reduce compute time
prop.cells <- 0.75
n.cells <- floor(ncol(pbmc.sce) * prop.cells)
set.seed(42)
keep.cells <- sample(colnames(pbmc.sce), size=n.cells)
pbmc.sce <- pbmc.sce[, colnames(pbmc.sce) %in% keep.cells]


# downsample the number of samples
colData(pbmc.sce)$ObsID <- paste(colData(pbmc.sce)$tenx_lane, colData(pbmc.sce)$sampleid, sep="_")
n.samps <- 80
keep.samps <- sample(unique(colData(pbmc.sce)$ObsID), size=n.samps)
keep.cells <- rownames(colData(pbmc.sce))[colData(pbmc.sce)$ObsID %in% keep.samps]
pbmc.sce <- pbmc.sce[, colnames(pbmc.sce) %in% keep.cells]

pbmc.sce

```

```{r, results="asis", eval=FALSE}
pbmc.milo <- Milo(pbmc.sce) # from the SCE object
```

```{r, results="asis", eval=FALSE}
pbmc.design <- data.frame(colData(pbmc.milo))[,c("tenx_lane", "adjmfc.time", "sample", "sampleid", "timepoint", "ObsID")]
pbmc.design <- distinct(pbmc.design)
rownames(pbmc.design) <- pbmc.design$ObsID        
```


```{r, results="asis", eval=FALSE}
pbmc.design      
```


```{r, results="asis", eval=FALSE}
pbmc.design %>% count(tenx_lane, ObsID, adjmfc.time)   
```


```{r, results="asis", eval=FALSE}
pbmc.design %>%
  group_by(tenx_lane) %>%
  summarise(count = n_distinct(adjmfc.time)) 
```

```{r, results="asis", eval=FALSE}
symp.design %>% count(tenx_lane, ObsID)   
```

```{r, results="asis"}

```

```{r, results="asis"}

```


```{r, results="asis"}
# gogo()
```
## MUSCAT


```{r, results="asis"}
## solution to error https://github.com/HelenaLC/muscat/issues/88

m2_asex_tr_sce_musct <- prepSCE(m2_asex_tr_sce_pt_chunks, 
               kid = "PT_grp_unbal", # subpopulation assignments
               gid = "condition",  # group IDs (ctrl/stim)
               sid = "donor",   # sample IDs (ctrl/stim.1234)
               drop = TRUE) #
```

```{r, results="asis"}
pb <- aggregateData(m2_asex_tr_sce_musct,
    assay = "counts", 
    fun = "sum")

# one sheet per subpopulation
assayNames(pb)
```

```{r, results="asis"}
t(head(assay(pb)))
```



```{r, results="asis"}
pb_mds <- pbMDS(pb)
# use very distinctive shaping of groups & change cluster colors
pb_mds <- pb_mds + 
  scale_shape_manual(values = c(17, 4)) +
  scale_color_manual(values = RColorBrewer::brewer.pal(8, "Set2"))
# change point size & alpha
pb_mds$layers[[1]]$aes_params$size <- 5
pb_mds$layers[[1]]$aes_params$alpha <- 0.6
# pb_mds
```

```{r, results="asis"}
# run DS analysis
res <- pbDS(pb, verbose = FALSE)
# access results table for 1st comparison
tbl <- res$table[[1]]
# one data.frame per cluster
names(tbl)
```

```{r, results="asis"}
# view results for 1st cluster
k1 <- tbl[[2]]
head(format(k1[, -ncol(k1)], digits = 2))
```

```{r, results="asis"}
# construct design & contrast matrix
ei <- metadata(m2_asex_tr_sce_musct)$experiment_info
mm <- model.matrix(~ 0 + ei$group_id)
dimnames(mm) <- list(ei$sample_id, levels(ei$group_id))
contrast <- makeContrasts("Yes-No", levels = mm)

# run DS analysis
pbDS(pb, design = mm, contrast = contrast)
```


```{r, results="asis"}
# filter FDR < 5%, abs(logFC) > 1 & sort by adj. p-value
tbl_fil <- lapply(tbl, function(u) {
  u <- dplyr::filter(u, p_adj.loc < 0.3, abs(logFC) > .01)
  dplyr::arrange(u, p_adj.loc)
})

# nb. of DS genes & % of total by cluster
n_de <- vapply(tbl_fil, nrow, numeric(1))
p_de <- format(n_de / nrow(m2_asex_tr_sce_musct) * 100, digits = 3)
data.frame("#DS" = n_de, "%DS" = p_de, check.names = FALSE)
```



```{r, results="asis"}
# view top 2 hits in each cluster
top2 <- bind_rows(lapply(tbl_fil, top_n, 2, p_adj.loc))
format(top2[, -ncol(top2)], digits = 2)
```



```{r, results="asis"}
frq <- calcExprFreqs(m2_asex_tr_sce_musct, assay = "counts", th = 0)
# one sheet per cluster
assayNames(frq)
```


```{r, results="asis"}
# expression frequencies in each
# sample & group; 1st cluster
t(head(assay(frq), 5))
```



```{r, results="asis"}
gids <- levels(m2_asex_tr_sce_musct$group_id)
frq10 <- vapply(as.list(assays(frq)), 
  function(u) apply(u[, gids] > 0.00001, 1, any), 
  logical(nrow(m2_asex_tr_sce_musct)))
t(head(frq10))
```

```{r, results="asis"}
nk <- length(kids <- levels(m2_asex_tr_sce_musct$cluster_id))
ns <- length(sids <- levels(m2_asex_tr_sce_musct$sample_id))
names(kids) <- kids; names(sids) <- sids
```



```{r, results="asis"}
tbl_fil2 <- lapply(kids, function(k)
  tryCatch(
    expr = { dplyr::filter(tbl_fil[[k]], 
    gene %in% names(which(frq10[, k])))}, error = function(e) NULL))

tbl_fil2<-tbl_fil2[!sapply(tbl_fil2,is.null)]

# nb. of DS genes & % of total by cluster
n_de <- vapply(tbl_fil2, nrow, numeric(1))
p_de <- format(n_de / nrow(m2_asex_tr_sce_musct) * 100, digits = 3)
data.frame("#DS" = n_de, "%DS" = p_de, check.names = FALSE)
```


```{r, results="asis", eval=FALSE}
# tidy format; attach pre-computed expression frequencies
resDS(m2_asex_tr_sce_musct, res, bind = "row", frq = frq)

# big-table (wide) format; attach CPMs
resDS(m2_asex_tr_sce_musct, res, bind = "col", cpm = TRUE)
```


```{r, results="asis"}
# compute expression frequencies on the fly
resDS(m2_asex_tr_sce_musct, res, frq = TRUE)
```


```{r, results="asis", eval=FALSE}
library(UpSetR)
de_gs_by_k <- map(tbl_fil, "gene")
upset(fromList(de_gs_by_k))
```


```{r, results="asis", eval=FALSE}
# pull top-8 DS genes across all clusters
top8 <- bind_rows(tbl_fil) %>% 
  slice_min(p_adj.loc, n = 8, 
    with_ties = FALSE) %>% 
  pull("gene")

# for ea. gene in 'top8', plot t-SNE colored by its expression 
ps <- lapply(top8, function(g)
  .plot_dr(m2_asex_tr_sce_musct[, cs100], "TSNE", g) + 
    ggtitle(g) + theme(legend.position = "none"))

# arrange plots
plot_grid(plotlist = ps, ncol = 4, align = "vh")
```


```{r, results="asis"}
plotExpression(m2_asex_tr_sce_musct[, m2_asex_tr_sce_musct$cluster_id == "PT.4"],
  features = tbl_fil$ptime_3$gene[seq_len(1)],
  x = "sample_id", colour_by = "group_id", ncol = 3) +
  guides(fill = guide_legend(override.aes = list(size = 5, alpha = 1))) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r, results="asis"}
plotExpression(m2_asex_tr_sce_musct[, m2_asex_tr_sce_musct$cluster_id == "PT.4"],
  features = tbl_fil$ptime_4$gene[seq_len(2)],
  x = "sample_id", colour_by = "group_id", ncol = 3) +
  guides(fill = guide_legend(override.aes = list(size = 5, alpha = 1))) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```


```{r, results="asis", eval=FALSE}
# top-5 DS genes per cluster
pbHeatmap(m2_asex_tr_sce_musct, res[[1]], k = "PT.4")
```


```{r, results="asis", eval=FALSE}
# top-5 DS genes per cluster
pbHeatmap(m2_asex_tr_sce_musct, res, top_n = 1)
```

## MUSCAT mixed

```{r, results="asis", eval=F}
## solution to error https://github.com/HelenaLC/muscat/issues/88

m2_asex_tr_sce_musct_sd <- prepSCE(m2_asex_tr_sce_pt_chunks, 
               kid = "PT_grp_unbal", # subpopulation assignments
               gid = "condition",  # group IDs (ctrl/stim)
               sid = "StrainDon",   # sample IDs (ctrl/stim.1234)
               # drop = TRUE
               ) #
```

```{r, results="asis", eval=F}
# 1st approach
mm <- mmDS(m2_asex_tr_sce_musct_sd, method = "nbinom",
  n_cells = 10, n_samples = 2,
  min_count = 1, min_cells = 20)

# 2nd & 3rd approach
# mm <- mmDS(m2_asex_tr_sce_musct, method = "vst", vst = "sctransform")
# mm <- mmDS(m2_asex_tr_sce_musct, method = "nbinom")
```



## MUSCAT

```{r, results="asis"}
## solution to error https://github.com/HelenaLC/muscat/issues/88

m2_asex_tr_sce_musct345 <- prepSCE(m2_asex_tr_sce_pt_chunks_345, 
               kid = "PT_grp_unbal", # subpopulation assignments
               gid = "condition",  # group IDs (ctrl/stim)
               sid = "donor",   # sample IDs (ctrl/stim.1234)
               drop = TRUE) #
```

```{r, results="asis"}
pb <- aggregateData(m2_asex_tr_sce_musct345,
    assay = "counts", 
    fun = "sum")

# one sheet per subpopulation
assayNames(pb)
```

```{r, results="asis"}
t(head(assay(pb)))
```



```{r, results="asis"}
pb_mds <- pbMDS(pb)
# use very distinctive shaping of groups & change cluster colors
pb_mds <- pb_mds + 
  scale_shape_manual(values = c(17, 4)) +
  scale_color_manual(values = RColorBrewer::brewer.pal(8, "Set2"))
# change point size & alpha
pb_mds$layers[[1]]$aes_params$size <- 5
pb_mds$layers[[1]]$aes_params$alpha <- 0.6
# pb_mds
```

```{r, results="asis"}
# run DS analysis
res <- pbDS(pb, verbose = FALSE)
# access results table for 1st comparison
tbl <- res$table[[1]]
# one data.frame per cluster
names(tbl)
```

```{r, results="asis"}
# view results for 1st cluster
k1 <- tbl[[2]]
head(format(k1[, -ncol(k1)], digits = 2))
```

```{r, results="asis"}
# construct design & contrast matrix
ei <- metadata(m2_asex_tr_sce_musct345)$experiment_info
mm <- model.matrix(~ 0 + ei$group_id)
dimnames(mm) <- list(ei$sample_id, levels(ei$group_id))
contrast <- makeContrasts("Yes-No", levels = mm)

# run DS analysis
pbDS(pb, design = mm, contrast = contrast)
```


```{r, results="asis"}
# filter FDR < 5%, abs(logFC) > 1 & sort by adj. p-value
tbl_fil <- lapply(tbl, function(u) {
  u <- dplyr::filter(u, p_adj.loc < 0.3, abs(logFC) > .01)
  dplyr::arrange(u, p_adj.loc)
})

# nb. of DS genes & % of total by cluster
n_de <- vapply(tbl_fil, nrow, numeric(1))
p_de <- format(n_de / nrow(m2_asex_tr_sce_musct345) * 100, digits = 3)
data.frame("#DS" = n_de, "%DS" = p_de, check.names = FALSE)
```



```{r, results="asis"}
# view top 2 hits in each cluster
top2 <- bind_rows(lapply(tbl_fil, top_n, 2, p_adj.loc))
format(top2[, -ncol(top2)], digits = 2)
```



```{r, results="asis"}
frq <- calcExprFreqs(m2_asex_tr_sce_musct345, assay = "counts", th = 0)
# one sheet per cluster
assayNames(frq)
```


```{r, results="asis"}
# expression frequencies in each
# sample & group; 1st cluster
t(head(assay(frq), 5))
```



```{r, results="asis"}
gids <- levels(m2_asex_tr_sce_musct345$group_id)
frq10 <- vapply(as.list(assays(frq)), 
  function(u) apply(u[, gids] > 0.00001, 1, any), 
  logical(nrow(m2_asex_tr_sce_musct345)))
t(head(frq10))
```

```{r, results="asis"}
nk <- length(kids <- levels(m2_asex_tr_sce_musct345$cluster_id))
ns <- length(sids <- levels(m2_asex_tr_sce_musct345$sample_id))
names(kids) <- kids; names(sids) <- sids
```



```{r, results="asis"}
tbl_fil2 <- lapply(kids, function(k)
  tryCatch(
    expr = { dplyr::filter(tbl_fil[[k]], 
    gene %in% names(which(frq10[, k])))}, error = function(e) NULL))

tbl_fil2<-tbl_fil2[!sapply(tbl_fil2,is.null)]

# nb. of DS genes & % of total by cluster
n_de <- vapply(tbl_fil2, nrow, numeric(1))
p_de <- format(n_de / nrow(m2_asex_tr_sce_musct345) * 100, digits = 3)
data.frame("#DS" = n_de, "%DS" = p_de, check.names = FALSE)
```


```{r, results="asis", eval=FALSE}
# tidy format; attach pre-computed expression frequencies
resDS(m2_asex_tr_sce_musct345, res, bind = "row", frq = frq)

# big-table (wide) format; attach CPMs
resDS(m2_asex_tr_sce_musct345, res, bind = "col", cpm = TRUE)
```


```{r, results="asis"}
# compute expression frequencies on the fly
resDS(m2_asex_tr_sce_musct345, res, frq = TRUE)
```


```{r, results="asis", eval=FALSE}
library(UpSetR)
de_gs_by_k <- map(tbl_fil, "gene")
upset(fromList(de_gs_by_k))
```


```{r, results="asis", eval=FALSE}
# pull top-8 DS genes across all clusters
top8 <- bind_rows(tbl_fil) %>% 
  slice_min(p_adj.loc, n = 8, 
    with_ties = FALSE) %>% 
  pull("gene")

# for ea. gene in 'top8', plot t-SNE colored by its expression 
ps <- lapply(top8, function(g)
  .plot_dr(m2_asex_tr_sce_musct345[, cs100], "TSNE", g) + 
    ggtitle(g) + theme(legend.position = "none"))

# arrange plots
plot_grid(plotlist = ps, ncol = 4, align = "vh")
```


```{r, results="asis"}
plotExpression(m2_asex_tr_sce_musct345[, m2_asex_tr_sce_musct345$cluster_id == "PT.4"],
  features = tbl_fil$ptime_3$gene[seq_len(1)],
  x = "sample_id", colour_by = "group_id", ncol = 3) +
  guides(fill = guide_legend(override.aes = list(size = 5, alpha = 1))) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r, results="asis"}
plotExpression(m2_asex_tr_sce_musct345[, m2_asex_tr_sce_musct345$cluster_id == "PT.4"],
  features = tbl_fil$ptime_4$gene[seq_len(2)],
  x = "sample_id", colour_by = "group_id", ncol = 3) +
  guides(fill = guide_legend(override.aes = list(size = 5, alpha = 1))) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```


```{r, results="asis", eval=FALSE}
# top-5 DS genes per cluster
pbHeatmap(m2_asex_tr_sce_musct345, res[[1]], k = "PT.4")
```


```{r, results="asis", eval=FALSE}
# top-5 DS genes per cluster
pbHeatmap(m2_asex_tr_sce_musct345, res, top_n = 1)
```


## !!NOTE- MISC CODE - TO BE DELETED IF NOT NEEDED PSEUDOBULK DE ON ENTIRE CLUSTER



```{r}
harmony_int <-  readRDS("data/processed/Pf/m2_all/harmony_int.RDS")
```


```{r}
harmony_int <- JoinLayers(harmony_int)
```


```{r, warning=FALSE, message=FALSE, eval=T}
# Display the combined datasets as an interactive 3D plot
harmony_int@meta.data %>% dplyr::count(condition, donor) 

```

```{r, warning=FALSE, message=FALSE, eval=T}
# Display the combined datasets as an interactive 3D plot
harmony_int@meta.data %>% dplyr::count(condition, donor, StagePrelimC)  %>% pivot_wider(names_from = StagePrelimC, values_from = n)

```

```{r, warning=FALSE, message=FALSE, eval=T}
# Display the combined datasets as an interactive 3D plot
harmony_int@meta.data %>% dplyr::count(condition, donor)

```

```{r, fig.width=15, fig.height=5}
# visualize by batch and cell type annotation
# pseudobulk the counts based on donor-condition-celltype
pseudo_m2 <- AggregateExpression(harmony_int, assays = "RNA", return.seurat = T, group.by = c("condition", "donor", "StageFL"))

# each 'cell' is a donor-condition-celltype pseudobulk profile
tail(Cells(pseudo_m2))
```

```{r, fig.width=15, fig.height=5}
# visualize by batch and cell type annotation
pseudo_m2$stage.sympt <- paste(pseudo_m2$StageFL, pseudo_m2$condition, sep = "_")
```


```{r, fig.width=15, fig.height=5}
Idents(pseudo_m2) <- "stage.sympt"

m2_bulk.Fem.de <- FindMarkers(object = pseudo_m2, 
                         ident.1 = "Female_Yes", 
                         ident.2 = "Female_No",
                         test.use = "DESeq2")

head(m2_bulk.Fem.de, n = 15)  %>% left_join(Pf3D7_genes_plasmoDB%>% select(gene_id, Gene))
```

```{r, fig.width=15, fig.height=5}
# create a new column to annotate sample-condition-celltype in the single-cell dataset
# ifnb$donor_id.stim <- paste0(ifnb$stim, "-", ifnb$donor_id)
harmony_int$donor_symptomatic <- paste0(harmony_int$donor, harmony_int$condition, sep = "_")

# generate violin plot 
Idents(harmony_int) <- "stage.sympt"
# print(merge_dat[merge_dat$gene%in%common[1:2],c('gene','p_val.sc','p_val.bulk')])
```

```{r, fig.width=15, fig.height=5}
harmony_int$stage.sympt <- paste(harmony_int$StageFL, harmony_int$condition, sep = "_")
Idents(harmony_int) <- "stage.sympt"

```

```{r, fig.width=15, fig.height=5}
# create a new column to annotate sample-condition-celltype in the single-cell dataset
VlnPlot(harmony_int, features = "PF3D7-1101300", idents = c("Female_Yes", "Female_No"), group.by = "donor_symptomatic") 
```


```{r, fig.width=10, fig.height=5}
FeaturePlot(harmony_int[,harmony_int$StageFL == "Female"],  reduction = "umap.harmony_PC_man", features = "PF3D7-1101300", split.by = c("condition"), ncol = 2, order = T)

```

```{r, fig.width=15, fig.height=5}
Idents(pseudo_m2) <- "stage.sympt"

m2_bulk.FLE.de <- FindMarkers(object = pseudo_m2, 
                         ident.1 = "Female LE_Yes", 
                         ident.2 = "Female LE_No",
                         test.use = "DESeq2")
head(m2_bulk.FLE.de, n = 15)  %>% left_join(Pf3D7_genes_plasmoDB%>% select(gene_id, Gene))
```

```{r, fig.width=15, fig.height=5}
Idents(pseudo_m2) <- "stage.sympt"

m2_bulk.M.de <- FindMarkers(object = pseudo_m2, 
                         ident.1 = "Male_Yes", 
                         ident.2 = "Male_No",
                         test.use = "DESeq2")
head(m2_bulk.M.de, n = 15)  %>% left_join(Pf3D7_genes_plasmoDB%>% select(gene_id, Gene))
```


```{r, fig.width=15, fig.height=5}
Idents(pseudo_m2) <- "stage.sympt"

m2_bulk.ET.de <- FindMarkers(object = pseudo_m2, 
                         ident.1 = "Early trophozoite_Yes", 
                         ident.2 = "Early trophozoite_No",
                         test.use = "DESeq2")
head(m2_bulk.ET.de, n = 15)  %>% left_join(Pf3D7_genes_plasmoDB%>% select(gene_id, Gene))
```

```{r, fig.width=15, fig.height=5}
Idents(pseudo_m2) <- "stage.sympt"

m2_bulk.LR.de <- FindMarkers(object = pseudo_m2, 
                         ident.1 = "Late ring_Yes", 
                         ident.2 = "Late ring_No",
                         test.use = "DESeq2")
head(m2_bulk.LR.de, n = 15) %>% left_join(Pf3D7_genes_plasmoDB%>% select(gene_id, Gene))
```



```{r, fig.width=15, fig.height=5}
VlnPlot(pseudo_m2, features = "PF3D7-0114500", idents = c("Late ring_Yes", "Late ring_No"), group.by = "stage.sympt", ncol = 1) 
```


```{r, fig.width=15, fig.height=5}
# create a new column to annotate sample-condition-celltype in the single-cell dataset
# ifnb$donor_id.stim <- paste0(ifnb$stim, "-", ifnb$donor_id)
harmony_int$donor_symptomatic <- paste0(harmony_int$donor, harmony_int$condition, sep = "_")

# generate violin plot 
Idents(harmony_int) <- "stage.sympt"
# print(merge_dat[merge_dat$gene%in%common[1:2],c('gene','p_val.sc','p_val.bulk')])
```

```{r, fig.width=15, fig.height=5}
harmony_int$stage.sympt <- paste(harmony_int$StageFL, harmony_int$condition, sep = "_")
Idents(harmony_int) <- "stage.sympt"

```

```{r, fig.width=15, fig.height=5}
# create a new column to annotate sample-condition-celltype in the single-cell dataset
VlnPlot(harmony_int, features = "PF3D7-1401300", idents = c("Late ring_Yes", "Late ring_No"), group.by = "donor_symptomatic") 
```


```{r, fig.width=15, fig.height=5}
# create a new column to annotate sample-condition-celltype in the single-cell dataset
VlnPlot(harmony_int, features = "PF3D7-1101300", idents = c("Female_Yes", "Female_No"), group.by = "donor_symptomatic") 
```


```{r, fig.width=15, fig.height=5}
# create a new column to annotate sample-condition-celltype in the single-cell dataset
VlnPlot(harmony_int, features = "PF3D7-1401300", idents = c("Early trophozoite_Yes", "Early trophozoite_No"), group.by = "donor_symptomatic") 
```

```{r, fig.width=15, fig.height=5}
# create a new column to annotate sample-condition-celltype in the single-cell dataset
VlnPlot(harmony_int, features = "PF3D7-1240400", idents = c("Early trophozoite_Yes", "Early trophozoite_No"), group.by = "donor_symptomatic") 
```

```{r, fig.width=15, fig.height=5}
# create a new column to annotate sample-condition-celltype in the single-cell dataset
VlnPlot(harmony_int, features = "PF3D7-1101300", idents = c("Female_Yes", "Female_No"), group.by = "donor_symptomatic") 
```

###################### MISCellanous DREAMLET START - to be deleted   ############################################

## Dreamlet

```{r, results="asis"}
# Since 'ind' is the individual and 'StimStatus' is the stimulus status,
# create unique identifier for each sample
# m2_asex_tr_sce$id <- paste0(m2_asex_tr_sce$condition, m2_asex_tr_sce$donor, m2_asex_tr_sce$StrainDon)
m2_asex_tr_sce_pt_chunks$id <- paste0(m2_asex_tr_sce_pt_chunks$condition, m2_asex_tr_sce_pt_chunks$donor)


# Create pseudobulk data by specifying cluster_id and sample_id
# Count data for each cell type is then stored in the `assay` field
# assay: entry in assayNames(sce) storing raw counts
# cluster_id: variable in colData(sce) indicating cell clusters
# sample_id: variable in colData(sce) indicating sample id for aggregating cells
m2_asex_tr_pb <- aggregateToPseudoBulk(m2_asex_tr_sce_pt_chunks,
  assay = "counts",
  # cluster_id = "StageFL",
  cluster_id = "PT_grp_unbal",
  sample_id = "id",
  verbose = FALSE
)

# one 'assay' per cell type
assayNames(m2_asex_tr_pb)
```



```{r, results="asis"}
# one 'assay' per cell type
assayNames(m2_asex_tr_pb)

metadata(m2_asex_tr_pb)$aggr_means
```


```{r, results="asis"}

# form <- ~ (1|donor) + condition + pseudotime
form <- ~ condition + pseudotime

# Normalize and apply voom/voomWithDreamWeights
res.proc <- processAssays(m2_asex_tr_pb, form, min.count = 2)

# the resulting object of class dreamletProcessedData stores
# normalized data and other information
res.proc
```

```{r, results="asis", eval=FALSE}
gogo()
```

```{r, results="asis"}
# Normalize and apply voom/voomWithDreamWeights
details(res.proc)
```

```{r, results="asis"}
# show voom plot for each cell clusters
plotVoom(res.proc)
```

```{r, results="asis"}

# run variance partitioning analysis
vp.lst <- fitVarPart(res.proc, form)
```

```{r, results="asis"}
# Show variance fractions at the gene-level for each cell type
genes <- vp.lst$gene[1:2]
plotPercentBars(vp.lst[vp.lst$gene %in% genes, ])
```


```{r, results="asis", fig.width=10, fig.height=10}
# Summarize variance fractions genome-wide for each cell type
plotVarPart(vp.lst, label.angle = 60)
```

```{r, results="asis"}
# Differential expression analysis within each assay,
# evaluated on the voom normalized data
res.dl <- dreamlet(res.proc, form, method = "random")

# names of estimated coefficients
coefNames(res.dl)
```

```{r, results="asis"}
# the resulting object of class dreamletResult
# stores results and other information
res.dl
```

```{r, results="asis"}
plotVolcano(res.dl, coef = "conditionSymptomatic")
```

```{r, results="asis"}
# genes <- c("PF3D7-1240400", "PF3D7-1240900")
# plotGeneHeatmap(res.dl, coef = "conditionSymptomatic", genes = genes)
```

```{r, results="asis"}
conflicts_prefer(dreamlet::topTable)
# results from full analysis
topTable(res.dl, coef = "conditionSymptomatic")
```

```{r, results="asis"}
# only B cells
topTable(res.dl[["PT.3"]], coef = "conditionSymptomatic")
```

```{r, results="asis"}
plotForest(res.dl, coef = "conditionSymptomatic", gene = c("PF3D7-1240400"))
```

```{r, results="asis"}
# get data
df <- extractData(res.proc, "PT.3", genes = "PF3D7-1240400")

# set theme
thm <- theme_classic() +
  theme(aspect.ratio = 1, plot.title = element_text(hjust = 0.5))

# make plot
ggplot(df, aes(condition, `PF3D7-1240400`)) +
  geom_boxplot() +
  thm +
  ylab(bquote(Expression ~ (log[2] ~ CPM))) +
  ggtitle("PF3D7-1240400")
```

```{r, results="asis"}
# create a contrasts called 'Diff' that is the difference between expression
# in the stimulated and controls.
# More than one can be specified
contrasts <- c(Diff = "conditionSymptomatic - conditionAsymptomatic")

# Evalaute the regression model without an intercept term.
# Instead estimate the mean expression in stimulated, controls and then
# set Diff to the difference between the two
res.dl2 <- dreamlet(res.proc, ~ 0 + condition, contrasts = contrasts)

# see estimated coefficients
coefNames(res.dl2)
```

```{r, results="asis"}
# Volcano plot of Diff
plotVolcano(res.dl2[1:2], coef = "Diff")
```

```{r, results="asis"}
# results from full analysis
topTable(res.dl, coef = "conditionSymptomatic")

```


```{r, results="asis"}
# results from full analysis
# topTable(res.dl, coef = "conditionAsymptomatic")

```


```{r, results="asis", eval=FALSE}
# results from full analysis
topTable(res.dl, coef = "conditionAsymptomatic")

```

```{r, results="asis", eval=FALSE}
gogo()
```

###################### MISCellanous DREAMLET END - to be deleted   ############################################

