---
title: "Integration of donor only and with lab"
date: "2024-01-05"
author: "Jesse Rop"
output:
  html_document:
    toc: true
    theme: united
    code_folding: hide
    df_print: paged
---


```{r setup}
sample_set = "Mali2"

knitr::opts_knit$set(root.dir = paste0("/lustre/scratch126/tol/teams/lawniczak/users/jr35/phd/", sample_set))
```

### Common functions

#### Source commong functions, varriables and tables
```{r}
##Boxplot labeling function
source("../multipurpose_scripts_lnk/seurat_sce/seu_sce_common_fns.R")
source("../multipurpose_scripts_lnk/plotting_common_fns.R")
source("mali22_code_lnk/pf_common_vars.R")
source("mali22_code_lnk/local_rmd_nbooks/lemieux_et_al_pipeline_functions.R")
```


```{r}
## Load reference set
# We use a cleaned version of the Bozdech et al. (2003) overview data set with
# NULL genes removed and duplicates averaged, yielding n = 3532 genes.
# Can also use 3D7 and Dd2 reference sets
# z <- read.csv("data/raw/Pf/andrade_lemiux_MLE/bozdech_Hb3_clean.csv", as.is = T)
z <- read.csv("data/raw/Pf/andrade_lemiux_MLE/bozdech_Hb3_clean.csv")
z
```
```{r}
z1 <- read.csv("data/raw/Pf/andrade_lemiux_MLE/bozdech_Hb3_clean.csv", as.is = T)
# z <- read.csv("data/raw/Pf/andrade_lemiux_MLE/bozdech_Hb3_clean.csv")
z1
```

```{r}
## separate old names and match to new names
Pf3D7_genes_plasmoDB_w_old_ids <- Pf3D7_genes_plasmoDB %>% mutate(across(`Previous ID(s)`, ~str_remove(., "Previous IDs: "))) %>% separate_wider_delim(`Previous ID(s)`, delim = ";", names = c("old_id1", "old_id2", "old_id3", "old_id4", "old_id5", "old_id6", "old_id7", "old_id8"), too_few = "align_start", too_many = "debug")

z_cln_id <- Pf3D7_genes_plasmoDB_w_old_ids %>% pivot_longer(cols = c("old_id1", "old_id2", "old_id3", "old_id4", "old_id5", "old_id6", "old_id7", "old_id8"), names_to = "id_cat", values_to = "old_ids", values_drop_na = T) %>% select(gene_id, old_ids) %>% inner_join(z, ., by = c("Name" = "old_ids"))

z_cln_id %>% arrange(Name) %>% group_by(Name) %>% add_count() %>% filter(n == 2) %>% select(gene_id, Name) %>% filter(!str_detect(gene_id, "\\.")) 
```

```{r}
## separate old names and match to new names
z_cln_id_uniq <- z_cln_id %>% arrange(Name) %>% group_by(Name) %>% add_count(name = "old_id_n") %>% ungroup() %>% filter(old_id_n == 1 | (old_id_n == 2 & str_detect(gene_id, "\\."))) %>% group_by(gene_id) %>% add_count(name = "new_id_n") %>% ungroup() %>% filter(new_id_n == 1) %>% select(-c(Name, new_id_n, old_id_n))

z_cln_id_uniq

```

```{r}
z_cln_id_uniq %>% arrange(gene_id) %>% group_by(gene_id) %>% add_count() %>% filter(n == 2)
```

```{r}
##Load the input data file
x1 <- read.csv("data/raw/Pf/andrade_lemiux_MLE/wgenes.csv", as.is=T, sep = ';')
x1 <- x1[,c(1,2:13)]
x1
```


```{r}
##Load the input data file
z = z_cln_id_uniq %>% rename("Name" = gene_id)
```

```{r}
##Load the input data file
x = cor_assay_bulk_painter %>% data.frame() %>% rownames_to_column("Name")
```



################################################################################
###
### Pre-process
###


```{r}
head(x$Name)
head(z$Name)
```

```{r}
## Sync data
# Ensures that data sets contain the same genes and are in the same order
# Note that first column must be gene description titled "Name"
data <- sync_data(x, z)
x <- data[[1]]
z <- data[[2]]
```

```{r}
x 
z
```

## Store names

```{r}
#genenames <- x$?..Name

#genenames <- x$Name
## Make data ordinal
# This is necessary for comparisons between one- and two-color arrays
# Returns a matrix and if `use.name = T', removes first "Name" column
x <- ordinal(x, use.name = T)
z <- ordinal(z, use.name = T)


## If same arry type, remove first column and make into a matrix
#x <- as.matrix(x[,-1])
#z <- as.matrix(z[,-1])
```

## Impute missing values

```{r}
# For later interpretability, we impute the missing time points (23 and 29 HPI)
# for the Hb3 reference set. This method interpolates missing values using
# smoothing splines, though other approaches are possible.
z.na <- cbind(z[,1:22], rep(NA, nrow(z)), z[,23:27],
              rep(NA, nrow(z)), z[,28:46])
z <- t(apply(z.na, 1, smooth.missing))
```



## Load sigma_epsilon
```{r}
# Estimate of background noise obtained by comparing Hb3 and 3D7 samples
sigma.epsilon <- 789.9056
```

################################################################################
###
### Get smooth reference data
###


## Check default log-likelihood parameters, stored in ll.par

```{r}
## Smooth reference data set
# This smooths the time course for each gene in the reference set using a
# smoothing spline. There is also a loess implementation.
# `spar' is the parameter of smoothing
z.smooth <- smooth.ref(z, method = "spline", spar = 0.5)
```


```{r}
## Estimate residual of smoothing
z.smooth.hourly <- z.smooth[,ll.par$hourly]
sigma.eta <- mean(sd(z - z.smooth.hourly, na.rm = T), na.rm=T)
```

```{r}
## Calculate new sigma based on sigma_eta and sigma_epsilon
new.sigma <- sqrt(sigma.eta^2 + sigma.epsilon^2)
```

################################################################################
###
### Get Log-Likelihood, MLE, and Confidence Intervals
###


#-----------------------
# Main function options
#-----------------------
# B: number of iterations; 100 is default; 500 is better
# sample.rate: rate of subsampling; 0.50 is default
# alpha: for confidence intervals; alpha = 0.05 is default

```{r}
ll <- compute.ll(x = x, z = z.smooth, sigma = new.sigma,
                 bootstrap = T,
                 B = 100,
                 sample.rate = 0.50)



# mle(ll) --> extracts MLEs
# loglik(ll) --> extracts log-likelihoods
# ci(ll) --> extracts confidence intervals
```

```{r}
## To obtain only the log-likelihoods without bootstrapped confidence intervals
ll <- compute.ll(x = x, z = z.smooth, sigma = new.sigma,
                 bootstrap = F)
```

###############################################################################
###
### Plot Logliks and MLEs
###
```{r}
## Plot the log-likelihood curves
plot.ll(ll)

## Plot the maximum likelihood estimates with 95% confidence intervals
plot.mle(ll)
```

```{r}
pt_zb_hpi <- map(ll, ~.x$mle) %>% data.frame() %>% t() %>% data.frame() %>% rownames_to_column("pseudotime") %>% rename_all(., ~str_replace(., "\\.", "ZB_hpi_corr")) %>% mutate(across(pseudotime, ~factor(., levels = paste0("ptime.", 1:12))))
```

```{r}
pt_zb_hpi %>% ggplot(aes(y = pseudotime, x=ZB_hpi_corr)) + geom_point() + theme_classic()
```

