---
title: "Integration of donor only and with lab"
date: "2024-01-05"
author: "Jesse Rop"
output:
  html_document:
    toc: true
    theme: united
    code_folding: hide
    df_print: paged
---

```{r}
suppressPackageStartupMessages({
  library(Seurat)
  # library(sctransform)
  # library(rgl)
  library(ggrepel)
  # library(ggpubr)
  library(dplyr) 
  library(stringr)
  library(purrr)
  library(readr)
  library(tibble)
  library(ggplot2)
  # library(ggnewscale)
  library(tidyr)
  library(cowplot)
  # library(scDblFinder)
  library(SeuratData)
  library(SeuratDisk)
  # library(scmap)
  library(SingleCellExperiment)
  library(pheatmap)
  library(plotly)
  # library(Hmisc)
  # library(rmarkdown)
  library(RColorBrewer)
  # library(MAST)
  library(ggvenn)
  # library("biomaRt")
  library(conflicted)
  # library(SoupX)
  library(MetBrewer)
  # library(dreamlet)
  # library(muscat)
  # library(ExperimentHub)
  # library(zenith)
  library(scater)
  library(slingshot)
  library(harmony)
  library(ggforce)
  # library(miloR)
  library(scater)
  library(scran)
  library(patchwork)
  # library(scRNAseq)
  library(scuttle)
  library(irlba)
  library(BiocParallel)
  library(tradeSeq)
  library(ComplexHeatmap)
  library(circlize)
  conflict_prefer("select", "dplyr")
  conflict_prefer("filter", "dplyr")
  conflict_prefer("rename", "dplyr")
  conflicts_prefer(dplyr::count)
  conflicts_prefer(base::as.data.frame)
  
})
```


```{r setup}
sample_set = "Mali2"

# knitr::opts_knit$set(root.dir = paste0("/Users/jr35/Google Drive/My Drive/PHD_documents/malaria_phd/", sample_set))
knitr::opts_knit$set(root.dir = paste0("/lustre/scratch126/tol/teams/lawniczak/users/jr35/phd/", sample_set))
```


#  integration of all MSC

### Common functions

#### Source commong functions, varriables and tables
```{r}
##Boxplot labeling function
source("../multipurpose_scripts_lnk/seurat_sce/seu_sce_common_fns.R")
source("../multipurpose_scripts_lnk/plotting_common_fns.R")
source("/nfs/users/nfs_j/jr35/Mali2/mali22_code/pf_common_vars.R")
```

```{r}
# optimum_k_qcd <- c("MSC33" =8, "MSC48" =4, "MSC49" =6, "MSC55" =7, "MSC60" =7, "MSC66" =8, "MSC38" =4, "MSC40" =8, "MSC50_MACS" =1, "MSC53" =1, "MSC57" =1, "MSC24" =1, "MSC37" =5, "MSC39" =4, "MSC45" =1, "MSC50_SLO" =1, "MSC54" =1, "MSC67" =1, "MSC68" =3, "MSC70" =1)
# 
# optimum_k_qcd <- optimum_k_qcd[sort(names(optimum_k_qcd))]
```


```{r}
# m2_seu_raw_cb_norm_harmony_PC_manl <-  readRDS("data/processed/Pf/m2_all/m2_seu_raw_cb_norm_harmony_PC_manl.RDS")
# all_msc_cb_no_dbt_mrg_noM12_sct_rpca_sct_PC10 <-  readRDS("data/processed/Pf/m2_all/all_msc_cb_no_dbt_mrg_noM12_sct_rpca_sct_PC10.RDS")
# all_msc_cb_no_dbt_mrg_noM12_sct_rpca_sct_PC10@meta.data[,'donor'] <- str_remove(all_msc_cb_no_dbt_mrg_noM12_sct_rpca_sct_PC10@meta.data[,'orig.ident'], ".mrna")
```


```{r}
m2_all_raw_mrg <- readRDS("data/processed/Pf/m2_all/m2_all_raw_mrg.RDS")
```

```{r}
DimPlot(m2_all_raw_mrg, reduction = "umap.rpca_sct_PC10", group.by = "sR_fine_stg", cols = sp_col_new)

```

```{r}
DimPlot(m2_all_raw_mrg, reduction = "umap.rpca_sct_PC10", group.by = "sR_fine_stg", cols = sp_col_new, dims = c(1,2))

```

```{r}
DimPlot(m2_all_raw_mrg, reduction = "umap.rpca_sct_PC10", group.by = "sR_fine_stg", cols = sp_col_new, dims = c(1,3))

```
```{r, warning=FALSE, message=FALSE}
VlnPlot(m2_all_raw_mrg, features = "nFeature_RNA", group.by = "stage", pt.size = 0) + geom_hline(yintercept = c(200, 400, 500, 1000))
```

```{r}
all_sct_intgrtd_fem <- m2_all_raw_mrg[, m2_all_raw_mrg$sR_fine_stg  == "Female"]
all_sct_intgrtd_femLE <- m2_all_raw_mrg[, m2_all_raw_mrg$sR_fine_stg  == "Female (LE)"]

```

```{r}
all_sct_intgrtd_fem_lst <- list(all_sct_intgrtd_fem, all_sct_intgrtd_femLE) %>% set_names(c("fem", "femLE"))

```

```{r}
# names(all_sct_intgrtd_fem_lst)

```
##############################START Female DE ##########################

```{r}
map(all_sct_intgrtd_fem_lst, ~DimPlot(.x, reduction = "umap.rpca_sct_PC10", group.by = "rpca_sct_PC10_clusters", label = T))
```


```{r}
pmap(list(all_sct_intgrtd_fem_lst, c(0,5.1), c(-10.5,-2)),  ~DimPlot(..1, reduction = "umap.rpca_sct_PC10", group.by = "rpca_sct_PC10_clusters", label = F, pt.size = 0.4) + geom_vline(xintercept =  ..2) + geom_hline(yintercept =  ..3))
```

```{r}
all_sct_intgrtd_fem <- pmap(list(all_sct_intgrtd_fem_lst, c(0,5.1), c(-10.5,-2)),  ~..1[, ..1@reductions[["umap.rpca_sct_PC10"]]@cell.embeddings[, "umaprpcasctPC10_1"] > ..2 & ..1@reductions[["umap.rpca_sct_PC10"]]@cell.embeddings[, "umaprpcasctPC10_2"] > ..3]
)
```

```{r}
map(all_sct_intgrtd_fem, ~DimPlot(.x, reduction = "umap.rpca_sct_PC10", cells.highlight = colnames(.x[,.x$rpca_sct_PC10_clusters == "25"])))
```

```{r}
map(all_sct_intgrtd_fem, ~DimPlot(.x, reduction = "umap.rpca_sct_PC10", group.by = "rpca_sct_PC10_clusters", label = T))
```

```{r, warning=FALSE, message=FALSE}
for (nm in names(all_sct_intgrtd_fem)) {
  DefaultAssay(all_sct_intgrtd_fem[[nm]]) <- "RNA"
  all_sct_intgrtd_fem[[nm]][["SCT"]] <- NULL
}
```


```{r}
# DefaultAssay(all_sct_intgrtd_fem) <- "RNA"
# all_sct_intgrtd_fem[["SCT"]] <- NULL
```


```{r}
map(all_sct_intgrtd_fem, ~table(.x$orig.ident))
```

```{r}
## Donors with more than 50 cells
don_50 <- map(all_sct_intgrtd_fem, ~table(.x$donor) %>% data.frame() %>% filter(Freq >= 100) %>% pull(Var1) %>% as.character)
```

```{r}
## Retain only donors with more than 50 cells
all_sct_intgrtd_fem <- map2(all_sct_intgrtd_fem, don_50, ~subset(.x, subset = donor %in% .y))
```

## !!NOTE - NOT SURE WHY, I have to remove donors with less than 100 cells for FindVariableFeatures to work
```{r, warning=FALSE, message=FALSE}
# run standard seurat normalization
all_sct_intgrtd_fem <- map(all_sct_intgrtd_fem, ~NormalizeData(.x))
all_sct_intgrtd_fem <- map2(all_sct_intgrtd_fem, c(1000, 200), ~FindVariableFeatures(.x, nfeatures = .y))
all_sct_intgrtd_fem <- map(all_sct_intgrtd_fem, ~ScaleData(.x, verbose = F))

```

```{r, warning=FALSE, message=FALSE}
# These are now standard steps in the Seurat workflow for visualization and clustering
all_sct_intgrtd_fem <- map(all_sct_intgrtd_fem, ~RunPCA(.x, verbose = FALSE, npcs = 20))

```

```{r, warning=FALSE, message=FALSE}

all_sct_intgrtd_fem <- map(all_sct_intgrtd_fem, ~FindNeighbors(.x, dims = 1:10, reduction = "pca"))
all_sct_intgrtd_fem <- map(all_sct_intgrtd_fem, ~FindClusters(.x, resolution = 0.3))
```


```{r, warning=FALSE, message=FALSE}
# run sctransform
map(all_sct_intgrtd_fem, ~DimPlot(.x, reduction = "pca"))
map(all_sct_intgrtd_fem, ~DimPlot(.x, reduction = "pca", group.by = "orig.ident"))


```



```{r, warning=FALSE, message=FALSE}
options(future.globals.maxSize = 1e+09)

all_sct_intgrtd_fem <- map(all_sct_intgrtd_fem, ~IntegrateLayers(
  object = .x,
  method = HarmonyIntegration,
  orig.reduction = "pca", 
  new.reduction = "harmony",
  # normalization.method = "SCT",
  verbose = F,
  # k.weight = 98,
))


```


```{r, warning=FALSE, message=FALSE}

all_sct_intgrtd_fem <- map(all_sct_intgrtd_fem, ~FindNeighbors(.x, dims = 1:10, reduction = "harmony"))
all_sct_intgrtd_fem <- map(all_sct_intgrtd_fem, ~FindClusters(.x, resolution = 0.3))
```

```{r, warning=FALSE, message=FALSE}
# These are now standard steps in the Seurat workflow for visualization and clustering
all_sct_intgrtd_fem <- map(all_sct_intgrtd_fem, ~RunUMAP(.x, dims = 1:6, verbose = FALSE, n.components = 3, seed.use = 867, reduction = "harmony"))

```


```{r, warning=FALSE, message=FALSE}
# run sctransform
map(all_sct_intgrtd_fem, ~DimPlot(.x, reduction = "umap"))
map(all_sct_intgrtd_fem, ~DimPlot(.x, group.by = "orig.ident"))

map(all_sct_intgrtd_fem, ~FeaturePlot(.x, reduction = "umap", features = "nCount_RNA"))

```

```{r, warning=FALSE, message=FALSE}
# run sctransform
map(all_sct_intgrtd_fem, ~DimPlot(.x, reduction = "harmony"))
map(all_sct_intgrtd_fem, ~DimPlot(.x, reduction = "harmony", dims = c(1,3)))

# FeaturePlot(all_sct_intgrtd_fem, reduction = "umap", features = "nCount_RNA")

```



```{r, warning=FALSE, message=FALSE, eval=T}
# Display the combined datasets as an interactive 3D plot

plot_ly(data = data.frame(all_sct_intgrtd_fem[[1]]@reductions[["umap"]]@cell.embeddings),
        x = ~umap_1, y = ~umap_2, z = ~umap_3,
        color= all_sct_intgrtd_fem[[1]]@meta.data$seurat_clusters,
        # colors = c('blue', 'red'),
        opacity = 0.8,
        size = 1, 
        type = 'scatter3d'
)

```


```{r, warning=FALSE, message=FALSE, eval=F}
# Display the combined datasets as an interactive 3D plot

plot_ly(data = data.frame(all_sct_intgrtd_fem@reductions[["umap"]]@cell.embeddings),
        x = ~umap_1, y = ~umap_2, z = ~umap_3,
        color= all_sct_intgrtd_fem@meta.data$donor,
        # colors = c('blue', 'red'),
        opacity = 0.8,
        size = 1, 
        type = 'scatter3d'
)

```

```{r, warning=FALSE, message=FALSE}
map(all_sct_intgrtd_fem, ~.x@meta.data %>% filter(seurat_clusters %in% c(4,6)) %>% count(seurat_clusters, orig.ident) %>% group_by(seurat_clusters) %>% arrange(desc(n), .by_group = T))
```

```{r, warning=FALSE, message=FALSE}
map(all_sct_intgrtd_fem, ~.x@meta.data %>% count(seurat_clusters, orig.ident) %>% group_by(seurat_clusters) %>% arrange(desc(n), .by_group = T))
```



```{r, warning=FALSE, message=FALSE}
all_sct_intgrtd_fem_jnd <- map(all_sct_intgrtd_fem, ~JoinLayers(.x)) 
```

```{r, warning=FALSE, message=FALSE}
for (nm in names(all_sct_intgrtd_fem_jnd)) {
  all_sct_intgrtd_fem_jnd[[nm]][["RNA"]]$counts <- as(all_sct_intgrtd_fem_jnd[[nm]][["RNA"]]$counts, Class = "dgCMatrix")
  all_sct_intgrtd_fem_jnd[[nm]][["RNA"]]$data <- as(all_sct_intgrtd_fem_jnd[[nm]][["RNA"]]$data, Class = "dgCMatrix")
  Idents(all_sct_intgrtd_fem_jnd[[nm]]) <- "seurat_clusters"
}
```

```{r, warning=FALSE, message=FALSE}
all_sct_intgrtd_fem_jnd <- map(all_sct_intgrtd_fem_jnd, ~FindClusters(.x, resolution = 0.35))
```

```{r, warning=FALSE, message=FALSE}
# run sctransform
map(all_sct_intgrtd_fem_jnd, ~DimPlot(.x, reduction = "umap", pt.size = 0.2, label = T, dims = c(1,3)))

```

```{r, warning=FALSE, message=FALSE}
fem_mks <- map(all_sct_intgrtd_fem_jnd, ~FindAllMarkers(.x, assay = "RNA", slot = "data"))
```


```{r, warning=FALSE, message=FALSE}
# run sctransform
map(all_sct_intgrtd_fem_jnd, ~DimPlot(.x, reduction = "umap", pt.size = 0.1, label = T))

```


```{r, warning=FALSE, message=FALSE}
fem_mks_anot <- map(fem_mks, ~.x %>%  left_join(., Pf3D7_genes_plasmoDB[,c("gene_id", "Gene")], by = c("gene" = "gene_id")))

fem_mks_anot
```


```{r, warning=FALSE, message=FALSE}
fem_mks_anot[["fem"]] %>% filter(cluster == "6" & avg_log2FC >0)# %>% filter( avg_log2FC >1)
fem_mks_anot[["fem"]] %>% filter(cluster == "6" & avg_log2FC >0 & pct.2 < 0.1)# %>% filter( avg_log2FC >1)

```

```{r, warning=FALSE, message=FALSE, fig.width=15, fig.height=4}
#KAHRP, HSP70x, GARP, HRPIII, 
map(all_sct_intgrtd_fem_jnd, ~FeaturePlot(.x, reduction = "umap", dims = c(1,3), pt.size = 0.15, label = T, features = c("PF3D7-0202000", "PF3D7-0831700", "PF3D7-0113000", "PF3D7-1372200"), ncol = 4))

```

```{r, warning=FALSE, message=FALSE}
fem_mks_anot[["fem"]] %>% filter(cluster == "4" & avg_log2FC >0)
```

```{r, warning=FALSE, message=FALSE}
# ap2-fg, pfg
fem_mks_anot[["fem"]] %>% filter(gene %in% c("PF3D7-1317200", "PF3D7-1146800") )
```

```{r, warning=FALSE, message=FALSE}
fem_mks_anot[["fem"]] %>% filter(cluster == "5" & avg_log2FC >0)# %>% filter( avg_log2FC >1)
# fem_mks_anot[["fem"]] %>% filter(cluster == "5" & avg_log2FC >0 & p_val_adj < 0.05) %>% arrange(desc(avg_log2FC))# %>% filter( avg_log2FC >1)

## cluster 5 could be activated-ish https://journals.plos.org/plospathogens/article?id=10.1371/journal.ppat.1010510#ppat-1010510-g002 
```

```{r, warning=FALSE, message=FALSE, fig.width=18, fig.height=4}
map(all_sct_intgrtd_fem_jnd, ~FeaturePlot(.x, reduction = "umap", dims = c(1,3), pt.size = 0.15, label = T, features = c("PF3D7-0411000", "PF3D7-0628100", "PF3D7-0715200", "PF3D7-1143100", "nFeature_RNA"), ncol = 5))

```

```{r, warning=FALSE, message=FALSE}
activated_gams_ngwa <- c("PF3D7-0827800", "PF3D7-1246200", "PF3D7-1218800", "PF3D7-1239400", "PF3D7-0704100", "PF3D7-0417400", "PF3D7-1321000")

fem_mks_anot[["fem"]] %>% filter(gene %in% activated_gams_ngwa & avg_log2FC > 0)

```


```{r, warning=FALSE, message=FALSE, fig.width=15, fig.height=4}
map(all_sct_intgrtd_fem_jnd, ~FeaturePlot(.x, reduction = "umap", dims = c(1,3), pt.size = 0.15, label = T, features = c("PF3D7-0411000", "PF3D7-1143100", "PF3D7-1107800", "PF3D7-0209000", "nFeature_RNA"), ncol = 4))

```

```{r, warning=FALSE, message=FALSE, fig.width=15, fig.height=4}
map(all_sct_intgrtd_fem_jnd, ~FeaturePlot(.x, reduction = "umap", dims = c(1,3), pt.size = 0.15, label = T, features = c("PF3D7-1143100", "PF3D7-1107800", "PF3D7-0209000"), ncol = 3))

```

```{r, warning=FALSE, message=FALSE}
fem_mks_anot[["femLE"]] %>% filter(cluster == "3" & avg_log2FC > 0) 

```

```{r, warning=FALSE, message=FALSE}
# map(fem_mks_anot, ~.x %>% filter(cluster == "6" & avg_log2FC >1))

```

```{r, warning=FALSE, message=FALSE}
# map(fem_mks_anot, ~.x %>% filter(cluster == "6" & avg_log2FC >1 & pct.2 < 0.1))

```


```{r, warning=FALSE, message=FALSE}
fem_clustrs_w_asexuals <- c("6", "3") %>% set_names(c("fem", "femLE"))
```

```{r, warning=FALSE, message=FALSE}
saveRDS(all_sct_intgrtd_fem_jnd, 'data/processed/Pf/m2_all/all_sct_intgrtd_fem_jnd.RDS')

```

```{r}
obj_names = c('M55', 'no_M55')

```

```{r}
m2_all_scvi_mdata <- map(obj_names, ~read_csv(paste0("/lustre/scratch126/tol/teams/lawniczak/users/jr35/phd/Mali2/data/processed/Pf/m2_all/m2_all_scvi_mdata_", .x,".csv"))) %>% set_names(obj_names)
```


```{r, warning=FALSE, message=FALSE}
all_sct_intgrtd_fem_jnd <- map(all_sct_intgrtd_fem_jnd, ~AddMetaData(.x, m2_all_scvi_mdata[["M55"]] %>% column_to_rownames('...1') %>% select(leiden, `_scvi_labels`)))
```


```{r, warning=FALSE, message=FALSE}
map(all_sct_intgrtd_fem_jnd, ~DimPlot(.x, group.by = "leiden"))
```

```{r, warning=FALSE, message=FALSE}
map(all_sct_intgrtd_fem_jnd, ~.x@meta.data %>% filter(seurat_clusters == 6) %>% count(leiden))
```

```{r}
# REmove poor QC asexual like cluster 
all_sct_intgrtd_fem_jnd_no_asx_dblts_cln <- map2(all_sct_intgrtd_fem_jnd, fem_clustrs_w_asexuals, ~.x[, .x@meta.data %>% 
                                               rownames_to_column('bcode') %>% 
                                               filter(seurat_clusters != .y)%>% 
                                               pull(bcode)])


```

```{r, warning=FALSE, message=FALSE}
saveRDS(all_sct_intgrtd_fem_jnd_no_asx_dblts_cln, 'data/processed/Pf/m2_all/all_sct_intgrtd_fem_jnd_no_asx_dblts_cln.RDS')

```

```{r, warning=FALSE, message=FALSE}
# all_sct_intgrtd_fem_jnd_no_asx_dblts_cln <- readRDS('data/processed/Pf/m2_all/all_sct_intgrtd_fem_jnd_no_asx_dblts_cln.RDS')

```

```{r}
# Retain only strains with sufficient numbers 
all_sct_intgrtd_fem_jnd_no_asx_dblts <- all_sct_intgrtd_fem_jnd_no_asx_dblts_cln[, all_sct_intgrtd_fem_jnd_no_asx_dblts_cln@meta.data %>% 
                                               rownames_to_column('bcode') %>% 
                                               group_by(seurat_clusters, donor, Strain_DN) %>% 
                                               filter(n() >=50) %>%
                                               group_by(seurat_clusters, donor) %>% 
                                               filter(n_distinct(Strain_DN) >1) %>%
                                               # filter(n() >=100) %>% 
                                               pull(bcode)]


```



```{r}
## subset where field parasites are for slingshot
all_sct_intgrtd_fem_jnd_no_asx_dblts@meta.data %>% count(seurat_clusters, donor, Strain_DN) %>% group_by(seurat_clusters, donor) %>% mutate(strn_n = n_distinct(Strain_DN)) %>% filter(strn_n >1)
```


Ω### NOTE - Donor
```{r}
DimPlot(all_sct_intgrtd_fem_jnd_no_asx_dblts, reduction = "umap", group.by = "seurat_clusters" )
```
```{r}
DimPlot(all_sct_intgrtd_fem_jnd_no_asx_dblts, reduction = "umap", group.by = "stage" )
```

```{r, warning=FALSE, message=FALSE}
saveRDS(all_sct_intgrtd_fem_jnd_no_asx_dblts, 'data/processed/Pf/m2_all/all_sct_intgrtd_fem_jnd_no_asx_dblts.RDS')

```


```{r, warning=FALSE, message=FALSE}
all_sct_intgrtd_fem_jnd_no_asx_dblts <- readRDS('data/processed/Pf/m2_all/all_sct_intgrtd_fem_jnd_no_asx_dblts.RDS')

```

## MAST DE adjusting for pseudotime

MAST DE on the entire dataset before subsampling. Subsampling is not necessary for MAST and also for tradeseq
### Prepare dataset
Normalizing and scaling the data in the subseted seurat objects - may have been done before on the larger parent object but just repeating since will not change anything as it starts from the raw counts.

Adding pseudotime metadata from the single cell experiment object 
Scaling is for visualization


```{r}
##REMEMBER TO REMOVE CLUSTERS WITH VERY FEW CELLS AND DOUBLETS SEURAT
# msc_gam_combi_noLE_de <- msc_gam_combi_noLE_de[, cells_de]
```


```{r, message=FALSE, warning=FALSE}
# Add to metadata the dataset/conditions and ptime cat
# all_sct_intgrtd_fem_jnd_no_asx_dblts <- all_sct_intgrtd_fem_jnd_no_asx_dblts
all_sct_intgrtd_fem_jnd_no_asx_dblts[["SCT"]] <- NULL
all_sct_intgrtd_fem_jnd_no_asx_dblts[["RNA"]]$scale.data <- NULL
DefaultAssay(all_sct_intgrtd_fem_jnd_no_asx_dblts) <- "RNA"

```

```{r, message=FALSE, warning=FALSE}
# Add to metadata the dataset/conditions and ptime cat
all_sct_intgrtd_fem_jnd_no_asx_dblts@meta.data[, "don_seu_clutrs"] <- paste0(all_sct_intgrtd_fem_jnd_no_asx_dblts$donor, "_", all_sct_intgrtd_fem_jnd_no_asx_dblts$seurat_clusters)

```

```{r, message=FALSE, warning=FALSE}
# Add to metadata the dataset/conditions and ptime cat
all_fem_de_bal_nosct <- SplitObject(all_sct_intgrtd_fem_jnd_no_asx_dblts, split.by = "don_seu_clutrs")

```

```{r, message=FALSE, warning=FALSE}
all_gns = map(all_fem_de_bal_nosct, ~unique(rownames(.x)))

fem_m2_bal_seu <- map2(all_fem_de_bal_nosct,all_gns, ~{
  .x %>%
         NormalizeData(., normalization.method = "LogNormalize", scale.factor = 10000, verbose = FALSE) %>%
         ScaleData(., features = .y,  verbose = FALSE)
  })

# m2_bal_seu <- list(m2_bal_seu) %>% set_names("gam")

```

### Actual MAST DE between pairwise strains and fieldvslab 

MAST DE adjusting for pseudotime as a latent variable

```{r, message=FALSE, warning=FALSE}
de_idents <- rep("Strain_DN", length = length(fem_m2_bal_seu))
fem_m2_bal_mast <- map2(fem_m2_bal_seu, de_idents, ~SetIdent(.x, value=.y))

```

```{r, message=FALSE, warning=FALSE}
saveRDS(fem_m2_bal_seu, 'data/processed/Pf/m2_all/fem_m2_bal_seu.RDS')

```

```{r, message=FALSE, warning=FALSE}
fem_m2_bal_seu <- readRDS('data/processed/Pf/m2_all/fem_m2_bal_seu.RDS')

```

```{r, message=FALSE, warning=FALSE}
fem_m2_bal_seu

```

```{r, message=FALSE, warning=FALSE}
##shortcut to run relevant donors subsets only
# m2_bal_mast_noM55M54 <- subset(m2_bal_mast[["gam"]], subset = donor %in% c("MSC68", "MSC66", "MSC70", "MSC60", "MSC49"))
```


```{r, message=FALSE, warning=FALSE, eval=F}
##!! NB Rerun only when the DE inputs change - takes looooonng

## Mast DE of dataset with male low (ml) included - shouldnt have any contribution as each stage handled differently
## Including lab and field strain DE
## !!!min.cells.group added for mh in the strain_dset where there's only balancing in lab VS field - we need to remove later

pts <- "slingPseudotime_1"
pts_all<- c(rep(pts,1))
pts_all<- pts

### !!! NOTE - RUNNING FOR NECESSARY DONORS, WILL REVERT LATER
fem_pairwise_mkrs_seu_clustr <- list(fem_m2_bal_seu, pts, de_idents, names(fem_m2_bal_seu)) %>%
# fem_pairwise_mkrs_seu_clustr <- list(list(m2_bal_mast_noM55M54), pts, de_idents, names(m2_bal_mast)) %>%
  pmap(.,  ~{
    # ide = ..3
    # a = combn(levels(droplevels(..1@meta.data[,..3])), 2,simplify = F) 
    a = combn(unique(..1@meta.data[,..3]), 2,simplify = F) 
    obj = ..1
    pseudos = ..2
    
    print(class(obj))
    print(paste("running ...", ..4))
    map(a, ~print(paste(.x[1], "n ", .x[2])))
    
    # map(a, ~FindMarkers(obj, ident.1 = .x[1], ident.2 = .x[2], latent.vars = pseudos, logfc.threshold = 0, test.use = 'MAST', min.cells.group = 1) %>% rownames_to_column('gene_id'))
    map(a, ~FindMarkers(obj, 
                        ident.1 = .x[1], 
                        ident.2 = .x[2], 
                        # latent.vars = c("donor", pseudos), 
                        # latent.vars = c(pseudos), 
                        logfc.threshold = 0, 
                        # test.use = 'MAST', 
                        min.cells.group = 1) %>% 
          rownames_to_column('gene_id')) %>%
      set_names(map(a, ~paste(., collapse = '_vs_') %>% str_remove_all(' '))) %>%
      bind_rows(., .id = 'pw_comparison')
    
  }#, 'missing genes')
  )

saveRDS(fem_pairwise_mkrs_seu_clustr,'data/processed/Pf/m2_all/fem_pairwise_mkrs_seu_clustr.RDS')
# saveRDS(fem_pairwise_mkrs_seu_clustr_don,'data/processed/Pf/m2_all/fem_pairwise_mkrs_seu_clustr_don.RDS')
# c(fem_pairwise_mkrs_seu_clustr_lf, fem_pairwise_mkrs_seu_clustr)



```



```{r, warning=FALSE, message=FALSE, eval=TRUE}
fem_pairwise_mkrs_seu_clustr <- readRDS('data/processed/Pf/m2_all/fem_pairwise_mkrs_seu_clustr.RDS')

```

```{r, warning=FALSE, message=FALSE}
fem_pairwise_mkrs_seu_clustr_anot <- map(fem_pairwise_mkrs_seu_clustr, ~.x %>%  left_join(., Pf3D7_genes_plasmoDB[,c("gene_id", "Gene")]))

fem_pairwise_mkrs_seu_clustr_anot
```

```{r, warning=FALSE, message=FALSE}
## cells that contain 2 or more gam UMI
map(fem_pairwise_mkrs_seu_clustr_anot, ~.x  %>% filter(p_val_adj < 0.05 & avg_log2FC > 0))

## https://www-nature-com.ezp.lib.cam.ac.uk/articles/nature24280.pdf - PF3D7-0406500 and HDA1 are high in gam positive cells

```

```{r, warning=FALSE, message=FALSE}
## cells that contain 2 or more gam UMI
imap(fem_pairwise_mkrs_seu_clustr_anot, ~.x %>% filter(p_val_adj < 0.05 & avg_log2FC < 0))

## https://www-nature-com.ezp.lib.cam.ac.uk/articles/nature24280.pdf - PF3D7-0406500 and HDA1 are high in gam positive cells

```


```{r, warning=FALSE, message=FALSE}
fem_pairwise_mkrs_seu_clustr_anot_all <- fem_pairwise_mkrs_seu_clustr_anot %>% 
  bind_rows(., .id = "sample") %>%
  filter(p_val_adj < 0.05) %>%
  group_by(gene_id) %>%
  mutate(donor = str_remove(sample, "_.*"),
         gene_DE_sample_prevalence = n_distinct(sample, pw_comparison),
         gene_DE_donor_prevalence = n_distinct(donor)) %>%
  ungroup() %>%
  mutate(gene_DE_sample_prevalence_cat = cut(gene_DE_sample_prevalence, breaks = c(3), labels = c("Low", "Medium", "High"), right = FALSE), 
         gene_DE_donor_prevalence_cat = cut(gene_DE_donor_prevalence, breaks = c(3), labels = c("Low", "Medium", "High"), right = FALSE))  %>% 
  filter(!str_detect(Gene, "VAR_"))
```


```{r, warning=FALSE, message=FALSE}
write_csv(fem_pairwise_mkrs_seu_clustr_anot_all, "../colleagues/chiyun/fem_pairwise_mkrs_seu_clustr_anot_all.csv")
```


```{r, warning=FALSE, message=FALSE}
fem_pairwise_mkrs_seu_clustr_anot_all  %>%
  arrange(desc(gene_DE_donor_prevalence), gene_id) %>%
  filter(gene_DE_donor_prevalence >= 5) %>%
  distinct(Gene, .keep_all = T)
```

```{r, warning=FALSE, message=FALSE}
fem_pairwise_mkrs_seu_clustr_anot_all  %>%
  arrange(desc(gene_DE_donor_prevalence), gene_id) %>%
  filter(gene_DE_donor_prevalence >= 5) %>%
  distinct(Gene, .keep_all = T)
```


```{r, warning=FALSE, message=FALSE}
fem_pairwise_mkrs_seu_clustr_anot_all  %>%
  filter(p_val_adj < 1e-5) %>% 
  ggplot(aes(x = avg_log2FC)) +
  geom_histogram(binwidth = 0.08) + 
  geom_vline(xintercept = c(-3, -2, -1, 1, 2, 3))
```

```{r, warning=FALSE, message=FALSE}
fem_pairwise_mkrs_seu_clustr_anot_all  %>%
  filter(p_val_adj < 1e-10) %>% 
  ggplot(aes(y = avg_log2FC, x = sample)) +
  geom_violin() + 
  geom_hline(yintercept = c(-3, -2, -1, 1, 2, 3))
```

```{r, warning=FALSE, message=FALSE, fig.width=5, fig.height=5}
fem_pairwise_mkrs_seu_clustr_anot_all  %>%
  filter(p_val_adj < 1e-10 ) %>% 
  ggplot(aes(y = avg_log2FC, x = 1)) +
  geom_violin() +
  geom_sina(size = 0.1) + 
  geom_hline(yintercept = seq(-8,8,1)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        text = element_text(size = 28))
```

```{r, warning=FALSE, message=FALSE, fig.width=7, fig.height=5}
fem_pairwise_mkrs_seu_clustr_anot_all  %>%
  filter(p_val_adj < 1e-10 ) %>% 
  ungroup() %>%
  ggplot(aes(y = avg_log2FC, x = gene_DE_donor_prevalence_cat, colour = gene_DE_donor_prevalence_cat)) +
  geom_violin() +
  geom_sina(size = 0.1) + 
  geom_hline(yintercept = seq(-8,8,1)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```


##############################END Female DE ##########################



```{r, warning=FALSE, message=FALSE}
gogo()
```

