---
title: "MSC QC and preprocessing"
date: '2022-07-11'
author: 'Jesse Rop'
output:
  html_document:
    toc: true
    theme: united
    code_folding: hide
    df_print: paged
---

```{r}
suppressPackageStartupMessages({
  library(Seurat)
  library(sctransform)
  # library(rgl)
  library(patchwork)
  # library(DoubletFinder)
  library(ggrepel)
  library(ggpubr)
  library(dplyr) 
  library(stringr)
  library(purrr)
  library(readr)
  library(tibble)
  library(ggplot2)
  library(tidyr)
  library(cowplot)
  # library(scDblFinder)
  library(SeuratData)
  library(SeuratDisk)
  library(plotly)
  library(Hmisc)
  library(rmarkdown)
  library(RColorBrewer)
  # library(Platypus)
  library(ggvenn)
  library(conflicted)
  # library(SoupX)
  library(MetBrewer)
  library(ggh4x)
  library(SingleCellExperiment)
  # library(scmap)
  # library(ggsankey)
  # library(findPC)
  library(BPCells)
  # library(furrr)
  # library(future.batchtools)
  library(parallel)
  library(EnhancedVolcano)
  library(DropletUtils)
  library(ggforce)
  library(slingshot)
  
  conflict_prefer("select", "dplyr")
  conflict_prefer("filter", "dplyr")
  conflict_prefer("rename", "dplyr")
  conflict_prefer("map", "purrr")
  conflict_prefer("count", "dplyr")
  conflicts_prefer(base::saveRDS)
  conflicts_prefer(SingleCellExperiment::colData)
  conflicts_prefer(base::as.data.frame)
})
```







# MSC Preprocessing


## setup, read input files and exploratory visualizations and general functions


```{r, setup}
# library(reticulate)
# # use_condaenv("/Users/jr35/envs/scv_jupcon/", required = TRUE)
# # matplotlib <- reticulate::import("matplotlib")
# # matplotlib$use("Agg", force = TRUE)
# 
# knitr::opts_knit$set(root.dir = '/lustre/scratch126/tol/teams/lawniczak/users/jr35/phd/Mali')
knitr::opts_knit$set(root.dir = '/lustre/scratch126/tol/teams/lawniczak/users/jr35/phd/Mali2')

```

### Common functions

#### Source commong functions, varriables and tables
```{r}
##Boxplot labeling function
source("/nfs/users/nfs_j/jr35/multipurpose_scripts/seurat_sce/seu_sce_common_fns.R")
source("/nfs/users/nfs_j/jr35/multipurpose_scripts/plotting_common_fns.R")
source("/nfs/users/nfs_j/jr35/Mali2/mali22_code/pf_common_vars.R")
```



########START TRY SCT ########
Integrated asexual to integrate with the rescued parasites
```{r}
## subset where field parasites are for slingshot
# v3_m2_asex_harmony_ss_res_fld_pt <- readRDS("data/processed/Pf/m2_all/v3_m2_asex_harmony_ss_res_fld_pt.RDS")
# v3_m2_asex_harmony_ss_res_fld_pt <- readRDS("data/processed/Pf/m2_all/v3_m2_asex_harmony_res_fld_pt.RDS")
v3_m2_asex_harmony_ss_res_fld_pt <- readRDS("data/processed/Pf/m2_all/all_msc_cb_no_dbt_mrg_noM12_asex.RDS")

# %in% c("MSC25" ,"MSC41", "MSC51")

```







```{r, warning=FALSE, message=FALSE}
# split layers
# v3_m2_asex_cr_filt[["RNA"]] <- split(v3_m2_asex_cr_filt[["RNA"]], f = v3_m2_asex_cr_filt$donor)

```


############################


```{r}
## read in 
# all_resc_asex_sct_ap2g <- readRDS("data/processed/Pf/m2_all/all_resc_asex_sct_ap2g.RDS")

```

Integrated asexual to integrate with the rescued parasites
```{r}
## subset where field parasites are for slingshot
# v3_m2_asex_harmony_ss_res_fld_pt <- readRDS("data/processed/Pf/m2_all/v3_m2_asex_harmony_ss_res_fld_pt.RDS")
# v3_m2_asex_harmony_ss_res_fld_pt <- readRDS("data/processed/Pf/m2_all/v3_m2_asex_harmony_res_fld_pt.RDS")
# %in% c("MSC25" ,"MSC41", "MSC51")

```


########START TRY SCT ########

Integrated asexual to integrate with the rescued parasites
```{r}
## subset where field parasites are for slingshot
# v3_m2_asex_harmony_ss_res_fld_pt <- readRDS("data/processed/Pf/m2_all/v3_m2_asex_harmony_ss_res_fld_pt.RDS")
# v3_m2_asex_harmony_ss_res_fld_pt <- readRDS("data/processed/Pf/m2_all/v3_m2_asex_harmony_res_fld_pt.RDS")
# %in% c("MSC25" ,"MSC41", "MSC51")

```

```{r, warning=FALSE, message=FALSE}
v3_m2_asex_harmony_ss_res_fld_pt@meta.data %>% count(sR_fine_stg)
```

```{r, warning=FALSE, message=FALSE}
v3_m2_asex_harmony_ss_res_fld_pt <- v3_m2_asex_harmony_ss_res_fld_pt@meta.data %>% 
  rownames_to_column('bcode') %>%
  mutate(dset_labmrg = "Field",
         stage = sR_fine_stg,
         donor = str_remove_all(orig.ident, ".mrna")) %>% 
  add_count(donor, name = "don_count") %>%
  select(bcode, dset_labmrg, donor, don_count, stage) %>%
  column_to_rownames('bcode') %>%
  AddMetaData(v3_m2_asex_harmony_ss_res_fld_pt, .)
  
```



```{r, warning=FALSE, message=FALSE}
v3_m2_asex_harmony_ss_res_fld_pt@meta.data %>% dplyr::count(donor, dset_labmrg)
```

```{r}
## remove lab and M12
v3_m2_asex_cr_filt <- v3_m2_asex_harmony_ss_res_fld_pt[, !v3_m2_asex_harmony_ss_res_fld_pt@meta.data$dset_labmrg %in% c("Lab", "MSC12")]
```

```{r, warning=FALSE, message=FALSE}
v3_m2_asex_cr_filt@meta.data %>% count(donor, sR_fine_stg) %>% group_by(donor) %>% mutate(prop = round(n/sum(n), 3)) %>% filter(sR_fine_stg == "Gametocyte (developing)")
```

```{r, warning=FALSE, message=FALSE}
v3_m2_asex_cr_filt@meta.data %>% add_count(sR_fine_stg, name = "stg_n") %>% group_by() %>% dplyr::count(donor)
```

```{r, warning=FALSE, message=FALSE}
v3_m2_asex_cr_filt@meta.data %>% filter(sR_fine_stg == "Gametocyte (developing)") %>% dplyr::count(donor)
```

```{r, eval=T}
## Making list of objects and renaming before integrating
# v3_m2_asex_cr_filt <- subset(v3_m2_asex_cr_filt, subset = donor != "MSC55")

```

```{r, warning=FALSE, message=FALSE}
v3_m2_asex_cr_filt@meta.data %>% dplyr::count(donor, dset_labmrg)
```

```{r, warning=FALSE, message=FALSE}
v3_m2_asex_cr_filt@meta.data %>% add_count(donor) %>% dplyr::filter( n < 100)
```

```{r}
## remove weird 2 cells with >9000 ncountRNA
VlnPlot(v3_m2_asex_cr_filt, features = "nCount_RNA") + geom_hline(yintercept = 9000)
```

```{r}
## remove weird 2 cells with >9000 ncountRNA
table(v3_m2_asex_cr_filt@meta.data$nCount_RNA >9000)
v3_m2_asex_cr_filt <- v3_m2_asex_cr_filt[, !v3_m2_asex_cr_filt@meta.data$nCount_RNA >9000]
```

```{r}
## remove any donor with less than 100 cells
v3_m2_asex_cr_filt <- v3_m2_asex_cr_filt[, !v3_m2_asex_cr_filt@meta.data$don_count < 150]
```


```{r}
## remove any donor with less than 100 cells
saveRDS(v3_m2_asex_cr_filt, "data/processed/Pf/m2_all/v3_m2_asex_cr_filt.RDS")
```

```{r, warning=FALSE, message=FALSE}
# split layers
# v3_m2_asex_cr_filt[["RNA"]] <- split(v3_m2_asex_cr_filt[["RNA"]], f = v3_m2_asex_cr_filt$donor)

```


```{r, warning=FALSE, message=FALSE}
# run sctransform
v3_m2_asex_cr_filt_sct <- SCTransform(v3_m2_asex_cr_filt, verbose = FALSE, variable.features.n = 400)
# v3_m2_asex_cr_filt_sct <- SCTransform(v3_m2_asex_cr_filt, verbose = FALSE, variable.features.n = 300)
```

```{r, warning=FALSE, message=FALSE}
# These are now standard steps in the Seurat workflow for visualization and clustering
v3_m2_asex_cr_filt_sct <- RunPCA(v3_m2_asex_cr_filt_sct, verbose = FALSE, npcs = 20)

```

```{r, warning=FALSE, message=FALSE}
# run sctransform
DimPlot(v3_m2_asex_cr_filt_sct, reduction = "pca")

```

```{r, warning=FALSE, message=FALSE}
# Error in getGlobalsAndPackages(expr, envir = envir, globals = globals) : 
options(future.globals.maxSize = 1000 * 1024^2)  # 1000 MiB (or 1 GiB)

v3_m2_asex_cr_filt_sct_int <- IntegrateLayers(
  object = v3_m2_asex_cr_filt_sct,
  method = RPCAIntegration,
  normalization.method = "SCT",
  verbose = F,
  # k.weight = 90,
)


```

```{r, warning=FALSE, message=FALSE}

v3_m2_asex_cr_filt_sct_int <- FindNeighbors(v3_m2_asex_cr_filt_sct_int, dims = 1:5, reduction = "integrated.dr")
v3_m2_asex_cr_filt_sct_int <- FindClusters(v3_m2_asex_cr_filt_sct_int, resolution = 0.5)
```

```{r, warning=FALSE, message=FALSE}
# These are now standard steps in the Seurat workflow for visualization and clustering
v3_m2_asex_cr_filt_sct_int <- RunUMAP(v3_m2_asex_cr_filt_sct_int, dims = 1:5, verbose = FALSE, n.components = 3, seed.use = 867)
v3_m2_asex_cr_filt_sct_int <- FindNeighbors(v3_m2_asex_cr_filt_sct_int, dims = 1:5, verbose = FALSE)
v3_m2_asex_cr_filt_sct_int <- FindClusters(v3_m2_asex_cr_filt_sct_int, verbose = FALSE, resolution = 0.1)
```

```{r, warning=FALSE, message=FALSE}
# saveRDS(v3_m2_asex_cr_filt_sct_int, "data/processed/Pf/m2_all/v3_m2_asex_cr_filt_sct_int.RDS")
```

```{r, warning=FALSE, message=FALSE}
v3_m2_asex_cr_filt_sct_int <- readRDS("data/processed/Pf/m2_all/v3_m2_asex_cr_filt_sct_int.RDS")
```


```{r, warning=FALSE, message=FALSE}
DimPlot(v3_m2_asex_cr_filt_sct_int, label = T, reduction = "pca")
DimPlot(v3_m2_asex_cr_filt_sct_int, label = T)
DimPlot(v3_m2_asex_cr_filt_sct_int, label = T, group.by = "stage") + scale_color_manual(values = sp_fld_colors)
DimPlot(v3_m2_asex_cr_filt_sct_int, group.by = "donor")
FeaturePlot(v3_m2_asex_cr_filt_sct_int, features = "nFeature_RNA")
FeaturePlot(v3_m2_asex_cr_filt_sct_int, features = "nCount_RNA")
VlnPlot(v3_m2_asex_cr_filt_sct_int, features = "nCount_RNA", group.by = "seurat_clusters")

```


```{r, warning=FALSE, message=FALSE, eval=T}
# Display the combined datasets as an interactive 3D plot

plot_ly(data = data.frame(v3_m2_asex_cr_filt_sct_int@reductions[["umap"]]@cell.embeddings),
        x = ~umap_1, y = ~umap_2, z = ~umap_3,
        color= v3_m2_asex_cr_filt_sct_int@meta.data$sR_fine_stg,
        # colors = c('blue', 'red'),
        opacity = 0.8,
        size = 0.1, 
        type = 'scatter3d'
)

```

```{r, warning=FALSE, message=FALSE}
DimPlot(v3_m2_asex_cr_filt_sct_int, label = T, group.by = "stage", dims = c(1,3))
DimPlot(v3_m2_asex_cr_filt_sct_int, label = T, group.by = "stage", dims = c(2,3))

```

```{r, warning=FALSE, message=FALSE}
DimPlot(v3_m2_asex_cr_filt_sct_int, label = T, group.by = "stage", dims = c(1,3))
DimPlot(v3_m2_asex_cr_filt_sct_int, label = T, dims = c(1,3))

```


```{r, warning=FALSE, message=FALSE}
DimPlot(v3_m2_asex_cr_filt_sct_int, label = T, group.by = "stage", dims = c(1,3))
DimPlot(v3_m2_asex_cr_filt_sct_int, label = T, dims = c(1,3))

```

```{r, warning=FALSE, message=FALSE}
## ap2g
DefaultAssay(v3_m2_asex_cr_filt_sct_int) <- "RNA"
v3_m2_asex_cr_filt_sct_int <- NormalizeData(v3_m2_asex_cr_filt_sct_int) 
```

```{r, warning=FALSE, message=FALSE}
## ap2g
FeaturePlot(v3_m2_asex_cr_filt_sct_int, features = "PF3D7-1222600", order = T) 
FeaturePlot(v3_m2_asex_cr_filt_sct_int, features = "PF3D7-1222600", reduction = "pca", order = T) 
```


```{r, warning=FALSE, message=FALSE}
## ap2g
v3_m2_asex_cr_filt_sct_int_ap2g <- subset(v3_m2_asex_cr_filt_sct_int, subset = `PF3D7-1222600` > 0, slot = 'counts')

table(v3_m2_asex_cr_filt_sct_int_ap2g$donor)
```

```{r, warning=FALSE, message=FALSE}
## ap2g
FeaturePlot(v3_m2_asex_cr_filt_sct_int, features = "PF3D7-1215500", order = T) 
FeaturePlot(v3_m2_asex_cr_filt_sct_int, features = "PF3D7-1215500", reduction = "pca", order = T) 
```


```{r, warning=FALSE, message=FALSE}

DefaultAssay(v3_m2_asex_cr_filt_sct_int) <- "SCT"
```


```{r, eval=T}
gogo()
# bsub < /lustre/scratch126/tol/teams/lawniczak/users/jr35/phd/Mali2/mali22_code_lnk/bash_scripts/submsn_seu_norm_integration_bsub.sh

```

```{r, eval=T}
# bsub < /lustre/scratch126/tol/teams/lawniczak/users/jr35/phd/Mali2/mali22_code_lnk/bash_scripts/submsn_intgrtn_pc_harmony_bsub.sh

```

```{r, warning=FALSE, message=FALSE}
# v3_m2_asex_cr_filt_norm_harmony <- readRDS("data/processed/Pf/m2_all/v3_m2_asex_cr_filt_norm_harmony.RDS")
v3_m2_asex_cr_filt_norm_harmony <- readRDS("data/processed/Pf/m2_all/v3_m2_asex_cr_filt_norm_harmony_pc_man.RDS")

```

```{r, warning=FALSE, message=FALSE}
v3_m2_asex_cr_filt_norm_harmony_elbo <- readRDS("data/processed/Pf/m2_all/v3_m2_asex_cr_filt_norm_harmony_elbo.RDS")
v3_m2_asex_cr_filt_norm_harmony_elbo
```


```{r, warning=FALSE, message=FALSE}
DimPlot(v3_m2_asex_cr_filt_norm_harmony, label = T, reduction = "integrated.harmony_pc_man")
DimPlot(v3_m2_asex_cr_filt_norm_harmony, label = T, reduction = "umap.harmony_pc_man")
DimPlot(v3_m2_asex_cr_filt_norm_harmony, label = T, group.by = "stage", reduction = "umap.harmony_pc_man") + scale_color_manual(values = sp_fld_colors)
DimPlot(v3_m2_asex_cr_filt_norm_harmony, group.by = "donor", reduction = "umap.harmony_pc_man")
FeaturePlot(v3_m2_asex_cr_filt_norm_harmony, features = "nFeature_RNA", reduction = "umap.harmony_pc_man")
FeaturePlot(v3_m2_asex_cr_filt_norm_harmony, features = "nCount_RNA", reduction = "umap.harmony_pc_man")
# VlnPlot(v3_m2_asex_cr_filt_norm_harmony, features = "nCount_RNA", group.by = "seurat_clusters", reduction = "umap.harmony_pc_man")

```


```{r, warning=FALSE, message=FALSE}
DimPlot(v3_m2_asex_cr_filt_norm_harmony, label = T, group.by = "stage", reduction = "umap.harmony_pc_man") + scale_color_manual(values = sp_col_new)

```

```{r, warning=FALSE, message=FALSE, fig.width= 15, fig.height=4}
DimPlot(v3_m2_asex_cr_filt_norm_harmony, label = T, split.by = "stage", reduction = "umap.harmony_pc_man", pt.size = 0.0001) 

```

```{r, warning=FALSE, message=FALSE, fig.width=12, fig.height=3}
DimPlot(v3_m2_asex_cr_filt_norm_harmony, label = T, reduction = "umap.harmony_pc_man", split.by = 'donor')

```

```{r, warning=FALSE, message=FALSE, eval=T}
# Display the combined datasets as an interactive 3D plot

plot_ly(data = data.frame(v3_m2_asex_cr_filt_norm_harmony@reductions[["umap.harmony_pc_man"]]@cell.embeddings),
        x = ~umapharmonypcman_1, y = ~umapharmonypcman_2, z = ~umapharmonypcman_3,
        color= v3_m2_asex_cr_filt_norm_harmony@meta.data$sR_fine_stg,
        # colors = c('blue', 'red'),
        opacity = 0.8,
        size = 0.1, 
        type = 'scatter3d'
)

```


############# START find all markers to test for merozoites #############

```{r, warning=FALSE, message=FALSE}
DimPlot(v3_m2_asex_cr_filt_norm_harmony, label = T, reduction = "umap.harmony_pc_man", group.by = "harmony_pc_man_clusters")

```

```{r, warning=FALSE, message=FALSE}
DimPlot(v3_m2_asex_cr_filt_norm_harmony, label = T, reduction = "umap.harmony_pc_man", group.by = "harmony_pc_man_clusters", dims = c(1,3))

```

```{r, warning=FALSE, message=FALSE}
DimPlot(v3_m2_asex_cr_filt_norm_harmony, label = T, reduction = "integrated.harmony_pc_man", group.by = "harmony_pc_man_clusters", dims = c(2,3))

```



```{r, warning=FALSE, message=FALSE}
v3_m2_asex_cr_filt_norm_harmony_jnd <- JoinLayers(v3_m2_asex_cr_filt_norm_harmony)
DefaultAssay(v3_m2_asex_cr_filt_norm_harmony_jnd)
unique(Idents(v3_m2_asex_cr_filt_norm_harmony_jnd))
```


```{r, warning=FALSE, message=FALSE}
gogo()
```

```{r, warning=FALSE, message=FALSE}
rm(v3_m2_asex_cr_filt_norm_harmony)
```


```{r, warning=FALSE, message=FALSE}
v3_m2_asex_cr_filt_norm_harmony_jnd_small <- v3_m2_asex_cr_filt_norm_harmony_jnd[,1:10]
```

```{r, warning=FALSE, message=FALSE}
v3_m2_asex_cr_filt_norm_harmony_jnd_small[["RNA"]]$data <- as(v3_m2_asex_cr_filt_norm_harmony_jnd_small[["RNA"]]$data, Class = "dgCMatrix")
```

```{r, warning=FALSE, message=FALSE}
saveRDS(v3_m2_asex_cr_filt_norm_harmony_jnd, "data/processed/Pf/m2_all/v3_m2_asex_cr_filt_norm_harmony_jnd.RDS")
```
