---
title: "MSC QC and preprocessing"
date: '2022-07-11'
author: 'Jesse Rop'
output:
  html_document:
    toc: true
    theme: united
    code_folding: hide
    df_print: paged
---

```{r}
suppressPackageStartupMessages({
  library(Seurat)
  library(sctransform)
  # library(rgl)
  library(patchwork)
  # library(DoubletFinder)
  library(ggrepel)
  library(ggpubr)
  library(dplyr) 
  library(stringr)
  library(purrr)
  library(readr)
  library(tibble)
  library(ggplot2)
  library(tidyr)
  library(cowplot)
  library(scDblFinder)
  library(SeuratData)
  library(SeuratDisk)
  library(plotly)
  library(Hmisc)
  library(rmarkdown)
  library(RColorBrewer)
  # library(Platypus)
  library(ggvenn)
  library(conflicted)
  library(SoupX)
  library(MetBrewer)
  library(ggh4x)
  library(SingleCellExperiment)
  library(scmap)
  # library(ggsankey)
  # library(findPC)
  library(BPCells)
  library(furrr)
  library(future.batchtools)
  library(parallel)
  
  conflict_prefer("select", "dplyr")
  conflict_prefer("filter", "dplyr")
  conflict_prefer("rename", "dplyr")
  conflict_prefer("map", "purrr")
  conflict_prefer("count", "dplyr")
  conflicts_prefer(base::saveRDS)
  conflicts_prefer(SingleCellExperiment::colData)
})
```


```{r, setup}

# knitr::opts_knit$set(root.dir = '/Users/jr35/Google Drive/My Drive/PHD_documents/malaria_phd/Mali2')
knitr::opts_knit$set(root.dir = '/lustre/scratch126/tol/teams/lawniczak/users/jr35/phd/Mali2/')
```


Copy the irods metadata file from local into farm
```{r, eval=FALSE}
system(
  paste0('rsync -av /Users/jr35/Google\\ Drive/My\\ Drive/PHD_documents/malaria_phd/Mali2/data/raw/irods_id_sk21_mdata_cln.csv jr35@farm22:/lustre/scratch126/tol/teams/lawniczak/users/jr35/phd/Mali2/data/raw/irods_id_sk21_mdata_cln.csv')
)
```


### Common functions

#### Source commong functions, varriables and tables
```{r}
##Boxplot labeling function
source("/lustre/scratch126/tol/teams/lawniczak/users/jr35/phd/multipurpose_scripts/seurat_sce/seu_sce_common_fns.R")
source("/lustre/scratch126/tol/teams/lawniczak/users/jr35/phd/multipurpose_scripts/plotting_common_fns.R")
source("/lustre/scratch126/tol/teams/lawniczak/users/jr35/phd/Mali2/mali22_code/pf_common_vars.R")
```


initialize variable for the sample names

```{r}
irods_id_sk21_mdata_cln <- read_csv("data/raw/irods_id_sk21_mdata_cln.csv") %>%
  mutate(sample_nm = str_remove_all(sample_name, "_JC")) %>%
  arrange(sample_nm)

smpl_nms <-irods_id_sk21_mdata_cln$sample_nm
#c("MSC24", "MSC33", "MSC37", "MSC38", "MSC39", "MSC40", "MSC45", "MSC48", "MSC49", "MSC50_MACS", "MSC50_SLO", "MSC53", "MSC54", "MSC55", "MSC57", "MSC60", "MSC66", "MSC67", "MSC68", "MSC70")
```

## Genotype clustering for MSC

Read in clusters.tsv files with assignment of cells to strains and PCA plots metrics

Transfer clusters.tsv files with assignment of cells to strains and PCA plots metrics for the best K from farm to my local Mac

```{r}
##Copy souporcell files for the optimum clusters

##Sample ID
sample_name <- irods_id_sk21_mdata_cln$sample_nm

##Sample IDs
source_irods <- irods_id_sk21_mdata_cln[,c('sample_nm', 'irods_id')] %>% deframe

```


NB-FARM_CODE
```{r, eval=F}
##Copy h5 files from cellranger and making relevant directories

for (i in 1:length(sample_name)) {
  
  ## Create new directories
  system(paste0('mkdir -p data/raw/Pf/', sample_name[i], '/outs/filtered_feature_bc_matrix'))
  system(paste0('mkdir -p data/processed/Pf/', sample_name[i], '/bpcells'))
  
  ##Transfers
  ##souporcell genotyping files
  system(
    paste0('cp /lustre/scratch126/tol/teams/lawniczak/projects/malaria_single_cell/mali_field_runs/2022/data/cellranger_runs/Pf/', source_irods[i], '/outs/filtered_feature_bc_matrix.h5 data/raw/Pf/', sample_name[i], '/outs/')
  )
  system(
    paste0('cp /lustre/scratch126/tol/teams/lawniczak/projects/malaria_single_cell/mali_field_runs/2022/data/cellranger_runs/Pf/', source_irods[i], '/outs/filtered_feature_bc_matrix/barcodes.tsv.gz data/raw/Pf/', sample_name[i], '/outs/filtered_feature_bc_matrix/')
  )
  
}


```


```{r}
## List filtered cellranger h5 matrices into R
list.files("data/raw/Pf/", pattern = "filtered_feature_bc_matrix.h5$", full.names = T, recursive = T, include.dirs = T)
```


```{r, message=F, warning=F, eval=F}
##Write BPcells matrices
list.files("data/raw/Pf/", pattern = "filtered_feature_bc_matrix.h5$", full.names = T, recursive = T, include.dirs = T) %>% 
  set_names(nm = (str_remove_all(.,'data/raw/Pf//|/outs.*'))) %>% 
  .[sample_name] %>%
  map(~open_matrix_10x_hdf5(path = .)) %>% 
  imap(., ~write_matrix_dir(mat = .x, dir = paste0('data/processed/Pf/', .y, '/bpcells/counts_mtx'), overwrite = T))
```


```{r, message=F, warning=F, eval=F}
##Read BPcells matrices
all_msc_bp <- list.files("data/processed/", pattern = "counts_mtx$", full.names = T, recursive = T, include.dirs = T) %>% 
  set_names(nm = (str_remove_all(.,'data/processed//Pf/|/bpcells.*'))) %>% 
  .[sample_name] %>%
  map(~open_matrix_dir(dir = .)) %>% 
  imap(., ~CreateSeuratObject(counts = .x, project = paste0(.y, ".mrna"))) 
```


```{r, message=F, warning=F, eval=F}
# Convert to dense object - might be necessary in future instances
all_msc <- all_msc_bp %>%
  map(., ~{
    .x[["RNA"]]$counts <- as(object = .x[["RNA"]]$counts, Class = "dgCMatrix")
    return(.x)
  })

```


```{r, eval=F}
##write out for normalization processing in farm

for (i in 1:length(sample_name)) {
  
  ## Create new directories
  system(paste0('mkdir -p data/processed/Pf/', sample_name[i], '/seu_obj'))
  
}


```

FARM CODE
```{r, eval=F}
## write out seurat object files for normalization processing in farm
imap(all_msc, ~saveRDS(.x, paste0("data/processed/Pf/",.y,"/seu_obj/seu.RDS")))
```

```{r, eval=F}
## write out seurat bpcells object files for normalization processing in farm
imap(all_msc_bp, ~saveRDS(.x, paste0("data/processed/Pf/",.y,"/seu_obj/bpc_seu.RDS")))
```

!!!NB - FARM CODE - DONT DELETE
```{r, eval=F}
## Run SCT on farm like below
# nextflow run submsn_seu_norm.nf --scrpt "/lustre/scratch126/tol/teams/lawniczak/users/jr35/phd/multipurpose_scripts/seurat_sce/seu_stdnorm.R" --input_seu "/lustre/scratch126/tol/teams/lawniczak/users/jr35/phd/Mali2/data/processed/Pf/MSC*/seu_obj/bpc_seu.RDS" --o_file_name "seu_norm.RDS" --umap_name "UMAP_" --hvg 500 --cluster_resoln 0.1
```

Read in seurat objects processed in farm
```{r, message=F, warning=F}
##Read BPcells matrices
all_msc <- list.files("data/processed/", pattern = "^seu_norm.RDS$", full.names = T, recursive = T, include.dirs = T) %>% 
  set_names(nm = (str_remove_all(.,'data/processed//Pf/|/seu_obj.*'))) %>% 
  # .[sample_name] %>%
  map(readRDS)
```


```{r, warning=FALSE, message=FALSE}
imap(all_msc, ~DimPlot(.x, dims = c(1,2), label = T) + labs(title = .y))
```


## scmap annotation

### Preprocessing references

#### msc_combi reference dataset
```{r, eval=FALSE}
##Read MCA V3 blood stage reference dataset
# msc_combi <- readRDS("../sc_velo/data/raw/RoadtoUMAP/pseudotime/raw_combined_cut_0.4.RDS")
# msc_combi <- readRDS("../Mali/data/raw/mca_v3_ref/raw_combined_cut_0.4_nodoublets_stagenew.RDS")
msc_combi <- readRDS("../Mali/data/processed/DB52e_star/msc_combi.RDS")
DefaultAssay(msc_combi) <- "RNA"

##Collapse asexual stages into 1 "Asexual"
msc_combi@meta.data$stageHL_ref_asx <- ifelse(msc_combi@meta.data$stageHL_ref %in% c("Early trophozoite", "Late ring"), "Asexual", msc_combi@meta.data$stageHL_ref)
```

```{r, message=FALSE, warning=FALSE, eval=FALSE}
##Convert to reference to sce
msc_combi.sce <- as.SingleCellExperiment(msc_combi)

##Place counts appropriately into single cell experiment object
SingleCellExperiment::logcounts(msc_combi.sce) <- log2(SingleCellExperiment::counts(msc_combi.sce) + 1)
rowData(msc_combi.sce)$feature_symbol <- rownames(msc_combi.sce)
msc_combi.sce <- msc_combi.sce[!duplicated(rownames(msc_combi.sce)), ]

##scmap feature selection
msc_combi.sce <- selectFeatures(msc_combi.sce, suppress_plot = T)

##Cell indexing
set.seed(155)
msc_combi.sce <- indexCell(msc_combi.sce)
```

NB-FARM_CODE
```{r, eval=F}
## transfer integrated object from sunil farm to local
# system('cp /lustre/scratch126/tol/teams/lawniczak/projects/malaria_single_cell/mali_field_runs/2022/data/rdata/for_scmap/v3labmscs.ref.seur.RDS data/raw/Pf/')
# system('cp /lustre/scratch126/tol/teams/lawniczak/projects/malaria_single_cell/mali_field_runs/2022/data/rdata/for_scmap/v3labmscs.ref.sce.RDS data/raw/Pf/')

```

#### full msc_atlas reference dataset
```{r}
##Read MCA V3 + field blood stage reference dataset
# v3labmscs.ref.sce <- readRDS("../Mali2/data/raw/Pf/v3labmscs.ref.sce.RDS")
v3labmscs.ref.sce <- readRDS("data/raw/Pf/v3labmscs.ref.sce.RDS")
```



```{r, message=FALSE, warning=FALSE}
##Collapse asexual stages into 1 "Asexual"
v3labmscs.ref.sce@colData[,c("stageHL_ref", "stageHL_ref_asx")] <- colData(v3labmscs.ref.sce) %>% 
  data.frame() %>% 
  mutate(stageHL_ref = case_when(str_detect(orig.ident, "msc") & stageLR == "late ring" ~ "Late ring", 
                                 str_detect(orig.ident, "msc") & stageLR == "early trophozoite" ~ "Early trophozoite", 
                                 str_detect(orig.ident, "msc") & stageLR == "gametocyte (female) LE" ~ "Female LE", 
                                 str_detect(orig.ident, "msc") & stageLR == "gametocyte (female)" ~ "Female", 
                                 str_detect(orig.ident, "msc") & stageLR == "gametocyte (male) LE" ~ "Male LE", 
                                 str_detect(orig.ident, "msc") & stageLR == "gametocyte (male)" ~ "Male", T ~ stageLR)) %>% 
  mutate(stageHL_ref_asx = case_when(stageHL_ref %in% c("Early trophozoite", "Late ring") ~ "Asexual", T ~ stageHL_ref)) %>%
  select(stageHL_ref, stageHL_ref_asx)

```


```{r, message=FALSE, warning=FALSE}
scater::plotReducedDim(v3labmscs.ref.sce, dimred = "PCA", colour_by = "stageHL_ref_asx", ncomponents = c(2,3))
```


```{r, message=FALSE, warning=FALSE, eval = F}
saveRDS(v3labmscs.ref.sce, "data/processed/Pf/v3labmscs.ref.sce.stgasx.RDS")
```

```{r, eval=T}
jkjkj()
## Run scmap on farm like below
# nextflow run submsn_scmap.nf --scrpt "/lustre/scratch126/tol/teams/lawniczak/users/jr35/phd/multipurpose_scripts/seurat_sce/scmap.R" --sce_ref "/lustre/scratch126/tol/teams/lawniczak/users/jr35/phd/Mali2/data/processed/Pf/v3labmscs.ref.sce.stgasx.RDS" --input_seu "/lustre/scratch126/tol/teams/lawniczak/users/jr35/phd/Mali2/data/processed/Pf/MSC*/seu_obj/seu_norm.RDS" --w 3 --cos_sim 0.5 --o_file "prelim_anotd.RDS"
```


```{r, message=F, warning=F}
##Read the annotations
all_msc_anots <- list.files("data/processed/", pattern = "prelim_anotd.RDS$", full.names = T, recursive = T, include.dirs = T) %>% 
  set_names(nm = (str_remove_all(.,'data/processed//Pf/|/seu_obj.*'))) %>% 
  .[sample_name] %>% 
  .[!is.na(.)] %>%
  map(readRDS)
```

```{r}
## Adding annotations to query cells
all_msc_anotd_preQC <- list(all_msc_anots, all_msc) %>%
  pmap(~ ..1 %>%
         column_to_rownames("cell_bc") %>%
         mutate(across(Stage_MSC, ~droplevels(factor(., levels = stg_refnd_lvls)))) %>%
         AddMetaData(..2, .) 
  )

```

```{r, warning=FALSE, message=FALSE}

imap(all_msc_anotd_preQC, ~DimPlot(.x, dims = c(1,2), label = T, group.by = 'Stage_MSC', cols = sp_fld_colors) + labs(title = .y))
```


```{r, warning=FALSE, message=FALSE}

imap(all_msc_anotd_preQC, ~DimPlot(.x, dims = c(1,2), label = T, group.by = 'stage_lab1', cols = sp_fld_colors)+ labs(title = .y))
```

```{r, warning=FALSE, message=FALSE}

map(all_msc_anotd_preQC, ~..1@meta.data %>% dplyr::count(Stage_MSC, stage_lab1))

```


```{r}
##Modify the stage labels for MSC14
# all_msc_anotd_preQC[[4]] <- all_msc_anotd_preQC[[4]]@meta.data %>%
all_msc_anotd_preQC <- pmap(list(all_msc_anotd_preQC, rep(0.4,length(sample_name)), c(rep('ring|trophozoite|developing', length(sample_name)))), ~{
  ..1@meta.data %>%
    mutate(across(Stage_MSC, as.character)) %>%
    # mutate(Stage_MSC = Stage_MSC) %>%
    mutate(StagePrelim_lf = case_when(
      is.na(Stage_MSC) & sim_lab1 > 0.35 & str_detect(stage_lab1, ..3) & str_detect(stage_lab2, ..3) & str_detect(stage_lab3, ..3) ~ stage_lab1, ## lower cos sim of 0.35 for asexuals instead of 0.4
      is.na(Stage_MSC) & sim_lab1 > ..2 & str_detect(stage_lab1, 'Male|male') & str_detect(stage_lab2, 'Male|male') & str_detect(stage_lab3, 'Male|male') ~ stage_lab1,
      is.na(Stage_MSC) & sim_lab1 > ..2 & str_detect(stage_lab1, 'Female|female') & str_detect(stage_lab2, 'Female|female') & str_detect(stage_lab3, 'Female|female') ~ stage_lab1,
      TRUE ~ Stage_MSC)#,
      # across(StagePrelim, str_to_sentence)
    )%>%
    mutate(StagePrelim = case_when(
      str_detect(StagePrelim_lf, "developing") ~ "Early trophozoite", ## !!NB - label developing gams as early trophs the closest stage
      str_detect(StagePrelim_lf, "schizont") & str_detect(stage_lab2, "ring|trophozoite")  & str_detect(stage_lab3, "ring|trophozoite") ~ str_to_sentence(stage_lab2),
      str_detect(StagePrelim_lf, "late|early") ~ str_to_sentence(StagePrelim_lf),
      str_detect(StagePrelim_lf, "gametocyte") ~ str_to_sentence(str_remove_all(StagePrelim_lf, "gametocyte \\(|\\)")),
      is.na(StagePrelim_lf) ~ "Unassigned",
      TRUE ~ StagePrelim_lf)#,
      # across(StagePrelim, str_to_sentence)
    ) %>%
    mutate(StagePrelimC = case_when(
      str_detect(StagePrelim, "ring|trophozoite|developing|schizont") ~ "Asexual",
      TRUE ~ StagePrelim)#,
      # across(StagePrelim, str_to_sentence)
    ) %>%
    mutate(across(c(StagePrelim, StagePrelimC), ~droplevels(factor(., levels = stg_refnd_lvls)))) %>%
    dplyr::select(StagePrelim, StagePrelim_lf, StagePrelimC) %>%
    # mutate(across(starts_with('Stage_V'), ~droplevels(factor(., levels = stg_refnd_lvls)))) %>%
    AddMetaData(.x, .)
})

```


```{r, warning=FALSE, message=FALSE}

# map(all_msc_anotd_preQC, ~.x@meta.data %>% dplyr::count(Stage_MSC, StagePrelim, stage_lab1))
map(all_msc_anotd_preQC, ~.x@meta.data %>% dplyr::count(StagePrelim, StagePrelimC))
```

```{r, warning=FALSE, message=FALSE}
imap(all_msc_anotd_preQC, ~DimPlot(.x, dims = c(1,2), label = F, group.by = 'StagePrelim', cols = sp_fld_colors)+ labs(title = .y))
```

```{r, warning=FALSE, message=FALSE}
imap(all_msc_anotd_preQC, ~FeaturePlot(.x, dims = c(1,2), features = "nFeature_RNA")+ labs(title = .y))
```

```{r, results='asis',  fig.width = 8, fig.height = 4}
##adding clusters that are used to display mitochondrial count

plts <- list(all_msc_anotd_preQC, names(all_msc_anotd_preQC)) %>%
  pmap(., ~{
    
    cell_qc_plot <-  ..1 %>% DimPlot(., pt.size = 1, reduction = 'umap', dims = c(1,2), group.by = 'StagePrelim', cols = sp_fld_colors) +labs(title = ..2)#,
    
    cell_qc_plot$data <- cell_qc_plot$data %>% rownames_to_column('bcode') %>% left_join(., ..1@meta.data[,'seurat_clusters', drop = F] %>% rownames_to_column('bcode'), by='bcode') %>% column_to_rownames('bcode')
    
    # print(str(cell_qc_plot$data))
    
    cell_qc_plot<- LabelClusters(cell_qc_plot, id = "seurat_clusters", size = 9, repel = T)
    
    return(cell_qc_plot)

  })
```


```{r, results='asis',  fig.width = 8, fig.height = 4}
plts

```

Umap with stage labels in legend

```{r, fig.width = 7, fig.height = 5, eval=F}
trunc_axis <- ggh4x::guide_axis_truncated(
  trunc_lower = unit(0, "npc"),
  trunc_upper = unit(2.2, "cm")
)

# Assign the female High_low stage of the projected cell to the nearest cell in the reference dataset
# msc_qcd_anotd_mali_mapped14HL = rep(list(rep("Not assigned", ncol(msc_combi.sce))), length(all_msc_jc))
msc_qcd_anotd_mali_mapped14HL = rep(list(rep("Not assigned", ncol(v3labmscs.ref.sce))), length(all_msc_jc))


for (i in 1:length(msc_qcd_anotd_mali_mapped14HL)){
  msc_qcd_anotd_mali_mapped14HL[[i]][all_msc_jc[[i]]$Stage_MSC$cells[1,]] <- all_msc_anotd_preQC[[i]]$StagePrelim
}


for (i in 1:length(all_msc_anotd_preQC)){
  msc_qcd_anotd_mali_mapped14HL[[i]][all_msc_jc[[i]]$Stage_MSC$cells[1,]] <- all_msc_anotd_preQC[[i]]$StagePrelim
  # msc_combi@meta.data[, paste0("stage_", names(all_msc_anotd_preQC[i]))] <-  droplevels(factor(msc_qcd_anotd_mali_mapped14HL[[i]], levels = stg_refnd_lvls))
  v3labmscs.ref.sce@colData[, paste0("stage_", names(all_msc_anotd_preQC[i]))] <-  droplevels(factor(msc_qcd_anotd_mali_mapped14HL[[i]], levels = stg_refnd_lvls))
  
}

labs_order <- c('Late ring', 'Early trophozoite','Female (HE)',  'Male (HE)', 'Female (LE)', 'Male (LE)','Gametocyte (developing)', 'Gametocyte (male)','Gametocyte (female)',  'Atlas')
names(labs_order)<- labs_order

umap_theme <- theme(legend.position = 'right',
                    axis.title=element_text(size=12,face="bold",hjust = 0),
                    axis.text=element_text(size=18), 
                    plot.title = element_blank(),
                    legend.title=element_text(size=14,face="bold"),
                    legend.text=element_text(size=14),
                    legend.spacing.y = unit(0.25, "cm"),
                    axis.line = element_line(arrow = arrow(angle = 20, length = unit(0.15, "inches"), ends = "last", type = "closed"))
)  


# Idents(msc_combi) <- 'MSC14_V3'
# m14_scmap_mca_umap <- map(names(all_msc_anotd_preQC), ~DimPlot(msc_combi, reduction = "umap",  group.by = paste0("stage_",.x), pt.size = 1, dims = c(2,3), 
#         order = labs_order) +
m14_scmap_mca_umap <- map(names(all_msc_anotd_preQC), ~{
  cbind(reducedDims(v3labmscs.ref.sce)$PCA, colData(v3labmscs.ref.sce)[, c("stageHL_ref", paste0("stage_",.x)), drop=F]) %>% 
    data.frame() %>% 
    mutate(od_stg = droplevels(factor(!!rlang::sym(paste0("stage_",.x)), levels = rev(stg_refnd_lvls)))) %>%
    arrange(od_stg) %>%
    ggplot(aes(y = PC2, x=PC3, color = od_stg)) +
    geom_point() +
    # scale_color_manual(breaks=labs_order, values = sp_fld_colors, labels = str_wrap(labs_order, 10))+
    scale_color_manual(values = sp_fld_colors)+
    umap_theme+
    guides(color = guide_legend(title = 'Stage',override.aes = list(size = 4), ncol = 1, byrow = TRUE),
           x= trunc_axis,
           y= trunc_axis)+
    scale_x_continuous(breaks = NULL) +
    scale_y_continuous(breaks = NULL)
})

##Save image for manuscript fig 4
# saveRDS(m14_scmap_mca_umap, 'plots/fig4/m14_scmap_mca_umap')

m14_scmap_mca_umap


```

Umap with stage labels inside
```{r, fig.width = 7, fig.height = 5, eval=F}

m14_scmap_mca_umap_labs_in <- map(names(all_msc_anotd_preQC), ~DimPlot(msc_combi, reduction = "umap",  group.by = paste0("stage_",.x), pt.size = 1, dims = c(2,3), label = T, order = labs_order, repel = T, label.size = 5) +
                                    scale_color_manual(values = sp_fld_colors)+ 
                                    umap_theme+
                                    theme(axis.title=element_text(size=18,face="bold"), 
                                          plot.title = element_blank(),
                                          legend.position='None'
                                    ) +
                                    guides(color = guide_legend(title = 'Stage',override.aes = list(size = 4), ncol = 1, byrow = TRUE),
                                           x= trunc_axis,
                                           y= trunc_axis)+
                                    scale_x_continuous(breaks = NULL) +
                                    scale_y_continuous(breaks = NULL)
)

##Save image for manuscript fig 4
# saveRDS(m14_scmap_mca_umap_labs_in, 'plots/fig4/m14_scmap_mca_umap_labs_in.RDS')


m14_scmap_mca_umap_labs_in


```


```{r, warning=FALSE, message=FALSE}

map(all_msc_anotd_preQC, ~..1@meta.data %>% dplyr::count(StagePrelim, Stage_MSC, stage_lab1, stage_lab2))

```


```{r, warning=FALSE, message=FALSE}
saveRDS(all_msc_anotd_preQC, "data/processed/Pf/all_msc_anotd_preQC.RDS")
```


```{r, warning=FALSE, message=FALSE}

map(all_msc_anotd_preQC, ~..1@meta.data) %>% saveRDS(., "data/processed/Pf/all_msc_anotd_preQC_mdata.RDS")

```

```{r, warning=FALSE, message=FALSE}

map(all_msc_anotd_preQC, ~..1@meta.data %>% dplyr::count(StagePrelim))

```



```{r, fig.width = 12, fig.height = 4}

##Plotting top two scmap stages among the top 10 with their similarities for MSC14
# 
# stage_fct_labs <- msc_qcd_anotd[[1]]@meta.data  %>% dplyr::group_by(stageHL, seurat_clusters) %>% dplyr::group_keys() %>% dplyr::mutate(across(c(stageHL, seurat_clusters), as.character), across(seurat_clusters, ~paste0("Cluster ",., ':\n',stageHL))) %>% deframe()
# 
# scm_barpt <- msc_qcd_anotd[[1]]@meta.data %>% 
#   dplyr::select(matches('s.*lab.'), stageHL) %>% 
#   rownames_to_column('cell_bc')%>%
#   pivot_longer(-c('cell_bc', 'stageHL'),
#                names_to = c( ".value", "Label"),
#                names_pattern = "(s.*)_(l.*)")%>% 
#   filter(!is.na(sim)) %>%
#   mutate(across(sim, as.numeric)) %>% 
#   mutate(across(Label, ~str_replace(., 'lab','Label '))) %>%
#   mutate(across(stage, ~droplevels(factor(., stg_refnd_lvls)))) %>%
#   ggplot(., aes(y = sim, x = stage, color = Label)) +
#   geom_boxplot(aes(fill = Label))+
#   scale_fill_manual(values =  c('Label 1' = '#007FFF', 'Label 2' = '#FFFF9F'), name = "Label",
#                     guide = guide_legend(direction = "horizontal", title.position = "top", ncol = 2)
#   )+
#   scale_color_manual(values =  c('Label 1' = 'black', 'Label 2' = 'black'), name = "Label")+ 
#   geom_point(aes(group=Label), size = 0.5, position = position_jitterdodge(jitter.width = 0.2))+
#   new_scale_fill()+
#   geom_tile(data = data.frame(x = c( 'Late ring', 'Early trophozoite', 'Gametocyte (developing)', 'Gametocyte (female)', 'Gametocyte (male)')), inherit.aes = F, aes(x = x, y = 0.37, fill = x, height = 0.03)) +
#   scale_fill_manual(values =  sp_fld_colors, name = "MCA stage",
#                     guide = guide_legend(direction = "horizontal", title.position = "top", ncol = 3)
#                     # guide = 'none'
#   )+
#   scale_y_continuous(expand = expansion(mult = c(0.01, 0.05))) +
#   facet_grid(.~stageHL, labeller = labeller(stageHL = stage_fct_labs)) +
#   geom_hline(yintercept = 0.4)+
#   labs(y = 'scmap \nsimilarity scores', x = 'MCA stage') +
#   theme_classic() +
#   theme(axis.text.x = element_blank(),
#         axis.title=element_text(size=16,face="plain"), 
#         axis.text = element_text(size=15), 
#         axis.ticks.x = element_blank(),
#         strip.text =element_text(size=16,face="plain", colour = 'black'),
#         panel.border=element_rect(colour="black",fill='transparent'),
#         legend.title = element_text(size=16,face="bold"),
#         legend.text = element_text(size=15),
#         legend.position = 'bottom')
# 
# scm_barpt
```
