---
title: "CellBender Learning Curve Assessment Analysis"
author: "jr35"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
    theme: flatly
    fig_width: 12
    fig_height: 8
params:
  filter_samples: NULL
---

```{r load-libraries}
# library(tidyverse)
library(gt)
library(plotly)
library(DT)
library(scales)
library(patchwork)
library(viridis)
library(conflicted)
library(readr)
library(dplyr)
library(stringr)
library(tidyr)
library(ggplot2)
conflicts_prefer(plotly::layout)
conflicts_prefer(dplyr::filter)

# Set theme
theme_set(theme_minimal(base_size = 12))

# Define color palette for assessment categories
ASSESSMENT_COLORS <- c(
  "Normal" = "#4CAF50",                # Green
  "Maybe Needs Lower LR" = "#FF9800",  # Orange
  "Needs Lower LR" = "#F44336",        # Red
  "Not Found" = "#FFC107",             # Yellow
  "Other" = "#9E9E9E"                  # Grey
)

# Define reusable theme elements
THEME_TITLE_CENTERED <- theme(
  plot.title = element_text(hjust = 0.5, face = "bold"),
  plot.subtitle = element_text(hjust = 0.5)
)

THEME_FACET_STRIP <- theme(
  strip.text = element_text(face = "bold", size = 12),
  strip.background = element_rect(fill = "#e0e0e0", color = NA)
)

THEME_LEGEND_TOP <- theme(legend.position = "top")
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.align = 'center'
)

knitr::opts_knit$set(root.dir = '/lustre/scratch126/tol/teams/lawniczak/users/jr35/phd/Mali2')

```


```{r, eval=T}
pf_solo_mixed_mdata_decode <- read_csv("data/raw/pf_all_id_decode.csv")
pf_solo_mixed_mdata_decode$sample_nm
```


```{r}
# Select sample names, remove poor quality and mixed infection samples
slct_sample_names <- pf_solo_mixed_mdata_decode$sample_nm

## poor quality samples
sample_rm <- c("MSC1", "MSC24", "MSC36", "MSC37", "MSC38", "MSC39")
sample_mxd <- pf_solo_mixed_mdata_decode %>% filter(mixed_infection2 == "mixed") %>% pull(sample_nm)
# slct_sample_names <- slct_sample_names[!(slct_sample_names %in% c(sample_rm, sample_mxd))]

all_samples2rm <- c(sample_rm, sample_mxd)
all_samples2rm <- c(all_samples2rm, paste0(all_samples2rm, "C"), paste0(all_samples2rm, "noJC"))
slct_sample_names <- slct_sample_names[!slct_sample_names %in% all_samples2rm]

# Filter metadata for selected samples
pf_solo_mixed_mdata_decode_slct <- pf_solo_mixed_mdata_decode %>% filter(sample_nm %in% slct_sample_names)

# Sorted sample names and irods IDs
sample_name <- str_sort(pf_solo_mixed_mdata_decode_slct$sample_nm)
```

```{r}

# ==============================================================================
# CONFIGURATION: Filter by specific samples
# ==============================================================================
# Set to NULL to include all samples, or provide a vector of sample names
# 
# Examples:
#   FILTER_SAMPLES <- NULL                              # All samples (default)
#   FILTER_SAMPLES <- c("MSC3", "MSC40")               # Only MSC3 and MSC40
#   FILTER_SAMPLES <- c("MSC36C", "MSC48C", "MSC49C")  # Multiple samples
#   FILTER_SAMPLES <- c("MSC50_SLOnoJC")               # Single sample
#
# You can also use grep patterns to select samples:
#   all_samples <- c("MSC36C", "MSC40noJC", "MSC48C", "MSC49C", "MSC50_SLOnoJC", ...)
#   FILTER_SAMPLES <- all_samples[grep("noJC", all_samples)]  # All samples with "noJC"

FILTER_SAMPLES <- sample_name  # Change this to filter specific samples
# ==============================================================================
```

```{r, eval=T}
# Run like 
# module load HGI/softpack/groups/team152/vietnam_vacxin/5

# Rscript extract_cellbender_learning_curve_assessment.R "/lustre/scratch126/tol/teams/lawniczak/users/jr35/phd/Mali2/data/processed/Pf/MSC*/cbender_custom2_wHsPf/cb_LR*/cb_report.html" "/lustre/scratch126/tol/teams/lawniczak/users/jr35/phd/Mali2/data/processed/Pf/m2_all/cbender_curve_assesment/wHsPf"

# Rscript extract_cellbender_learning_curve_assessment.R "/lustre/scratch126/tol/teams/lawniczak/users/jr35/phd/Mali2/data/processed/Pf/MSC*/cbender_custom2_rmHs/cb_LR*/cb_report.html" "/lustre/scratch126/tol/teams/lawniczak/users/jr35/phd/Mali2/data/processed/Pf/m2_all/cbender_curve_assesment/rmHs"
```

# Overview

This document analyzes CellBender learning curve assessments across different samples and conditions.

```{r overview-note, results='asis', echo=FALSE}
if (!is.null(FILTER_SAMPLES)) {
  cat("\n> **Note:** This report is filtered to show only the following samples: **", 
      paste(FILTER_SAMPLES, collapse = ", "), "**\n\n")
}
```

# Load Data

```{r load-data}
# Function to load CSV files
load_assessment_data <- function(file_path) {
  if (file.exists(file_path)) {
    read_csv(file_path, show_col_types = FALSE) %>%
      mutate(
        Learning_Rate = as.numeric(Learning_Rate),
        Epochs = as.numeric(Epochs)
      )
  } else {
    message(paste("File not found:", file_path))
    return(NULL)
  }
}

# Try to load both conditions
rmHs_file <- "/lustre/scratch126/tol/teams/lawniczak/users/jr35/phd/Mali2/data/processed/Pf/m2_all/cbender_curve_assesment/rmHs/cellbender_learning_curve_assessments_rmHs.csv"
wHsPf_file <- "/lustre/scratch126/tol/teams/lawniczak/users/jr35/phd/Mali2/data/processed/Pf/m2_all/cbender_curve_assesment/wHsPf/cellbender_learning_curve_assessments_wHsPf.csv"

# Load available files
data_list <- list()

if (file.exists(rmHs_file)) {
  data_list$rmHs <- load_assessment_data(rmHs_file) %>%
    mutate(Condition = "rmHs")
}

if (file.exists(wHsPf_file)) {
  data_list$wHsPf <- load_assessment_data(wHsPf_file) %>%
    mutate(Condition = "wHsPf")
}

# Combine data if multiple conditions exist
if (length(data_list) > 0) {
  combined_data <- bind_rows(data_list)
  message(paste("Loaded", nrow(combined_data), "assessments from", length(data_list), "condition(s)"))
  
  # Apply sample filter if specified
  if (!is.null(FILTER_SAMPLES)) {
    original_count <- nrow(combined_data)
    original_samples <- length(unique(combined_data$Sample))
    
    combined_data <- combined_data %>%
      filter(Sample %in% FILTER_SAMPLES)
    
    filtered_count <- nrow(combined_data)
    filtered_samples <- length(unique(combined_data$Sample))
    
    if (filtered_count == 0) {
      stop(paste("No data found for specified samples:", paste(FILTER_SAMPLES, collapse = ", ")))
    }
    
    message(paste("\n*** SAMPLE FILTER APPLIED ***"))
    message(paste("Requested samples:", paste(FILTER_SAMPLES, collapse = ", ")))
    message(paste("Found samples:", paste(unique(combined_data$Sample), collapse = ", ")))
    message(paste("Filtered from", original_count, "to", filtered_count, "assessments"))
    message(paste("Filtered from", original_samples, "to", filtered_samples, "samples"))
  }
} else {
  stop("No data files found. Please run the extraction script first.")
}

# Display loaded data info
cat("\n**Data Summary:**\n")
if (!is.null(FILTER_SAMPLES)) {
  cat("- **FILTERED SAMPLES:**", paste(FILTER_SAMPLES, collapse = ", "), "\n")
}
cat("- Total assessments:", nrow(combined_data), "\n")
cat("- Conditions:", paste(unique(combined_data$Condition), collapse = ", "), "\n")
cat("- Samples:", length(unique(combined_data$Sample)), "\n")
cat("- Learning rates:", paste(unique(combined_data$Learning_Rate), collapse = ", "), "\n")
cat("- Epochs:", paste(unique(combined_data$Epochs), collapse = ", "), "\n")
```

# Data Table


```{r data-table}
# Remove PNG hyperlink column, just show the main assessment table
table_data <- combined_data %>%
  select(Sample, Condition, Learning_Rate, Epochs, Assessment_Category, Full_Assessment)

datatable(
  table_data,
  filter = 'top',
  options = list(
    pageLength = 20,
    scrollX = TRUE,
    autoWidth = TRUE
  ),
  rownames = FALSE,
  escape = TRUE,
  caption = "Interactive table of all CellBender learning curve assessments"
)
```

## Learning Curve Images by Sample (Grid)
```{r learning-curve-image-grid, results='asis', fig.width=10, fig.height=10}

# Show a grid of learning curve images for each sample, up to four per sample (for different parameter combos)
library(gridExtra)
img_df <- combined_data %>%
  filter(!is.na(Learning_Curve_Image)) %>%
  mutate(Exists = file.exists(Learning_Curve_Image))

if (nrow(img_df) == 0) {
  cat("No learning curve images found in the loaded data.\n")
} else {
  samples <- unique(img_df$Sample)
  for (s in samples) {
    sample_imgs <- img_df %>% filter(Sample == s & Exists)
    if (nrow(sample_imgs) == 0) next
    # Order by Condition, Learning_Rate, Epochs for consistency
    sample_imgs <- sample_imgs %>% arrange(Condition, Learning_Rate, Epochs)
    # Titles for each image with assessment category
    img_titles <- paste0(sample_imgs$Condition, ", LR=", sample_imgs$Learning_Rate, ", E=", sample_imgs$Epochs, 
                         "\n[", sample_imgs$Assessment_Category, "]")
    # List of grobs for images
    grob_list <- lapply(seq_len(nrow(sample_imgs)), function(i) {
      grid::rasterGrob(png::readPNG(sample_imgs$Learning_Curve_Image[i]), interpolate=TRUE)
    })
    # Arrange in up to 2x2 grid
    n_imgs <- length(grob_list)
    ncol <- ifelse(n_imgs > 2, 2, n_imgs)
    nrow <- ceiling(n_imgs / ncol)
    # Add titles above each image with assessment category
    grid_plots <- mapply(function(g, t) gridExtra::arrangeGrob(g, top = t), grob_list, img_titles, SIMPLIFY = FALSE)
    # Title for the sample
    cat(paste0("\n\n### ", s, "\n"))
    # grid::grid.newpage()
    gridExtra::grid.arrange(grobs = grid_plots, ncol = ncol, top = grid::textGrob(s, gp = grid::gpar(fontsize = 16, fontface = "bold")))
  }
}
```

## Learning Curve Images

```{r learning-curve-images, results='asis', eval=F}
# Display extracted learning curve PNGs, if present
img_df <- combined_data %>%
  filter(!is.na(Learning_Curve_Image)) %>%
  mutate(Exists = file.exists(Learning_Curve_Image)) %>%
  arrange(Condition, Sample, Learning_Rate, Epochs)

if (nrow(img_df) == 0) {
  cat("No learning curve images found in the loaded data.\n")
} else {
  for (i in seq_len(nrow(img_df))) {
    row <- img_df[i, ]
    cat(paste0("\n### ", row$Sample, " | ", row$Condition, " | LR=", row$Learning_Rate, " | E=", row$Epochs, "\n\n"))
    if (isTRUE(row$Exists)) {
      print(knitr::include_graphics(row$Learning_Curve_Image))
    } else {
      cat("Image file not found: ", row$Learning_Curve_Image, "\n\n")
    }
  }
}
```

# Summary Statistics

```{r summary-stats}
# Overall summary
overall_summary <- combined_data %>%
  count(Assessment_Category) %>%
  mutate(
    Percentage = n / sum(n) * 100,
    Percentage_Label = sprintf("%.1f%%", Percentage)
  )

# Create summary table
overall_summary %>%
  gt() %>%
  tab_header(
    title = "Overall Assessment Categories",
    subtitle = "Across all samples and conditions"
  ) %>%
  cols_label(
    Assessment_Category = "Assessment",
    n = "Count",
    Percentage = "Percentage",
    Percentage_Label = "Percentage"
  ) %>%
  fmt_number(
    columns = c(Percentage),
    decimals = 1
  ) %>%
  cols_hide(columns = c(Percentage_Label)) %>%
  tab_style(
    style = cell_fill(color = "#f0f0f0"),
    locations = cells_body(columns = everything(), rows = Assessment_Category == "Normal")
  ) %>%
  tab_style(
    style = cell_fill(color = "#fff3e0"),
    locations = cells_body(columns = everything(), rows = Assessment_Category == "Maybe Needs Lower LR")
  ) %>%
  tab_style(
    style = cell_fill(color = "#ffe6e6"),
    locations = cells_body(columns = everything(), rows = Assessment_Category == "Needs Lower LR")
  )
```

```{r condition-summary}
# Summary by condition
if (length(unique(combined_data$Condition)) > 1) {
  combined_data %>%
    count(Condition, Assessment_Category) %>%
    group_by(Condition) %>%
    mutate(
      Total = sum(n),
      Percentage = n / Total * 100
    ) %>%
    ungroup() %>%
    gt() %>%
    tab_header(
      title = "Assessment Categories by Condition",
      subtitle = "Comparison across processing conditions"
    ) %>%
    cols_label(
      Condition = "Condition",
      Assessment_Category = "Assessment",
      n = "Count",
      Total = "Total",
      Percentage = "Percentage"
    ) %>%
    fmt_number(
      columns = c(Percentage),
      decimals = 1
    ) %>%
    tab_style(
      style = cell_fill(color = "#f0f0f0"),
      locations = cells_body(columns = everything(), rows = Assessment_Category == "Normal")
    ) %>%
    tab_style(
      style = cell_fill(color = "#fff3e0"),
      locations = cells_body(columns = everything(), rows = Assessment_Category == "Maybe Needs Lower LR")
    ) %>%
    tab_style(
      style = cell_fill(color = "#ffe6e6"),
      locations = cells_body(columns = everything(), rows = Assessment_Category == "Needs Lower LR")
    )
}
```

# Visualizations

## Assessment Category Distribution

```{r assessment-distribution, fig.height=6}
# Pie chart of overall distribution
p1 <- overall_summary %>%
  ggplot(aes(x = "", y = n, fill = Assessment_Category)) +
  geom_bar(stat = "identity", width = 1, color = "white") +
  coord_polar("y", start = 0) +
  geom_text(aes(label = paste0(Assessment_Category, "\n", n, " (", Percentage_Label, ")")),
            position = position_stack(vjust = 0.5),
            size = 4) +
  scale_fill_manual(values = ASSESSMENT_COLORS) +
  labs(
    title = "Overall Assessment Distribution",
    fill = "Assessment"
  ) +
  theme_void() +
  THEME_TITLE_CENTERED +
  theme(
    plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
    legend.position = "right"
  )

# Bar chart
p2 <- combined_data %>%
  count(Assessment_Category) %>%
  ggplot(aes(x = reorder(Assessment_Category, -n), y = n, fill = Assessment_Category)) +
  geom_col(color = "white") +
  geom_text(aes(label = n), vjust = -0.5, size = 4) +
  scale_fill_manual(values = ASSESSMENT_COLORS) +
  labs(
    title = "Assessment Category Counts",
    x = "Assessment Category",
    y = "Count"
  ) +
  THEME_TITLE_CENTERED +
  theme(legend.position = "none")

p1 + p2
```

## Assessment by Condition

```{r assessment-by-condition, fig.height=6}
if (length(unique(combined_data$Condition)) > 1) {
  combined_data %>%
    count(Condition, Assessment_Category) %>%
    ggplot(aes(x = Condition, y = n, fill = Assessment_Category)) +
    geom_col(position = "dodge", color = "white") +
    geom_text(aes(label = n), position = position_dodge(width = 0.9), vjust = -0.5) +
    scale_fill_manual(values = ASSESSMENT_COLORS) +
    labs(
      title = "Assessment Categories by Processing Condition",
      x = "Condition",
      y = "Count",
      fill = "Assessment"
    ) +
    THEME_TITLE_CENTERED +
    THEME_LEGEND_TOP
} else {
  cat("Only one condition present in the data.\n")
}
```

## Assessment by Learning Rate

```{r assessment-by-lr, fig.height=6}
if (length(unique(combined_data$Condition)) > 1) {
  # Plot with condition split
  combined_data %>%
    count(Condition, Learning_Rate, Assessment_Category) %>%
    ggplot(aes(x = factor(Learning_Rate), y = n, fill = Assessment_Category)) +
    geom_col(position = "stack", color = "white") +
    geom_text(aes(label = n), position = position_stack(vjust = 0.5), size = 3) +
    scale_fill_manual(values = ASSESSMENT_COLORS) +
    facet_wrap(~Condition, ncol = 2) +
    labs(
      title = "Assessment Categories by Learning Rate and Condition",
      x = "Learning Rate",
      y = "Count",
      fill = "Assessment"
    ) +
    THEME_TITLE_CENTERED +
    THEME_LEGEND_TOP +
    THEME_FACET_STRIP
} else {
  # Single condition plot
  combined_data %>%
    count(Learning_Rate, Assessment_Category) %>%
    ggplot(aes(x = factor(Learning_Rate), y = n, fill = Assessment_Category)) +
    geom_col(position = "stack", color = "white") +
    geom_text(aes(label = n), position = position_stack(vjust = 0.5), size = 3.5) +
    scale_fill_manual(values = ASSESSMENT_COLORS) +
    labs(
      title = "Assessment Categories by Learning Rate",
      x = "Learning Rate",
      y = "Count",
      fill = "Assessment"
    ) +
    THEME_TITLE_CENTERED +
    THEME_LEGEND_TOP
}
```

## Assessment by Epochs

```{r assessment-by-epochs, fig.height=6}
if (length(unique(combined_data$Condition)) > 1) {
  # Plot with condition split
  combined_data %>%
    count(Condition, Epochs, Assessment_Category) %>%
    ggplot(aes(x = factor(Epochs), y = n, fill = Assessment_Category)) +
    geom_col(position = "stack", color = "white") +
    geom_text(aes(label = n), position = position_stack(vjust = 0.5), size = 3) +
    scale_fill_manual(values = ASSESSMENT_COLORS) +
    facet_wrap(~Condition, ncol = 2) +
    labs(
      title = "Assessment Categories by Number of Epochs and Condition",
      x = "Epochs",
      y = "Count",
      fill = "Assessment"
    ) +
    THEME_TITLE_CENTERED +
    THEME_LEGEND_TOP +
    THEME_FACET_STRIP
} else {
  # Single condition plot
  combined_data %>%
    count(Epochs, Assessment_Category) %>%
    ggplot(aes(x = factor(Epochs), y = n, fill = Assessment_Category)) +
    geom_col(position = "stack", color = "white") +
    geom_text(aes(label = n), position = position_stack(vjust = 0.5), size = 3.5) +
    scale_fill_manual(values = ASSESSMENT_COLORS) +
    labs(
      title = "Assessment Categories by Number of Epochs",
      x = "Epochs",
      y = "Count",
      fill = "Assessment"
    ) +
    THEME_TITLE_CENTERED +
    THEME_LEGEND_TOP
}
```

## Heatmap: Learning Rate vs Epochs

```{r heatmap-lr-epochs, fig.height=8}
# Create a heatmap showing assessment categories
heatmap_data <- combined_data %>%
  count(Condition, Learning_Rate, Epochs, Assessment_Category) %>%
  group_by(Condition, Learning_Rate, Epochs) %>%
  mutate(
    Total = sum(n),
    Category_Pct = n / Total * 100
  ) %>%
  ungroup()

if (length(unique(combined_data$Condition)) > 1) {
  # Plot for "Normal" assessments - split by condition
  heatmap_data %>%
    filter(Assessment_Category == "Normal") %>%
    ggplot(aes(x = factor(Learning_Rate), y = factor(Epochs), fill = Category_Pct)) +
    geom_tile(color = "white") +
    geom_text(aes(label = sprintf("%.0f%%\n(%d/%d)", Category_Pct, n, Total)), size = 3) +
    scale_fill_gradient(low = "#fff5f0", high = "#4CAF50", name = "% Normal") +
    facet_wrap(~Condition, ncol = 2) +
    labs(
      title = "Percentage of 'Normal' Learning Curve Assessments",
      subtitle = "By Learning Rate, Epochs, and Condition",
      x = "Learning Rate",
      y = "Epochs"
    ) +
    THEME_TITLE_CENTERED +
    THEME_FACET_STRIP
} else {
  # Single condition heatmap
  heatmap_data %>%
    filter(Assessment_Category == "Normal") %>%
    ggplot(aes(x = factor(Learning_Rate), y = factor(Epochs), fill = Category_Pct)) +
    geom_tile(color = "white") +
    geom_text(aes(label = sprintf("%.0f%%\n(%d/%d)", Category_Pct, n, Total)), size = 3) +
    scale_fill_gradient(low = "#fff5f0", high = "#4CAF50", name = "% Normal") +
    labs(
      title = "Percentage of 'Normal' Learning Curve Assessments",
      subtitle = "By Learning Rate and Epochs",
      x = "Learning Rate",
      y = "Epochs"
    ) +
    THEME_TITLE_CENTERED
}
```


```{r, fig.height= 4, fig.width= 4, eval=T}
# Copy to 'cell_calling_figs.Rmd' for figure generation
combined_data_cbender_assesment <- combined_data
saveRDS(combined_data_cbender_assesment, "data/processed/Pf/m2_all/cell_calling_plotting/combined_data_cbender_assesment.RDS")

```

## Sample-Level Assessment

```{r sample-assessment, fig.height=12, fig.width=14}
# Summary by sample with parameter details
sample_summary <- combined_data %>%
  count(Sample, Condition, Learning_Rate, Epochs, Assessment_Category) %>%
  mutate(
    Param_Combo = paste0("LR: ", Learning_Rate, "\nE: ", Epochs)
  )

if (length(unique(combined_data$Condition)) > 1) {
  # Plot with condition and parameter combinations
  sample_summary %>%
    ggplot(aes(x = Sample, y = n, fill = Assessment_Category)) +
    geom_col(position = "stack", color = "white", width = 0.7) +
    geom_text(aes(label = ifelse(n > 0, n, "")), 
              position = position_stack(vjust = 0.5), size = 2) +
    scale_fill_manual(values = ASSESSMENT_COLORS) +
    facet_grid(Condition ~ Param_Combo, scales = "free_y", space = "free") +
    labs(
      title = "Assessment Categories by Sample, Condition, and Parameters",
      subtitle = "Each panel shows a different learning rate and epoch combination",
      x = "Sample",
      y = "Count",
      fill = "Assessment"
    ) +
    coord_flip() +
    theme_minimal(base_size = 10) +
    THEME_TITLE_CENTERED +
    THEME_LEGEND_TOP +
    theme(
      strip.text = element_text(face = "bold", size = 9),
      strip.background = element_rect(fill = "#e0e0e0", color = NA),
      axis.text.x = element_text(size = 8),
      axis.text.y = element_text(size = 7),
      panel.spacing = unit(0.5, "lines")
    )
} else {
  # Single condition plot with parameter facets
  sample_summary %>%
    ggplot(aes(x = Sample, y = n, fill = Assessment_Category)) +
    geom_col(position = "stack", color = "white", width = 0.7) +
    geom_text(aes(label = ifelse(n > 0, n, "")), 
              position = position_stack(vjust = 0.5), size = 2.5) +
    scale_fill_manual(values = ASSESSMENT_COLORS) +
    facet_wrap(~ Param_Combo, ncol = 2, scales = "free_y") +
    labs(
      title = "Assessment Categories by Sample and Parameters",
      subtitle = "Each panel shows a different learning rate and epoch combination",
      x = "Sample",
      y = "Count",
      fill = "Assessment"
    ) +
    coord_flip() +
    theme_minimal(base_size = 10) +
    THEME_TITLE_CENTERED +
    THEME_LEGEND_TOP +
    theme(
      strip.text = element_text(face = "bold", size = 10),
      strip.background = element_rect(fill = "#e0e0e0", color = NA)
    )
}
```

## Interactive Sample Exploration

```{r interactive-sample}
if (length(unique(combined_data$Condition)) > 1) {
  # Interactive plot with condition faceting
  plot_ly(
    data = sample_summary,
    x = ~Sample,
    y = ~n,
    color = ~Assessment_Category,
    colors = ASSESSMENT_COLORS,
    type = "bar",
    text = ~paste(
      "Sample:", Sample,
      "<br>Condition:", Condition,
      "<br>Learning Rate:", Learning_Rate,
      "<br>Epochs:", Epochs,
      "<br>Assessment:", Assessment_Category,
      "<br>Count:", n
    ),
    hoverinfo = "text",
    transforms = list(
      list(
        type = 'groupby',
        groups = ~Condition
      )
    )
  ) %>%
    layout(
      title = "Interactive Sample-Level Assessment by Condition",
      barmode = "stack",
      xaxis = list(title = "Sample"),
      yaxis = list(title = "Count"),
      legend = list(title = list(text = "Assessment")),
      updatemenus = list(
        list(
          type = "buttons",
          direction = "down",
          xanchor = "left",
          yanchor = "top",
          x = 0,
          y = 1.15,
          buttons = list(
            list(method = "restyle",
                 args = list("transforms[0].value", unique(sample_summary$Condition)[1]),
                 label = unique(sample_summary$Condition)[1]),
            list(method = "restyle",
                 args = list("transforms[0].value", unique(sample_summary$Condition)[2]),
                 label = unique(sample_summary$Condition)[2])
          )
        )
      )
    )
} else {
  # Single condition interactive plot
  plot_ly(
    data = sample_summary,
    x = ~Sample,
    y = ~n,
    color = ~Assessment_Category,
    colors = ASSESSMENT_COLORS,
    type = "bar",
    text = ~paste(
      "Sample:", Sample,
      "<br>Learning Rate:", Learning_Rate,
      "<br>Epochs:", Epochs,
      "<br>Assessment:", Assessment_Category,
      "<br>Count:", n
    ),
    hoverinfo = "text"
  ) %>%
    layout(
      title = "Interactive Sample-Level Assessment",
      barmode = "stack",
      xaxis = list(title = "Sample"),
      yaxis = list(title = "Count"),
      legend = list(title = list(text = "Assessment"))
    )
}
```

# Detailed Analysis

## Parameter Combinations

```{r parameter-combinations}
if (length(unique(combined_data$Condition)) > 1) {
  # Show parameter combinations by condition
  param_combos <- combined_data %>%
    count(Condition, Learning_Rate, Epochs, Assessment_Category) %>%
    pivot_wider(
      names_from = Assessment_Category,
      values_from = n,
      values_fill = 0
    ) %>%
    mutate(
      Total = rowSums(across(where(is.numeric) & !matches("Condition"))),
      Normal_Pct = if_else(Total > 0, Normal / Total * 100, 0)
    ) %>%
    arrange(Condition, desc(Normal_Pct))
  
  param_combos %>%
    gt() %>%
    tab_header(
      title = "Parameter Combinations Performance by Condition",
      subtitle = "Sorted by percentage of Normal assessments within each condition"
    ) %>%
    cols_label(
      Condition = "Condition",
      Learning_Rate = "Learning Rate",
      Epochs = "Epochs",
      Normal_Pct = "% Normal"
    ) %>%
    fmt_number(
      columns = c(Normal_Pct),
      decimals = 1
    ) %>%
    data_color(
      columns = c(Normal_Pct),
      colors = scales::col_numeric(
        palette = c("#F44336", "#FFC107", "#4CAF50"),
        domain = c(0, 100)
      )
    ) %>%
    tab_style(
      style = cell_fill(color = "#f5f5f5"),
      locations = cells_body(columns = everything(), rows = Condition == "rmHs")
    )
} else {
  # Single condition
  param_combos <- combined_data %>%
    count(Learning_Rate, Epochs, Assessment_Category) %>%
    pivot_wider(
      names_from = Assessment_Category,
      values_from = n,
      values_fill = 0
    ) %>%
    mutate(
      Total = rowSums(across(where(is.numeric))),
      Normal_Pct = if_else(Total > 0, Normal / Total * 100, 0)
    ) %>%
    arrange(desc(Normal_Pct))
  
  param_combos %>%
    gt() %>%
    tab_header(
      title = "Parameter Combinations Performance",
      subtitle = "Sorted by percentage of Normal assessments"
    ) %>%
    cols_label(
      Learning_Rate = "Learning Rate",
      Epochs = "Epochs",
      Normal_Pct = "% Normal"
    ) %>%
    fmt_number(
      columns = c(Normal_Pct),
      decimals = 1
    ) %>%
    data_color(
      columns = c(Normal_Pct),
      colors = scales::col_numeric(
        palette = c("#F44336", "#FFC107", "#4CAF50"),
        domain = c(0, 100)
      )
    )
}
```

## Samples Needing Attention

```{r samples-attention}
# Identify samples with "Needs Lower LR" or "Maybe Needs Lower LR" assessments
samples_needs_lr <- combined_data %>%
  filter(Assessment_Category == "Needs Lower LR") %>%
  select(Sample, Condition, Learning_Rate, Epochs, Full_Assessment) %>%
  arrange(Sample, Learning_Rate, Epochs)

samples_maybe_needs_lr <- combined_data %>%
  filter(Assessment_Category == "Maybe Needs Lower LR") %>%
  select(Sample, Condition, Learning_Rate, Epochs, Full_Assessment) %>%
  arrange(Sample, Learning_Rate, Epochs)

if (nrow(samples_needs_lr) > 0) {
  cat("\n**Samples with 'Needs Lower LR' assessments (definite):**\n\n")
  
  samples_needs_lr %>%
    gt() %>%
    tab_header(
      title = "Samples Definitely Requiring Lower Learning Rate",
      subtitle = "These runs showed unusual behavior - re-run recommended"
    ) %>%
    cols_label(
      Sample = "Sample",
      Condition = "Condition",
      Learning_Rate = "Learning Rate",
      Epochs = "Epochs",
      Full_Assessment = "Assessment"
    ) %>%
    tab_style(
      style = cell_fill(color = "#ffe6e6"),
      locations = cells_body()
    )
} else {
  cat("\n**No samples definitely require lower learning rates!**\n")
}

if (nrow(samples_maybe_needs_lr) > 0) {
  cat("\n\n**Samples with 'Maybe Needs Lower LR' assessments (slightly unusual):**\n\n")
  
  samples_maybe_needs_lr %>%
    gt() %>%
    tab_header(
      title = "Samples That Might Benefit from Lower Learning Rate",
      subtitle = "These runs showed slightly unusual behavior - consider re-running"
    ) %>%
    cols_label(
      Sample = "Sample",
      Condition = "Condition",
      Learning_Rate = "Learning Rate",
      Epochs = "Epochs",
      Full_Assessment = "Assessment"
    ) %>%
    tab_style(
      style = cell_fill(color = "#fff3e0"),
      locations = cells_body()
    )
} else {
  cat("\n\n**No samples show slightly unusual behavior!**\n")
}

if (nrow(samples_needs_lr) == 0 && nrow(samples_maybe_needs_lr) == 0) {
  cat("\n**All learning curves look normal! No action required.**\n")
}
```

## Recommended Parameters

```{r recommended-parameters}
if (length(unique(combined_data$Condition)) > 1) {
  # Find best performing parameter combinations per condition
  recommended <- combined_data %>%
    filter(Assessment_Category == "Normal") %>%
    count(Condition, Learning_Rate, Epochs) %>%
    group_by(Condition) %>%
    arrange(Condition, desc(n)) %>%
    slice_head(n = 3) %>%
    ungroup()
  
  cat("\n**Top 3 Parameter Combinations per Condition (by number of Normal assessments):**\n\n")
  
  recommended %>%
    gt() %>%
    tab_header(
      title = "Recommended Parameter Combinations by Condition",
      subtitle = "Based on frequency of Normal assessments"
    ) %>%
    cols_label(
      Condition = "Condition",
      Learning_Rate = "Learning Rate",
      Epochs = "Epochs",
      n = "Normal Count"
    ) %>%
    tab_style(
      style = cell_fill(color = "#e8f5e9"),
      locations = cells_body()
    ) %>%
    tab_style(
      style = cell_fill(color = "#f5f5f5"),
      locations = cells_body(columns = everything(), rows = Condition == "rmHs")
    )
} else {
  # Single condition
  recommended <- combined_data %>%
    filter(Assessment_Category == "Normal") %>%
    count(Learning_Rate, Epochs) %>%
    arrange(desc(n)) %>%
    head(5)
  
  cat("\n**Top 5 Parameter Combinations (by number of Normal assessments):**\n\n")
  
  recommended %>%
    gt() %>%
    tab_header(
      title = "Recommended Parameter Combinations",
      subtitle = "Based on frequency of Normal assessments"
    ) %>%
    cols_label(
      Learning_Rate = "Learning Rate",
      Epochs = "Epochs",
      n = "Normal Count"
    ) %>%
    tab_style(
      style = cell_fill(color = "#e8f5e9"),
      locations = cells_body()
    )
}
```

# Condition Comparison

```{r condition-comparison}
if (length(unique(combined_data$Condition)) > 1) {
  # Compare conditions side by side
  condition_comparison <- combined_data %>%
    count(Condition, Learning_Rate, Epochs, Assessment_Category) %>%
    pivot_wider(
      names_from = Assessment_Category,
      values_from = n,
      values_fill = 0
    ) %>%
    mutate(
      Total = rowSums(across(where(is.numeric))),
      Normal_Pct = if_else(Total > 0, Normal / Total * 100, 0)
    )
  
  # Plot comparison
  condition_comparison %>%
    ggplot(aes(x = interaction(Learning_Rate, Epochs), y = Normal_Pct, fill = Condition)) +
    geom_col(position = "dodge") +
    geom_text(aes(label = sprintf("%.0f%%", Normal_Pct)), 
              position = position_dodge(width = 0.9), vjust = -0.5, size = 3) +
    scale_fill_brewer(palette = "Set2") +
    labs(
      title = "Comparison of Normal Assessment Rates by Condition",
      subtitle = "Across different parameter combinations",
      x = "Learning Rate × Epochs",
      y = "% Normal Assessments",
      fill = "Condition"
    ) +
    THEME_TITLE_CENTERED +
    THEME_LEGEND_TOP +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
}
```

# Summary

```{r final-summary}
cat("\n## Key Findings\n\n")

# Calculate key metrics
total_assessments <- nrow(combined_data)
normal_count <- sum(combined_data$Assessment_Category == "Normal")
needs_lr_count <- sum(combined_data$Assessment_Category == "Needs Lower LR")
maybe_needs_lr_count <- sum(combined_data$Assessment_Category == "Maybe Needs Lower LR")
normal_pct <- normal_count / total_assessments * 100

cat("- **Total assessments analyzed:**", total_assessments, "\n")
cat("- **Normal learning curves:**", normal_count, sprintf("(%.1f%%)", normal_pct), "\n")
cat("- **Definitely needs lower learning rate:**", needs_lr_count, sprintf("(%.1f%%)", needs_lr_count / total_assessments * 100), "\n")
cat("- **Maybe needs lower learning rate:**", maybe_needs_lr_count, sprintf("(%.1f%%)", maybe_needs_lr_count / total_assessments * 100), "\n")

if (needs_lr_count > 0) {
  cat("\n**High Priority Action Items:**\n")
  cat("- Re-run", needs_lr_count, "assessments with half the current learning rate\n")
  cat("- Focus on samples:", paste(unique(samples_needs_lr$Sample), collapse = ", "), "\n")
}

if (maybe_needs_lr_count > 0) {
  cat("\n**Consider Action Items:**\n")
  cat("- Consider re-running", maybe_needs_lr_count, "assessments with half the current learning rate\n")
  cat("- Samples to consider:", paste(unique(samples_maybe_needs_lr$Sample), collapse = ", "), "\n")
}

if (needs_lr_count == 0 && maybe_needs_lr_count == 0) {
  cat("\n**All learning curves look normal! No action required.**\n")
}

if (length(unique(combined_data$Condition)) > 1) {
  cat("\n**Best performing parameters by condition:**\n")
  for (cond in unique(recommended$Condition)) {
    cat("\n", cond, ":\n", sep = "")
    best_params <- recommended %>% filter(Condition == cond) %>% slice(1)
    cat("  - Learning Rate:", best_params$Learning_Rate, "\n")
    cat("  - Epochs:", best_params$Epochs, "\n")
    cat("  - Normal count:", best_params$n, "\n")
  }
} else {
  cat("\n**Best performing parameters:**\n")
  best_params <- recommended %>% slice(1)
  cat("- Learning Rate:", best_params$Learning_Rate, "\n")
  cat("- Epochs:", best_params$Epochs, "\n")
}
```

---

**Report generated:** `r Sys.time()`

**Data files analyzed:** `r paste(names(data_list), collapse = ", ")`
